### 2014-05-10 paper/05_conclusions.tex

\section{Conclusions}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% TeX-PDF-mode: t
%%% End: