### 2014-08-11 paper/04_discussion.tex
Added short description of results. Renamed results section to discussion.

\section{Discussion}
\label{sec:discussion}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% TeX-PDF-mode: t
%%% End: