### 2014-08-11 paper/paper.tex

\documentclass[fleqn,reqno,10pt]{article}

%========================================
% Packages
%========================================

\usepackage[]{../helpers/mypackages}
%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
\bibliography{../helpers/MyRefGlobal}
\usepackage{../helpers/myenvironments}
\usepackage{../helpers/mycommands}
\usepackage{todonotes}
\usepackage{subcaption}

%========================================
% Standard Layout
%========================================

% Itemize
\renewcommand{\labelitemi}{\large{$\mathbf{\cdot}$}}    % itemize symbols
\renewcommand{\labelitemii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiv}{\large{$\mathbf{\cdot}$}}
% Description
\renewcommand{\descriptionlabel}[1]{\hspace\labelsep\textsc{#1}}

% Figure Captions
\usepackage{caption} % use corresponding myfiguresize!
\setlength{\captionmargin}{20pt}
\renewcommand{\captionfont}{\small}
\setlength{\belowcaptionskip}{7pt} % standard is 0pt

%========================================
% Additional layout & commands
%========================================

\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}

% Annotations
\newcommand{\mytodo}[2]{\todo[inline,color=yellow,author=#1]{#2}}
\newcommand{\question}[2]{\todo[inline,color=blue,author=#1]{#2}}
\newcommand{\answer}[2]{\todo[inline,color=green,author=#1]{#2}}

\title{Vagueness, Noise, and Communication}
\author{Jos\'e Pedro Correia and Michael Franke}
\date{}

\begin{document}
\maketitle

\begin{abstract}
\mytodo{JPC}{This start needs to be adjusted depending on the audience}
Signaling games have been productively used as models for studying natural language meaning.
One important aspect that is, however, typically missing from most models in the literature is the possibility for vagueness as a feature of successful languages.
Recently, some models have been proposed to explicitly address this limitation.
In this paper we explore one such models.
Namely, we establish a link between the dynamics in the model and the replicator-mutator equations.
Furthermore, we study the long term behavior of the model using computer simulations.
\mytodo{JPC}{Kind of a big claim below\ldots}
Our results indicate that a possible explanation of why vagueness is pervasive in natural language is that vague languages converge faster than non-vague ones, thus providing an evolutionary advantage in an changing environment.
\end{abstract}

\input{01_intro}

\input{02_noisyRMD}

\input{03_simulations}

\input{04_discussion}

\input{05_conclusions}

\input{06_appendix}

\printbibliography[heading=bibintoc]

\end{document}