### 2014-08-22 paper/paper.tex
Reshaped abstract.

\documentclass[fleqn,reqno,10pt]{article}

%========================================
% Packages
%========================================

\usepackage[]{../helpers/mypackages}
%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
\bibliography{../helpers/MyRefGlobal}
\bibliography{paper}
\usepackage{../helpers/myenvironments}
\usepackage{../helpers/mycommands}
\usepackage{todonotes}
\usepackage{subcaption}

%========================================
% Standard Layout
%========================================

% Itemize
\renewcommand{\labelitemi}{\large{$\mathbf{\cdot}$}}    % itemize symbols
\renewcommand{\labelitemii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiv}{\large{$\mathbf{\cdot}$}}
% Description
\renewcommand{\descriptionlabel}[1]{\hspace\labelsep\textsc{#1}}

% Figure Captions
\usepackage{caption} % use corresponding myfiguresize!
\setlength{\captionmargin}{20pt}
\renewcommand{\captionfont}{\small}
\setlength{\belowcaptionskip}{7pt} % standard is 0pt

%========================================
% Additional layout & commands
%========================================

\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}

% Annotations
\newcommand{\mytodo}[2]{\todo[inline,color=yellow,author=#1]{#2}}
\newcommand{\question}[2]{\todo[inline,color=blue,author=#1]{#2}}
\newcommand{\answer}[2]{\todo[inline,color=green,author=#1]{#2}}

\title{Vagueness, Noise, and Communication}
\author{Jos\'e Pedro Correia and Michael Franke}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Signaling games have been garnering a lot of attention as models for studying natural language meaning.
One important aspect that is nevertheless typically missing from most models in the literature is the possibility for vagueness as a feature of successful languages.
Recently, some models have been proposed to explicitly address this limitation.
In this paper we explore one such model and establish a link between the dynamics in the model and the replicator-mutator equations.
Furthermore, we study the long term behavior of the model using computer simulations.
The results indicate that vague languages, despite having lower expected utility and informativity, are faster to converge to an equilibrium than non-vague ones.
We argue that this could be an importante advantage that contributes to the pervasiveness of vagueness in natural language.
\end{abstract}

\input{01_intro}

\input{02_noisyRMD}

\input{03_simulations}

\input{04_discussion}

\input{05_conclusions}

\input{06_appendix}

\printbibliography[heading=bibintoc]

\end{document}