### 2014-08-26 data_exploration_tmp.Rmd
data exploration MF

Data Exploration
================

load and pre-process the data

``` {r}

require('ellipse')
require('ggplot2')

data = read.csv('results/20140813-194306.csv', 
                colClasses=c('numeric','factor','numeric','numeric','numeric','factor','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','character','character'))

ds = data[,c(1,4,5,7:16)] # only relevant data

attach(ds)
```

plot all the relevant variables against each other

``` {r warning = FALSE, fig.width=12, fig.height=12}

pairs(~ Number.of.states + Impairment + Speaker.entropy + Hearer.entropy +
        Speaker.convexity + Hearer.convexity +
        Speaker.informativity + Hearer.informativity + Expected.utility + Iterations, data=ds)
```

some properties are very clearly positively correlated:

* speaker informativity, hearer informativity & expected utility
* impairment & speaker entropy

these two "clusters" are also strongly negatively correlated

``` {r warning = FALSE, fig.width=7, fig.height=7}

dss = ds[,c(2,4,10,11,12)]

ctab <- cor(dss)

plotcorr(ctab)

# pairs(~ Impairment + Speaker.entropy +
#         Speaker.informativity + Hearer.informativity + Expected.utility, data=dss)
```

The benefit of impairment (some thoughts)
-------------------------

we want to argue that (small?) levels of impairment can be helpful (i.e., boost EU); but unfortunately, impairment and EU are strongly negatively correlated; so we should look for stages before final convergence

impairment does seem to speed up learning! that's different from Cailin's stuff, because there learning speed could be argued to be only due to more balls being added per learning step

also speaker and hearer voronoiness seems to be much higher for at least some bit of impairment, but we left this notion out [why what that again?]

Faster learning by impairment
-----------------------------

if learning is faster for some impairment, we should see these effects also in terms of higher EU at earlier stages

- plot EU at different stages in the nRD
- show that early on slight impairment is beneficial to no impairment

Next up
-------
- evaluate impact of number of states
- replot Fig 1 with seperated numbers of states

``` {r}

x = data

ggplot(x, aes(x=factor(Impairment), y=Expected.utility, color=factor(Number.of.states))) + geom_boxplot() + xlab('Impairment') + ylim(0, 1) + ylab(expression(EU))

```