### 2014-08-27 data_exploration_round_based.Rmd
data exploration round based

Data Exploration (round-based data)
===================================

load and pre-process the data

``` {r warnings=FALSE, message=FALSE}

require('ellipse')
require('ggplot2')
require('reshape2')
source('~//Desktop//data//svn//programming//R//helpers//helpers.R')

data = read.csv('results/20140827-124230.round_based.csv')

ds = data[,c(1,4,7:8)] # only relevant data
dm = melt(ds,id.vars = c("Number.of.states", "Impairment", "Iterations")) # melted data

attach(ds)
```

Plot EU by round
----------------

We plot EU at every fifth round, depending on number of states and impairments.

```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=10}

dms = summarySE(dm, groupvars = c("Number.of.states", "Impairment", "Iterations", "variable"), measurevar="value")
dsub = subset(dms, dms$variable=="Expected.utility")

pd <- position_dodge(.05)
sp = ggplot(dsub, aes(x=Iterations, y=value, colour=factor(Impairment))) + 
     geom_point(position = pd) +
     geom_errorbar(aes(ymin=value-ci, ymax=value+ci), width=0, position=pd) +
     geom_line(position=pd)

sp + facet_wrap(~ Number.of.states, ncol=2, scales="free")

```

Unfortunately, we see that lower impairment always trumps higher impairment.