### 2014-09-12 snippets/iteration_diffusion.Rmd
investigating iterations of diffusion only

```{r warning=FALSE, echo=FALSE, message=FALSE}
library(MCMCpack) # for Dirichlet distribution
source('~/Desktop/data/svn/programming/R/fuzzy-action/dev.r')
```

Iterated Diffusion
==================

The (discrete) Replicator Diffusion Dynamics first calculates the replicator dynamics and then applies diffusion, so to speak. To understand what diffusion does, we'd like to see it run without the replicator dynamics. We look at the sender and the receiver roles seperately.

Diffusion
---------

Diffusion is defined in terms of

```{r}

get.nbr = function(response,noise.matrix){
    if (dim(noise.matrix)[1] == dim(response)[2]){  # receiver part
        return(response %*% noise.matrix)
    }
    if (dim(noise.matrix)[2] == dim(response)[1]){  # sender part
        return(noise.matrix %*% response)
    }
}

```

Preliminaries & Parameters
--------------------------

```{r}

alpha= 0.1
rounds = 1000
ns = 6
nm = 3
states = seq(from=0,to=1,length.out = ns)
C = get.confusion.matrix(states,alpha)
```

Initialization
--------------

Create an arbitrary sender and receiver strategy.

```{r}

sen = rdirichlet(ns,rep(1,nm))
rec = rdirichlet(nm,rep(1,ns))

show(rec)

for (i in 1:rounds){
#   sen = get.nbr(sen,C)
  rec = get.nbr(rec,C)
}

show(rec)

```