### 2014-09-16 snippets/data_exploration/data_exploration_convexity.Rmd

Data Exploration
================

``` {r warnings=FALSE, message=FALSE, echo=FALSE}

require('ellipse')
require('plyr')
require('ggplot2')
require('reshape2')
source('~/Desktop/data/svn/programming/R/helpers/helpers.R')

data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140916-134358.csv', 
                colClasses=c('numeric','factor','numeric','numeric','numeric',
                             'factor','numeric','numeric','numeric','numeric',
                             'numeric','numeric','numeric','numeric','numeric',
                             'numeric','numeric','numeric','character','character'))

ddply(data, .(Impairment,Number.of.states), summarise, no.convex = sum(Speaker.Convex.Cat)) -> sumconv

qplot(Impairment,no.convex,data=sumconv, geom=c("point","line"),ylab="percentage of convex outcomes", color=factor(Number.of.states))

```