### 2014-09-25 snippets/data_exploration/data_exploration_tmp.Rmd
restructured snippets

Data Exploration
================

``` {r warnings=FALSE, message=FALSE, echo=FALSE}

require('ellipse')
require('ggplot2')
require('reshape2')
require('plyr')
source('~/Desktop/data/svn/programming/R/helpers/helpers.R')
source('~/Desktop/data/svn/vagueness-games/data-analysis.R')
source('~/Desktop/data/svn/vagueness-games/helper_functions.R')

```

plot all the relevant variables against each other

``` {r warning = FALSE, fig.width=12, fig.height=12}

pairs(~ Number.of.states + Impairment + Tolerance + Speaker.entropy + Hearer.entropy + 
        Speaker.Voronoiness + Hearer.Voronoiness +
        Speaker.informativity + Hearer.informativity + Expected.utility + Iterations +
        Speaker.Convex.Cat, data=ds)
```

some properties are very clearly positively correlated:

* speaker informativity, hearer informativity & expected utility
* impairment & speaker entropy

these two "clusters" are also strongly negatively correlated

``` {r warning = FALSE, fig.width=7, fig.height=7}

dss = ds[c("Impairment", "Speaker.informativity", "Hearer.informativity", "Expected.utility",
           "Speaker.entropy")]

ctab <- cor(dss)

plotcorr(ctab)

```

Replotting data for each combination of values of independent variables
-----------------------------------------------------------------------

We have three independent variables: impairment, tolerance & number of states. The following plots show the means of all 50 data points for each combination of values for independent variables, together with the estimated confidence intervals.

```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=10}

dms = summarySE(dm, groupvars = c("Number.of.states", "Impairment", "Tolerance", "variable"), measurevar="value")
dsub = dms

pd <- position_dodge(.01)
sp = ggplot(dsub, aes(x=Impairment, y=value, colour=factor(Number.of.states))) + 
     geom_point(position = pd) +
     geom_errorbar(aes(ymin=value-ci, ymax=value+ci), width=0, position=pd) +
     geom_line(position=pd)

sp + facet_grid(variable ~ Tolerance, scales="free")

```

Informativity
-------------

This is the notion from Skyrms. It seems very well-behaved. There is a clear positive correlation between speaker and hearer informativity.

```{r fig.width=5, fig.height=5}
cor.test(data$Speaker.informativity, data$Hearer.informativity)
qplot(data$Speaker.informativity, data$Hearer.informativity) + geom_smooth(method=lm)
```

There is a clear negative correlation between impairment and informativity. The remaining variance here must and can be accounted for in terms of Number.of.states (see main plot).

```{r fig.width=5, fig.height=5}
cor.test(data$Impairment, data$Speaker.informativity)
qplot(data$Impairment, data$Speaker.informativity) + geom_smooth(method=lm)
```