### 2014-09-26 paper/03_simulations.tex
simulation results write up

\section{Exploring the RDD}

The previous section showed that the \rdd can be conceptualized as a
two-step procedure: we first calculate where the replicator dynamic
would take us, then apply diffusion. Since the latter step is the
novel addition, we should look at diffusion in isolation to understand
what the \rdd does. This is what we will do in
Section~\ref{sec:iterated-diffusion}. Section~\ref{sec:simulations}
 then explores how diffusion and fitness-based selection interact.

\subsection{(Iterated) Diffusion}
\label{sec:iterated-diffusion}

Confusion of states should intuitively be a function of their
perceptual similarity. Here and in the following, we assume that the
state space of the sim-max game consists of $n \ge 2$ states that are
equally spaced across the unit interval, including $0$ and $1$. We
assume here that the distance $\card{\mystate{i} - \mystate{j}}$ is
the objective, physical similarity between two states $\mystate{i}$
and $\mystate{j}$. Distance in physical space feeds into a perceptual
similarity function, as described by
\citet{Nosofsky1986:Attention-Simil}:
\begin{align}
  \label{eq:NosofskySim}
  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \impairment) =
      \begin{cases}
    1 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \impairment^2} \right ) \,,& \textrm{otherwise} \\
    \end{cases}
\end{align}
where $\impairment >= 0$ is an imprecision or indiscriminability
parameter. When $\impairment=0$ agents perfectly discriminate between
states; when $\impairment \rightarrow \infty$ agents cannot
discriminate states at all. Figure~\ref{fig:NosofskySim} gives an
impression of Nosofsky-similarity for different parameter
values. Other formalizations of perceptual similarity are possible,
including ones that allow for different discriminability in different
areas of the state space, but we stick with Nosofsky's similarity
function for the time being, because it is a mathematically simple,
yet established notion in mathematical psychology.

\begin{figure}
  \centering

  \includegraphics[width=\textwidth]{plots/NosofskySim.pdf}

  \caption{Examples for Nosofsky-similarity (non-normalized) and
    derived probability of state confusion (normalized) for
    $\card{\States} = 50$ and different
    values of $\impairment$.}
  \label{fig:NosofskySim}
\end{figure}

We further assume that the probability of confusing any two states
$\mystate{i}$ and $\mystate{j}$ is proportional to their perceived
similarity and therefore obtained by normalization:
\begin{align*}
  C_{ij} = \frac{\similarity(\mystate{i}, \mystate{j} \myts ; \myts
  \impairment)} {\sum_j \similarity(\mystate{i}, \mystate{j} \myts ; \myts
  \impairment)} \,.
\end{align*}
Confusion of states is then also a function of imprecision parameter
$\impairment$ (see Figure~\ref{fig:NosofskySim}, bottom part). For $\impairment
= 0$ the confusion matrix has $C_{ii} = 1$; for $\impairment > 0$ $C$ is
positive, i.e., $C_{ij} >0$ for all $i$ and $j$; for $\impairment
\rightarrow \infty$ we find $C_{ij} = \nicefrac{1}{\card{\States}}$ for all
$i$ and $j$.

Difusion of behavior under confusion of states was defined in
Equation~(\ref{eq:confusion-function}), repeated here:
\begin{align*}
  \Diff_C(\Sstrat) & = C \Sstrat &    \Diff_C(\Rstrat) & = \Rstrat C\,.
\end{align*}
If states are confused on average with a probability proportional to
their similarity, the effect on noise-perturbed behavior is that
behavioral dispositions gradually diffuse along a gradient of
similarity of states as well. Consequently, iterated diffusion leads
to a smoothing out and an eventual equalization of behavioral
strategies, for both the sender and the
receiver. Figure~\ref{fig:confusion-SenRec} shows an arbitrary sender
and receiver strategy after one or several rounds of iterated
application of diffusion, for different values of perceptual
imprecision $\impairment$.

The example in Figure~\ref{fig:confusion-SenRec} suggests that
iterated diffusion leads to sender and receiver behavior that is the
same at all choice points. This holds in general. After $n$ steps of
diffusion an initial sender strategy $\Sstrat$ will be $C^n \Sstrat$,
and an initial receiver strategy $\Rstrat$ will be $\Rstrat C^n$. If
we assume that $\impairment > 0$, $C$ is a positive row-stochastic
matrix (each state could in principle be confused as any other state,
albeit with possibly a very low probability). It then follows from the
Perron-Frobenius theorem that the limit $C^\infty = \lim_{n
  \rightarrow \infty} C^n$ exists (it is the Perron-projection) and
all of its rows are identical. That is why, in the limit of iterated
diffusion, $C^\infty S$ has identical rows (messages are sent with the
same probability in each state), and $R C^\infty$ likewise has
identical rows (every message is interpreted equally). All of this is
in line with the intuition that diffusion of behavior under confusion
of states iteratively equalizes behavior for similar states; if all
states can be confused for one another in principle, behavior smoothes
out entirely in the limit.

\begin{figure}
  \centering

  \includegraphics[width=\textwidth]{plots/confusion_sender.pdf}

  \includegraphics[width=\textwidth]{plots/confusion_receiver.pdf}

  \caption{Iterated diffusion for one arbitrary sender (top) and one
    receiver strategy (bottom), for a game with 50 states and 2
    messages. Each row corresponds to a different confusion level
    $\impairment$, each colum to a different number of applications of
    iterated diffusion.}
  \label{fig:confusion-SenRec}
\end{figure}

Now that we understand better what diffusion does, let us look at how
diffusion interacts with optimization of behavior, as described by the
replicator dynamics.

\subsection{Exploring the replicator diffusion dynamics}
\label{sec:simulations}

The main question we should address is whether diffusion and
fitness-based replication reasonably interact, and if so, whether the
inclusion of diffusion has any noteworthy effects on the evolving
meaning of signals. Clearly, if $\impairment = 0$, the \rdd reduces to
the \rd. If $\impairment \rightarrow \infty$, the diffusion component
takes over and the \rdd reduces to the trivial iterated diffusion that
we looked at in the previous section. For reasonable, in-between
levels of imprecision $\impairment$, the \rdd leads to communicatively
successful, yet vague signal meaning. Moreover, suitable levels of
imprecision can have further accelerating and, therefore, unifying
effects on meaning evolution.

To explore the \rdd, we looked at numerical simulations in sim-max
games with two signals whose states were evenly spaced elements of the
unit interval. The state space always included 0 and 1. We assumed
that priors were flat: $\Pr(\state) = \Pr(\state')$ for all $\state,
\state'$. As for utility functions, we implemented another
Nosofsky-style function:
\begin{align}
  \label{eq:NosofskyUtil}
  \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
      \begin{cases}
    1 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \toler^2} \right ) \,,& \textrm{otherwise} \\
    \end{cases}
\end{align}
The tolerance parameter $\toler$ models the amount of tolerable
pragmatic slack or communicative imprecision. For This choice of
utility function is governed partly by convenience in parallel to the
similarity function, but also because it has, as we believe, the right
general properties for a communicative payoff function. Unlike
utilities that are, say, linearly or quadratically decreasing
\citep[as used
by][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa},
utilities that are exponentially decreasing in negative quadratic
distance can model situations where a small amount of imprecision in
communication is tolerable, whereas similarly small differences in
intolerably off interpretations matter very little, with a smooth
transition between these regimes.

\begin{figure}
  \centering
  \subfloat[][$\ns = 90$, $\toler = 0.2$, $\impairment = 0.1$]{
    \includegraphics[width = 0.5\textwidth]{plots/strat_example_NS-90_tol-2_imp1_ind3501.pdf}
  }
  \subfloat[][$\ns = 90$, $\toler = 0.05$, $\impairment = 0$]{
    \includegraphics[width = 0.5\textwidth]{plots/strat_example_NS-90_tol-005_imp0_ind3002.pdf}
  }

  \subfloat[][$\ns = 10$, $\toler = 0.05$, $\impairment = 0$]{
    \includegraphics[width = 0.5\textwidth]{plots/strat_example_NS-10_tol-005_imp0_ind1001.pdf}
  }
  \subfloat[][$\ns = 6$, $\toler = 0.05$, $\impairment = 0.05$]{
    \includegraphics[width = 0.5\textwidth]{plots/strat_example_NS-10_tol-005_imp005_ind1201.pdf}
  }
  \caption{Example strategies under \rdd at stopping time of our simulations.}
  \label{fig:example_strats}
\end{figure}

We ran 50 trials of the \rdd, starting with randomly samples sender
and receiver strategies, for each triplet of independent parameter
values: $\ns = \card{T} \in \set{6, 10, 50, 90}$, $\impairment \in
\set{0, 0.05, 0.1, 0.2, 0.3}$, $\toler \in \set{0.05, 0.1, 0.2,
  0.3}$. Each trial ran for a maximum of 200 update steps of the
\rdd. This may not be enough to guarantee convergence to the eventual
attracting state, but we were interested especially in whether the
dynamics could lead to reasonable outcomes in reasonable time. A trial
was considered converged if the total amount of change between
strategies before and after the current \rdd step was smaller than
0.001. Examples for the resulting strategy pairs at stopping time are
given in Figure~\ref{fig:example_strats}.

\begin{itemize}
\item we see some interesting stuff in these strategies
\item now we look at the whole bunch
\item introduce some measures
\end{itemize}

\newpage

In order to study the long-term behavior of the noise-perturbed sim-max games, we conducted computer simulations for different combinations of initial parameters.
Namely, we were interested in investigating the impact of different values of impairment and its relation with the size of the state space.
\mytodo{JPC}{Justify why size of state space is interesting}

\subsubsection{Metrics}
\label{sec:metrics}
To be able to quantitively analyze the results of our simulations, we define a number of metrics to capture different characteristics of sim-max languages.

\paragraph{Entropy.}
This classic information-theoretic metric captures the amount of uncertainty in a signaling strategy.
The most natural way to define it is in terms of mixed strategies rather than behavioral strategies.
Let us recall that mixed sender (receiver) strategies are functions $\Smixed \in \Delta(\Messgs^\States)$ ($\Rmixed \in (\Acts^\Messgs)$).
We can define the entropy of a sender strategy as
\begin{align*}
  E(\Smixed) = \sum_{\Spure \in \Messgs^\States} \Smixed(\Spure) \cdot \log(\Smixed(\Spure))
\end{align*} 
and the entropy of a receiver strategy as
\begin{align*}
  E(\Rmixed) = \sum_{\Rpure \in \Acts^\Messgs} \Rmixed(\Rpure) \cdot \log(\Rmixed(\Rpure)) \,.
\end{align*} 
These metrics are computationally expensive to calculate, since the size of the domain over which the sum is computed grows exponentially with the number of choice points.
Therefore, we converted\footnote{See Appendix~\ref{sec:proofs}.} the above definitions into equivalent metrics defined in terms of behavioral strategies.
These are as follows:
\begin{align*}
  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs} \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar \state))
\end{align*} 
\begin{align*}
  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\act \in \Acts} \Rstrat(\act \probbar \messg) \cdot \log(\Rstrat(\act \probbar \messg)) \,.
\end{align*}
The metrics are lower bounded by $0$ and upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States| \cdot \log(|\Messgs|)$ and $\log(|\Acts^\Messgs|) = |\Messgs| \cdot \log(|\Acts|)$, thus we can normalize them by dividing by these values.

\paragraph{Informativity.}
\mytodo{JPC}{Informativity is perhaps a bad name for the receiver...}
The quantity of information in a signal is an old notion that also goes back to the start of information theory.
Skyrms~\citet[ch.~3]{Skyrms2010:Signals} discusses its use in the context of signaling games.
The main idea is that we can quantify the amount of information about a state $\state$ in a message $\messg$ via the relation between the probability of the state given the message $\Pr(\state \,|\, \messg)$ and the unconditional probability of the state $\Pr(\state)$.
Following Bayes' theorem, we can express $\Pr(\state \,|\, \messg)$ as $\frac{\Pr(\state) \cdot \Pr(\messg \,|\, \state)}{\Pr(\messg)}$.
We then have $\Pr(\messg \,|\, \state) = \Sstrat(\state,\messg)$ and $\Pr(\messg) = \sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)$.
Finally, we equate sender infomativity with the average overall information about states in each signal.
Based on the definition by Skyrms~\citet[p.~36]{Skyrms2010:Signals}, and the considerations above, we define sender informativity as follows:
\begin{align*}
  I(\Sstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs} \sum_{\state \in \States} \frac{\Pr(\state) \cdot \Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \cdot \log \left(\frac{\Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \right) \,.
\end{align*}
Conversely, we can quantify the amount of information about an act in a message.
We equate receiver informativity with the average overall information about acts in each signal.
Based on Skyrms' definition~\citet[p.~39]{Skyrms2010:Signals}, we define receiver informativity as follows:
\begin{align*}
  I(\Sstrat,\Rstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs} \sum_{\act \in \Acts} \Rstrat(\messg,\act) \cdot \log \left(\frac{\Rstrat(\messg,\act)}{\sum_{\state \in \States} \Pr(\state) \cdot \sum_{\messg^\prime \in \Messgs} \Sstrat(\state,\messg^\prime) \cdot \Rstrat(\messg^\prime,\act)} \right) \,.
\end{align*}
Both metrics are bounded between $0$ and $1$.

\paragraph{Voronoiness.}
This metric aims to quantify how close a strategy is to being a part of a Voronoi tessallation of the state space.
Recalling the results by \cite{Jager2007}, the stable outcomes of a population of agents playing a similarity-maximization game using pure strategies and evolving according to the replicator dynamics are those where ``the sender strategy is consistent with the Voronoi tessallation that is induced by the image of the receiver strategy.''~(\cite{Jager2007}, p. 562).
\mytodo{JPC}{Explain further?}
Given we are working with behavioral strategies, rather than pure strategies, and that we perform numerical simulations, rather than calculate analytical results, instead of looking into a binary criterion of whether a simulation result is such a Voronoi tessallation of the state space or not, we want a more progressive metric to characterize that in terms of degrees.
With that in mind, and given that this metric only makes sense for $\Acts = \States$, we define Voronoiness as follows.
Let $\pi(\Rstrat, \messg) = \argmax_{\state \in \States}\Rstrat(\messg,\state)$ be the prototype of a message $\messg \in \Messgs$, then
\begin{multline*}
  V(\Sstrat, \Rstrat) = \sum_{\state \in \States} \sum_{\messg_1 \in \Messgs} \sum_{\messg_2 \in \Messgs} \textrm{if } s(\state, \pi(\Rstrat, \messg_1)) > s(\state, \pi(\Rstrat, \messg_2)) \bicond \Sstrat(\state, \messg_1) > \Sstrat(\state, \messg_2) \textrm{ then } 1 \textrm{ else } 0
\end{multline*}
is the Voronoiness of the sender strategy $\Sstrat$ given the receiver strategy $\Rstrat$ and
\begin{align*}
  V(\Rstrat) = \sum_{\messg \in \Messgs} \sum_{\state_1 \in \States} \sum_{\state_2 \in \States} \textrm{if } s(\state_1, \pi(\Rstrat, \messg)) > s(\state_2, \pi(\Rstrat, \messg)) \bicond \Rstrat(\messg, \state_1) > \Rstrat(\messg, \state_2) \textrm{ then } 1 \textrm{ else } 0
\end{align*}
is the Voronoiness of the receiver strategy $\Rstrat$.
\mytodo{JPC}{Give some intuitions on these?}
The metrics are lower bounded by $0$ and upper bounded by, respectively, $|\States| \times |\Messgs|^2$ and $|\Messgs| \times |\States|^2$, thus we can normalize them by dividing by these values.

\subsubsection{Methodology}

In our simulations we focus on a concrete setup in line with \cite{Correia2013:The-Bivalent-Tr}.
Namely, we define a similarity-maximization game with $\States = \Acts$, where $\States$ is a discrete set of equidistant points within the unit interval $[0,1]$, including the boundaries (thus we always use $n \geq 2$).
As similarity metric $s:\States \times \States \rightarrow [0,1]$, we use a negative exponential of the of the square distance, \emph{i.e.}~for $\state,\state^\prime \in \States$:
\begin{align*}
  s(\state,\state^\prime) =
    \begin{cases}
    1 & \textrm{if } \impairment = 0 \textrm{ and } \state = \state^\prime \\
    0 & \textrm{if } \impairment = 0 \textrm{ and } \state \neq \state^\prime \\
    e^{-\frac{\lVert \state - \state^\prime \rVert ^2}{\impairment^2}} & \textrm{otherwise} \\
    \end{cases}
\end{align*}
where $\impairment$ is an \emph{impairment} factor.
Here we deviate from \cite{Correia2013:The-Bivalent-Tr} by working with impairment $\impairment$ rather than acuity $\theta$, since we are interested in exploring the case of having $0$ impairment, and acuity does not have a natural maximum as defined.
Confusion is directly related with similarity, \emph{i.e.}~the more similar two states are, the more likely an agent is to perceive one as the other.
For simplicity, we equate one with the other by defining $C_{ij} = s(\state_i,\state_j)$.

We run the noise-perturbed replicator dynamics with random initial strategies $\Sstrat_0$ and $\Rstrat_0$ until a convergence criterion is met, namely that the total absolute change in both sender and receiver strategies does not amount to more than $0.01$, \emph{i.e.}:
\begin{align*}
  \sum_{\state \in \States} \sum_{\messg \in \Messgs} \left| \Sstrat^\prime(\state,\messg) - \Sstrat(\state,\messg) \right| < 0.01
\end{align*}
and
\begin{align*}
  \sum_{\messg \in \Messgs} \sum_{\state \in \States} \left| \Rstrat^\prime(\messg,\state) - \Rstrat(\messg,\state) \right| < 0.01 \,.
\end{align*}

We perform a parameter sweep with size of the state space $n \in \{10,
50, 90, 130\}$ and impairment $\impairment \in \{0, 0.1, 0.2, 0.3,
0.4, 0.5\}$.
For each parameter combination, we perform $100$ trials and record the values of each of the metrics introduced in Section~\ref{sec:metrics} after convergence, as well as two additional indicators: number of iterations taken until convergence, and the normalized expected utility.
Expected utility is normalized by dividing by the expected utility of the theoretical optimum without impairment, \emph{i.e.}~a Voronoi tessalation of the state space.

\subsubsection{Results}
In order to visualize the impact of the impairment factor on the outcome of simulations, we plot in Figure~\ref{fig:metrics-boxplots} the distribution of each of the measurements per impairment value.
%
% \begin{figure}
%         \centering
%         \begin{subfigure}{0.45\textwidth}
%                 \includegraphics[width=\textwidth]{plots/Speaker-entropy-20140813-194306}
%                 \caption{Entropy of sender strategy.}
% %                \label{fig:gull}
%         \end{subfigure}
%         \begin{subfigure}{0.45\textwidth}
%                 \includegraphics[width=\textwidth]{plots/Hearer-entropy-20140813-194306}
%                 \caption{Entropy of receiver strategy.}
%         \end{subfigure}

%         \begin{subfigure}{0.45\textwidth}
%                 \includegraphics[width=\textwidth]{plots/Speaker-informativity-20140813-194306}
%                 \caption{Informativity of sender strategy.}
%         \end{subfigure}
%         \begin{subfigure}{0.45\textwidth}
%                 \includegraphics[width=\textwidth]{plots/Hearer-informativity-20140813-194306}
%                 \caption{Informativity of receiver strategy.}
%         \end{subfigure}

%         \begin{subfigure}{0.45\textwidth}
%                 \includegraphics[width=\textwidth]{plots/Speaker-Voronoiness-20140813-194306}
%                 \caption{Voronoiness of sender strategy.}
%         \end{subfigure}
%         \begin{subfigure}{0.45\textwidth}
%                 \includegraphics[width=\textwidth]{plots/Hearer-Voronoiness-20140813-194306}
%                 \caption{Voronoiness of receiver strategy.}
%         \end{subfigure}

%         \begin{subfigure}{0.45\textwidth}
%                 \includegraphics[width=\textwidth]{plots/Expected-utility-20140813-194306}
%                 \caption{Expected utility.}
%         \end{subfigure}
%         \begin{subfigure}{0.45\textwidth}
%                 \includegraphics[width=\textwidth]{plots/Iterations-20140813-194306}
%                 \caption{Number of iterations until convergence.}
%         \end{subfigure}
%         \caption{Boxplots of metrics per impairment value.}
%         \label{fig:metrics-boxplots}
% \end{figure}
%
Most trends are quite clear.
Entropy seems to increase with higher impairment values for both sender and receiver.
This is an expected outcome, since the higher the impairment value, the more likely each player is to confuse two states, thus the more uncertain he will be in unequivocally associating a state with a message, or vice-versa.
The trend for informativity is also expected: more vague strategies convey less information, be it from the sender or the receiver perspective.

Regarding Voronoiness, we observe that most simulation results approximate perfect Voronoi tessalations.
This is not the case, however, for impairment of $0$ (both sender and receiver) and $0.5$ (mostly sender only).
\mytodo{JPC}{For impairment 0, the languages with the lowest sender voronoiness look like non-convex equilibria. Discuss in relation with Elliott's results?}
For impairment of $0$,  the explanation is usually that strategies can achieve a high expected utility without having perfect Voronoiness.
Although the expected long term outcome in these situations is perfect Voronoiness (given J\"ager's analytical results~\cite{Jager2007}), actually reaching it involves minor refinements which at some point do not qualify as significant change given our convergence criterion.
The result is thus an artifact of stopping the simulations before they reach perfect convergence.
For impairment of $0.5$ (and some cases with impairment of $0.4$), the reason is simply that the languages are representatives of something between babbling and pooling equilibria.
These are equilibria where $\Sstrat(\state,\messg_1) \geq \Sstrat(\state,\messg_2)$ for all $\state \in \States$, but vary from situations where $\Sstrat(\state,\messg)$ is close to $0.5$ for both messages (babbling), or is significantly higher for one of them (pooling).
\mytodo{JPC}{Plot examples}

The trend for expected utility reflects Lipman's observation that vague languages are less optimal than non-vague ones~\cite{Lipman2009:Why-is-Language}: more impairment necessarily leads to lower expected utility.
Finally, the plot for number of iterations until convergence shows that a certain level of impairment enables faster convergence.
We will discuss this further in more detail in Section~\ref{sec:discussion}.

\mytodo{JPC}{Present and discuss results per size of state space? Don't forget O'Connor's claim that speed of convergence is more important for larger \#state/\#message ratio}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% TeX-PDF-mode: t
%%% End: