### 2014-09-26 paper/paper.tex

\documentclass[fleqn,reqno,10pt]{article}

%========================================
% Packages
%========================================

\usepackage[]{../helpers/mypackages}
%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
\bibliography{../helpers/MyRefGlobal}
\bibliography{paper}
\usepackage{../helpers/myenvironments}
\usepackage{../helpers/mycommands}
\usepackage{todonotes}
\usepackage{subfig}

%========================================
% Standard Layout
%========================================

% Itemize
\renewcommand{\labelitemi}{\large{$\mathbf{\cdot}$}}    % itemize symbols
\renewcommand{\labelitemii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiv}{\large{$\mathbf{\cdot}$}}
% Description
\renewcommand{\descriptionlabel}[1]{\hspace\labelsep\textsc{#1}}

% Figure Captions
\usepackage{caption} % use corresponding myfiguresize!
\setlength{\captionmargin}{20pt}
\renewcommand{\captionfont}{\small}
\setlength{\belowcaptionskip}{7pt} % standard is 0pt

%========================================
% Additional layout & commands
%========================================

\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}

% Annotations
\newcommand{\mytodo}[2]{\todo[inline,color=yellow,author=#1]{#2}}
\newcommand{\question}[2]{\todo[inline,color=blue,author=#1]{#2}}
\newcommand{\answer}[2]{\todo[inline,color=green,author=#1]{#2}}

\newcommand{\rd}{\acro{rd}} % replicator dynamic
\newcommand{\rmd}{\acro{rmd}} % replicator mutator dynamic
\newcommand{\rdd}{\acro{rdd}} % replicator diffusion dynamic
\newcommand{\RD}{\ensuremath{\mathrm{RD}}} % replicator dynamic
\newcommand{\RDD}{\ensuremath{\mathrm{RDD}}} % replicator diffusion dynamic
\newcommand{\RMD}{\ensuremath{\mathrm{RMD}}} % replicator mutator

\newcommand{\Diff}{\ensuremath{\mathrm{D}}} % Difusion 
\newcommand{\Mutate}{\ensuremath{\mathrm{M}}} % Mutation

\newcommand{\impairment}{\ensuremath{\alpha_i}} % impairment
\newcommand{\toler}{\ensuremath{\alpha_t}} % tolerance
\newcommand{\ns}{\ensuremath{n_s}} % number of states

\newcommand{\similarity}{\ensuremath{\mathrm{Sim}}} % similarity function

\title{Vagueness, Noise, and Communication}
\author{Jos\'e Pedro Correia and Michael Franke}
\date{}

\begin{document}
\maketitle

\begin{abstract}
  Signaling games have been garnering a lot of attention as models for
  studying natural language meaning.  One important aspect that is
  nevertheless typically missing from most models in the literature is
  the possibility for vagueness as a feature of successful languages.
  Recently, some models have been proposed to explicitly address this
  limitation.  In this paper we explore one such model and establish a
  link between the dynamics in the model and the replicator-mutator
  equations.  Furthermore, we study the long term behavior of the
  model using computer simulations.  The results indicate that vague
  languages, despite having lower expected utility and informativity,
  are faster to converge to an equilibrium than non-vague ones.  We
  argue that this could be an important advantage that contributes to
  the pervasiveness of vagueness in natural language.
\end{abstract}

\input{01_intro}

\input{02_noisyRMD}

\input{03_simulations}

\input{04_discussion}

\input{05_conclusions}

\input{06_appendix}

\printbibliography[heading=bibintoc]

\end{document}