### 2014-09-28 paper/03_simulations.tex
results report

\section{Exploring the RDD}

The previous section showed that the \rdd can be conceptualized as a
two-step procedure: we first calculate where the replicator dynamic
would take us, then apply diffusion. Since the latter step is the
novel addition, we should look at diffusion in isolation to understand
what the \rdd does. This is what we will do in
Section~\ref{sec:iterated-diffusion}. Section~\ref{sec:simulations}
 then explores how diffusion and fitness-based selection interact.

\subsection{(Iterated) Diffusion}
\label{sec:iterated-diffusion}

Confusion of states should intuitively be a function of their
perceptual similarity. Here and in the following, we assume that the
state space of the sim-max game consists of $n \ge 2$ states that are
equally spaced across the unit interval, including $0$ and $1$. We
assume here that the distance $\card{\mystate{i} - \mystate{j}}$ is
the objective, physical similarity between two states $\mystate{i}$
and $\mystate{j}$. Distance in physical space feeds into a perceptual
similarity function, as described by
\citet{Nosofsky1986:Attention-Simil}:
\begin{align}
  \label{eq:NosofskySim}
  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \impairment) =
      \begin{cases}
    1 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \impairment^2} \right ) \,,& \textrm{otherwise} \\
    \end{cases}
\end{align}
where $\impairment >= 0$ is an imprecision or indiscriminability
parameter. When $\impairment=0$ agents perfectly discriminate between
states; when $\impairment \rightarrow \infty$ agents cannot
discriminate states at all. Figure~\ref{fig:NosofskySim} gives an
impression of Nosofsky-similarity for different parameter
values. Other formalizations of perceptual similarity are possible,
including ones that allow for different discriminability in different
areas of the state space, but we stick with Nosofsky's similarity
function for the time being, because it is a mathematically simple,
yet established notion in mathematical psychology.

\begin{figure}
  \centering

  \includegraphics[width=\textwidth]{plots/NosofskySim.pdf}

  \caption{Examples for Nosofsky-similarity (non-normalized) and
    derived probability of state confusion (normalized) for
    $\card{\States} = 50$ and different
    values of $\impairment$.}
  \label{fig:NosofskySim}
\end{figure}

We further assume that the probability of confusing any two states
$\mystate{i}$ and $\mystate{j}$ is proportional to their perceived
similarity and therefore obtained by normalization:
\begin{align*}
  C_{ij} = \frac{\similarity(\mystate{i}, \mystate{j} \myts ; \myts
  \impairment)} {\sum_j \similarity(\mystate{i}, \mystate{j} \myts ; \myts
  \impairment)} \,.
\end{align*}
Confusion of states is then also a function of imprecision parameter
$\impairment$ (see Figure~\ref{fig:NosofskySim}, bottom part). For $\impairment
= 0$ the confusion matrix has $C_{ii} = 1$; for $\impairment > 0$ $C$ is
positive, i.e., $C_{ij} >0$ for all $i$ and $j$; for $\impairment
\rightarrow \infty$ we find $C_{ij} = \nicefrac{1}{\card{\States}}$ for all
$i$ and $j$.

Difusion of behavior under confusion of states was defined in
Equation~(\ref{eq:confusion-function}), repeated here:
\begin{align*}
  \Diff_C(\Sstrat) & = C \Sstrat &    \Diff_C(\Rstrat) & = \Rstrat C\,.
\end{align*}
If states are confused on average with a probability proportional to
their similarity, the effect on noise-perturbed behavior is that
behavioral dispositions gradually diffuse along a gradient of
similarity of states as well. Consequently, iterated diffusion leads
to a smoothing out and an eventual equalization of behavioral
strategies, for both the sender and the
receiver. Figure~\ref{fig:confusion-SenRec} shows an arbitrary sender
and receiver strategy after one or several rounds of iterated
application of diffusion, for different values of perceptual
imprecision $\impairment$.

The example in Figure~\ref{fig:confusion-SenRec} suggests that
iterated diffusion leads to sender and receiver behavior that is the
same at all choice points. This holds in general. After $n$ steps of
diffusion an initial sender strategy $\Sstrat$ will be $C^n \Sstrat$,
and an initial receiver strategy $\Rstrat$ will be $\Rstrat C^n$. If
we assume that $\impairment > 0$, $C$ is a positive row-stochastic
matrix (each state could in principle be confused as any other state,
albeit with possibly a very low probability). It then follows from the
Perron-Frobenius theorem that the limit $C^\infty = \lim_{n
  \rightarrow \infty} C^n$ exists (it is the Perron-projection) and
all of its rows are identical. That is why, in the limit of iterated
diffusion, $C^\infty S$ has identical rows (messages are sent with the
same probability in each state), and $R C^\infty$ likewise has
identical rows (every message is interpreted equally). All of this is
in line with the intuition that diffusion of behavior under confusion
of states iteratively equalizes behavior for similar states; if all
states can be confused for one another in principle, behavior smoothes
out entirely in the limit.

\begin{figure}
  \centering

  \includegraphics[width=\textwidth]{plots/confusion_sender.pdf}

  \includegraphics[width=\textwidth]{plots/confusion_receiver.pdf}

  \caption{Iterated diffusion for one arbitrary sender (top) and one
    receiver strategy (bottom), for a game with 50 states and 2
    messages. Each row corresponds to a different confusion level
    $\impairment$, each colum to a different number of applications of
    iterated diffusion.}
  \label{fig:confusion-SenRec}
\end{figure}

\subsection{Exploring the replicator diffusion dynamics}
\label{sec:simulations}

Now that we understand better what diffusion does, let us look at how
diffusion interacts with optimization of behavior, as described by the
replicator dynamics. The main question we should address is whether
diffusion and fitness-based replication reasonably interact, and if
so, whether the inclusion of diffusion has any noteworthy effects on
the evolving meaning of signals. Clearly, if $\impairment = 0$, the
\rdd reduces to the \rd. If $\impairment \rightarrow \infty$, the
diffusion component takes over and the \rdd reduces to the trivial
iterated diffusion that we looked at in the previous section. For
reasonable in-between levels of imprecision $\impairment$, the \rdd
leads to communicatively successful, yet vague signal
meaning. Moreover, suitable levels of imprecision can have further
accelerating and, therefore, unifying effects on meaning evolution.

To explore the \rdd, we looked at numerical simulations in sim-max
games with two signals whose states were evenly spaced elements of the
unit interval that always included 0 and 1. We assumed that priors
were flat: $\Pr(\state) = \Pr(\state')$ for all $\state, \state'$. As
for utility functions, we implemented another Nosofsky-style function:
\begin{align}
  \label{eq:NosofskyUtil}
  \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
      \begin{cases}
    1 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \toler^2} \right ) \,,& \textrm{otherwise} \\
    \end{cases}
\end{align}
The tolerance parameter $\toler$ models the amount of tolerable
pragmatic slack or communicative imprecision. This choice of utility
function is governed partly by convenience in parallel to the
similarity function, but also because it has, as we believe, the right
general properties for a communicative payoff function. Unlike
utilities that are, say, linearly or quadratically decreasing
\citep[as used
by][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa},
utilities that are exponentially decreasing in negative quadratic
distance can model situations where a small amount of imprecision in
communication is tolerable, whereas similarly small differences in
intolerably off interpretations matter very little, with a smooth
transition between these regimes \citep{OConnor2013:The-Evolution-o}.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_NS-90_tol-01_imp0_ind3098.pdf}
    \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0$}
    \label{fig:example_stratsA}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_NS-90_tol-01_imp01_ind3452.pdf}
    \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0.1$}
    \label{fig:example_stratsB}
  \end{subfigure}

  % \begin{subfigure}[]{0.45\textwidth}
  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp0_ind1001.pdf}
  %   \caption{$\ns = 10$, $\toler = 0.1$, $\impairment = 0$}
  %   \label{fig:example_stratsC}
  % \end{subfigure}
  % \hfill
  % \begin{subfigure}[]{0.45\textwidth}
  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp005_ind1201.pdf}
  %   \caption{$\ns = 10$, $\toler = 0.1$, $\impairment = 0.05$}
  %   \label{fig:example_stratsD}
  % \end{subfigure}

  \caption{Example strategies under \rdd at stopping time of our simulations.}
  \label{fig:example_strats}
\end{figure}

We ran 50 trials of the \rdd, starting with randomly sampled sender
and receiver strategies, for each triplet of independent parameter
values: $\ns = \card{T} \in \set{6, 10, 50, 90}$, $\impairment \in
\set{0, 0.05, 0.1, 0.2, 0.3}$, $\toler \in \set{0.05, 0.1, 0.2,
  0.3}$. Each trial ran for a maximum of 200 update steps of the
\rdd. This may not be enough to guarantee convergence to the eventual
attracting state, but we were interested especially in whether the
dynamics could lead to reasonable outcomes in reasonable time. A trial
was considered converged and stopped before the maximum of 200 rounds
if the total amount of change between strategies before and after the
current \rdd step was smaller than 0.001, i.e., if both:\todo{say what
the prime means}
\begin{align*}
  \sum_{\state \in \States} \sum_{\messg \in \Messgs} \left|
    \Sstrat^\prime(\state,\messg) - \Sstrat(\state,\messg) \right| <
  0.001 & \text{\ \ and} &
  \sum_{\messg \in \Messgs} \sum_{\state \in \States} \left| \Rstrat^\prime(\messg,\state) - \Rstrat(\messg,\state) \right| < 0.001 \,.
\end{align*}
Representative examples for the resulting strategy pairs at stopping
time are given in Figure~\ref{fig:example_strats}.

Figure~\ref{fig:example_stratsA} shows a strategy pair at stopping
time with 90 states, tolerance $\toler = 0.1$ and imprecision
$\impairment = 0$. The latter means that the trial was effectively a
plain application of the \rd. Noteworthily, the given sender strategy
approximate a pure sender strategy that crisply partitions the state
space into non-convex sets. The irregular shape of the receiver
strategy shows clearly that the pictured strategy pair has not yet
reached a dynamically stable state under the \rd. Indeed, the trial
did not converge in our technical sense, but was obtained by stopping
after 200 rounds of iterated \rdd. In contrast, the outcome of a trial
with identical parameters, except that imprecision was $\impairment =
0.1$, is shown in Figure~\ref{fig:example_stratsB} had converged after
98 rounds of iterated \rdd. The sender strategy shows a smooth
blending from one ``category'' to the other, and the receiver's
interpretations are rather extended curves, peaking at a central point
in the relevant ``categories''. What these examples show is that (i)
(speaker) strategies can approach non-convex strategies and that (ii)
inclusion of imprecision can lead to seemingly well-behaved, yet vague
strategies in the sense that we are after.

To explore further what happened in all of our 50 trials for each
parameter triple, we defined a small number of metrics that aim to
capture numerically how vague, communicatively efficient and generally
well-structured the resulting strategy pairs were.

\paragraph{Entropy.} This classic information-theoretic metric captures
the amount of uncertainty in a signaling strategy.  The most natural
way to define it is in terms of mixed strategies rather than
behavioral strategies.  Let us recall that mixed sender (receiver)
strategies are functions $\Smixed \in \Delta(\Messgs^\States)$
($\Rmixed \in (\Acts^\Messgs)$).  We can define the entropy of a
sender and receiver strategy as
\begin{align*}
  E(\Smixed) & = \sum_{\Spure \in \Messgs^\States} \Smixed(\Spure) \cdot
  \log(\Smixed(\Spure)) & 
  E(\Rmixed) & = \sum_{\Rpure \in \Acts^\Messgs} \Rmixed(\Rpure) \cdot \log(\Rmixed(\Rpure)) \,.
\end{align*} 
These metrics are computationally expensive to calculate, since the
size of the domain over which the sum is computed grows exponentially
with the number of choice points.  Therefore, we
converted\footnote{See Appendix~\ref{sec:proofs}.}\todo{include this!?} the above
definitions into equivalent metrics defined in terms of behavioral
strategies.  These are as follows:
\begin{align*}
  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs}
  \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar
  \state)) 
\end{align*} 
\begin{align*}
  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\act \in \Acts}
  \Rstrat(\act \probbar \messg) \cdot \log(\Rstrat(\act \probbar
  \messg)) \,. 
\end{align*}
The metrics are lower bounded by $0$ and upper bounded by,
respectively, $\log(|\Messgs^\States|) = |\States| \cdot
\log(|\Messgs|)$ and $\log(|\Acts^\Messgs|) = |\Messgs| \cdot
\log(|\Acts|)$, thus we can normalize them by dividing by these
values. The reported values are these normalized values.

\paragraph{Informativity.} The quantity of information in a signal is
an old notion that also goes back to the start of information theory.
Skyrms~\citet[ch.~3]{Skyrms2010:Signals} discusses its use in the
context of signaling games.  The main idea is that we can quantify the
amount of information about a state $\state$ in a message $\messg$ via
the relation between the probability of the state given the message
$\Pr(\state \,|\, \messg)$ and the unconditional probability of the
state $\Pr(\state)$.  Following Bayes' theorem, we can express
$\Pr(\state \,|\, \messg)$ as $\frac{\Pr(\state) \cdot \Pr(\messg
  \,|\, \state)}{\Pr(\messg)}$.  We then have $\Pr(\messg \,|\,
\state) = \Sstrat(\state,\messg)$ and $\Pr(\messg) =
\sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot
\Sstrat(\state^\prime,\messg)$.  Finally, we equate sender
infomativity with the average overall information about states in each
signal.  Based on the definition by
Skyrms~\citet[p.~36]{Skyrms2010:Signals}, and the considerations
above, we define sender informativity as follows:
\begin{align*}
  I(\Sstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs} \sum_{\state \in \States} \frac{\Pr(\state) \cdot \Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \cdot \log \left(\frac{\Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \right) \,.
\end{align*}
Conversely, we can quantify the amount of information about an act in a message.
We equate receiver informativity with the average overall information about acts in each signal.
Based on Skyrms' definition~\citet[p.~39]{Skyrms2010:Signals}, we define receiver informativity as follows:
\begin{align*}
  I(\Sstrat,\Rstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs} \sum_{\act \in \Acts} \Rstrat(\messg,\act) \cdot \log \left(\frac{\Rstrat(\messg,\act)}{\sum_{\state \in \States} \Pr(\state) \cdot \sum_{\messg^\prime \in \Messgs} \Sstrat(\state,\messg^\prime) \cdot \Rstrat(\messg^\prime,\act)} \right) \,.
\end{align*}
Both metrics are bounded between $0$ and $1$.

\paragraph{Voronoiness.}
This metric aims to quantify how close a strategy is to being a part of a Voronoi tessallation of the state space.
Recalling the results by \cite{Jager2007}, the stable outcomes of a population of agents playing a similarity-maximization game using pure strategies and evolving according to the replicator dynamics are those where ``the sender strategy is consistent with the Voronoi tessallation that is induced by the image of the receiver strategy.''~(\cite{Jager2007}, p. 562).
\mytodo{JPC}{Explain further?}
Given we are working with behavioral strategies, rather than pure strategies, and that we perform numerical simulations, rather than calculate analytical results, instead of looking into a binary criterion of whether a simulation result is such a Voronoi tessallation of the state space or not, we want a more progressive metric to characterize that in terms of degrees.
With that in mind, and given that this metric only makes sense for $\Acts = \States$, we define Voronoiness as follows.
Let $\pi(\Rstrat, \messg) = \argmax_{\state \in \States}\Rstrat(\messg,\state)$ be the prototype of a message $\messg \in \Messgs$, then
\begin{multline*}
  V(\Sstrat, \Rstrat) = \sum_{\state \in \States} \sum_{\messg_1 \in \Messgs} \sum_{\messg_2 \in \Messgs} \textrm{if } s(\state, \pi(\Rstrat, \messg_1)) > s(\state, \pi(\Rstrat, \messg_2)) \bicond \Sstrat(\state, \messg_1) > \Sstrat(\state, \messg_2) \textrm{ then } 1 \textrm{ else } 0
\end{multline*}
is the Voronoiness of the sender strategy $\Sstrat$ given the receiver strategy $\Rstrat$ and
\begin{align*}
  V(\Rstrat) = \sum_{\messg \in \Messgs} \sum_{\state_1 \in \States} \sum_{\state_2 \in \States} \textrm{if } s(\state_1, \pi(\Rstrat, \messg)) > s(\state_2, \pi(\Rstrat, \messg)) \bicond \Rstrat(\messg, \state_1) > \Rstrat(\messg, \state_2) \textrm{ then } 1 \textrm{ else } 0
\end{align*}
is the Voronoiness of the receiver strategy $\Rstrat$.
\mytodo{JPC}{Give some intuitions on these?}
The metrics are lower bounded by $0$ and upper bounded by, respectively, $|\States| \times |\Messgs|^2$ and $|\Messgs| \times |\States|^2$, thus we can normalize them by dividing by these values.

\subsubsection{Methodology}

In our simulations we focus on a concrete setup in line with \cite{Correia2013:The-Bivalent-Tr}.
Namely, we define a similarity-maximization game with $\States = \Acts$, where $\States$ is a discrete set of equidistant points within the unit interval $[0,1]$, including the boundaries (thus we always use $n \geq 2$).
As similarity metric $s:\States \times \States \rightarrow [0,1]$, we use a negative exponential of the of the square distance, \emph{i.e.}~for $\state,\state^\prime \in \States$:
\begin{align*}
  s(\state,\state^\prime) =
    \begin{cases}
    1 & \textrm{if } \impairment = 0 \textrm{ and } \state = \state^\prime \\
    0 & \textrm{if } \impairment = 0 \textrm{ and } \state \neq \state^\prime \\
    e^{-\frac{\lVert \state - \state^\prime \rVert ^2}{\impairment^2}} & \textrm{otherwise} \\
    \end{cases}
\end{align*}
where $\impairment$ is an \emph{impairment} factor.
Here we deviate from \cite{Correia2013:The-Bivalent-Tr} by working with impairment $\impairment$ rather than acuity $\theta$, since we are interested in exploring the case of having $0$ impairment, and acuity does not have a natural maximum as defined.
Confusion is directly related with similarity, \emph{i.e.}~the more similar two states are, the more likely an agent is to perceive one as the other.
For simplicity, we equate one with the other by defining $C_{ij} = s(\state_i,\state_j)$.

We run the noise-perturbed replicator dynamics with random initial strategies $\Sstrat_0$ and $\Rstrat_0$ until a convergence criterion is met, namely that the total absolute change in both sender and receiver strategies does not amount to more than $0.01$, \emph{i.e.}:

We perform a parameter sweep with size of the state space $n \in \{10,
50, 90, 130\}$ and impairment $\impairment \in \{0, 0.1, 0.2, 0.3,
0.4, 0.5\}$.
For each parameter combination, we perform $100$ trials and record the values of each of the metrics introduced in Section~\ref{sec:metrics} after convergence, as well as two additional indicators: number of iterations taken until convergence, and the normalized expected utility.
Expected utility is normalized by dividing by the expected utility of the theoretical optimum without impairment, \emph{i.e.}~a Voronoi tessalation of the state space.

\subsubsection{Results}
In order to visualize the impact of the impairment factor on the outcome of simulations, we plot in Figure~\ref{fig:metrics-boxplots} the distribution of each of the measurements per impairment value.
%
% \begin{figure}
%         \centering
%         \begin{subfigure}{0.45\textwidth}
%                 \includegraphics[width=\textwidth]{plots/Speaker-entropy-20140813-194306}
%                 \caption{Entropy of sender strategy.}
% %                \label{fig:gull}
%         \end{subfigure}
%         \begin{subfigure}{0.45\textwidth}
%                 \includegraphics[width=\textwidth]{plots/Hearer-entropy-20140813-194306}
%                 \caption{Entropy of receiver strategy.}
%         \end{subfigure}

%         \begin{subfigure}{0.45\textwidth}
%                 \includegraphics[width=\textwidth]{plots/Speaker-informativity-20140813-194306}
%                 \caption{Informativity of sender strategy.}
%         \end{subfigure}
%         \begin{subfigure}{0.45\textwidth}
%                 \includegraphics[width=\textwidth]{plots/Hearer-informativity-20140813-194306}
%                 \caption{Informativity of receiver strategy.}
%         \end{subfigure}

%         \begin{subfigure}{0.45\textwidth}
%                 \includegraphics[width=\textwidth]{plots/Speaker-Voronoiness-20140813-194306}
%                 \caption{Voronoiness of sender strategy.}
%         \end{subfigure}
%         \begin{subfigure}{0.45\textwidth}
%                 \includegraphics[width=\textwidth]{plots/Hearer-Voronoiness-20140813-194306}
%                 \caption{Voronoiness of receiver strategy.}
%         \end{subfigure}

%         \begin{subfigure}{0.45\textwidth}
%                 \includegraphics[width=\textwidth]{plots/Expected-utility-20140813-194306}
%                 \caption{Expected utility.}
%         \end{subfigure}
%         \begin{subfigure}{0.45\textwidth}
%                 \includegraphics[width=\textwidth]{plots/Iterations-20140813-194306}
%                 \caption{Number of iterations until convergence.}
%         \end{subfigure}
%         \caption{Boxplots of metrics per impairment value.}
%         \label{fig:metrics-boxplots}
% \end{figure}
%
Most trends are quite clear.
Entropy seems to increase with higher impairment values for both sender and receiver.
This is an expected outcome, since the higher the impairment value, the more likely each player is to confuse two states, thus the more uncertain he will be in unequivocally associating a state with a message, or vice-versa.
The trend for informativity is also expected: more vague strategies convey less information, be it from the sender or the receiver perspective.

Regarding Voronoiness, we observe that most simulation results approximate perfect Voronoi tessalations.
This is not the case, however, for impairment of $0$ (both sender and receiver) and $0.5$ (mostly sender only).
\mytodo{JPC}{For impairment 0, the languages with the lowest sender voronoiness look like non-convex equilibria. Discuss in relation with Elliott's results?}
For impairment of $0$,  the explanation is usually that strategies can achieve a high expected utility without having perfect Voronoiness.
Although the expected long term outcome in these situations is perfect Voronoiness (given J\"ager's analytical results~\cite{Jager2007}), actually reaching it involves minor refinements which at some point do not qualify as significant change given our convergence criterion.
The result is thus an artifact of stopping the simulations before they reach perfect convergence.
For impairment of $0.5$ (and some cases with impairment of $0.4$), the reason is simply that the languages are representatives of something between babbling and pooling equilibria.
These are equilibria where $\Sstrat(\state,\messg_1) \geq \Sstrat(\state,\messg_2)$ for all $\state \in \States$, but vary from situations where $\Sstrat(\state,\messg)$ is close to $0.5$ for both messages (babbling), or is significantly higher for one of them (pooling).
\mytodo{JPC}{Plot examples}

The trend for expected utility reflects Lipman's observation that vague languages are less optimal than non-vague ones~\cite{Lipman2009:Why-is-Language}: more impairment necessarily leads to lower expected utility.
Finally, the plot for number of iterations until convergence shows that a certain level of impairment enables faster convergence.
We will discuss this further in more detail in Section~\ref{sec:discussion}.

\mytodo{JPC}{Present and discuss results per size of state space? Don't forget O'Connor's claim that speed of convergence is more important for larger \#state/\#message ratio}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% TeX-PDF-mode: t
%%% End: