### 2014-09-29 paper/03_simulations.tex
results section sec com

\section{Exploring the RDD}

The previous section showed that the \rdd can be conceptualized as a
two-step procedure: we first calculate where the replicator dynamic
would take us, then apply diffusion. Since the latter step is the
novel addition, we should look at diffusion in isolation to understand
what the \rdd does. This is what we will do in
Section~\ref{sec:iterated-diffusion}. Section~\ref{sec:simulations}
 then explores how diffusion and fitness-based selection interact.

\subsection{(Iterated) Diffusion}
\label{sec:iterated-diffusion}

Confusion of states should intuitively be a function of their
perceptual similarity. Here and in the following, we assume that the
state space of the sim-max game consists of $n \ge 2$ states that are
equally spaced across the unit interval, including $0$ and $1$. We
assume here that the distance $\card{\mystate{i} - \mystate{j}}$ is
the objective, physical similarity between two states $\mystate{i}$
and $\mystate{j}$. Distance in physical space feeds into a perceptual
similarity function, as described by
\citet{Nosofsky1986:Attention-Simil}:
\begin{align}
  \label{eq:NosofskySim}
  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \impairment) =
      \begin{cases}
    1 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \impairment^2} \right ) \,,& \textrm{otherwise} \\
    \end{cases}
\end{align}
where $\impairment >= 0$ is an imprecision or indiscriminability
parameter. When $\impairment=0$ agents perfectly discriminate between
states; when $\impairment \rightarrow \infty$ agents cannot
discriminate states at all. Figure~\ref{fig:NosofskySim} gives an
impression of Nosofsky-similarity for different parameter
values. Other formalizations of perceptual similarity are possible,
including ones that allow for different discriminability in different
areas of the state space, but we stick with Nosofsky's similarity
function for the time being, because it is a mathematically simple,
yet established notion in mathematical psychology.

\begin{figure}
  \centering

  \includegraphics[width=\textwidth]{plots/NosofskySim.pdf}

  \caption{Examples for Nosofsky-similarity (non-normalized) and
    derived probability of state confusion (normalized) for
    $\card{\States} = 50$ and different
    values of $\impairment$.}
  \label{fig:NosofskySim}
\end{figure}

We further assume that the probability of confusing any two states
$\mystate{i}$ and $\mystate{j}$ is proportional to their perceived
similarity and therefore obtained by normalization:
\begin{align*}
  C_{ij} = \frac{\similarity(\mystate{i}, \mystate{j} \myts ; \myts
  \impairment)} {\sum_j \similarity(\mystate{i}, \mystate{j} \myts ; \myts
  \impairment)} \,.
\end{align*}
Confusion of states is then also a function of imprecision parameter
$\impairment$ (see Figure~\ref{fig:NosofskySim}, bottom part). For $\impairment
= 0$ the confusion matrix has $C_{ii} = 1$; for $\impairment > 0$ $C$ is
positive, i.e., $C_{ij} >0$ for all $i$ and $j$; for $\impairment
\rightarrow \infty$ we find $C_{ij} = \nicefrac{1}{\card{\States}}$ for all
$i$ and $j$.

Difusion of behavior under confusion of states was defined in
Equation~(\ref{eq:confusion-function}), repeated here:
\begin{align*}
  \Diff_C(\Sstrat) & = C \Sstrat &    \Diff_C(\Rstrat) & = \Rstrat C\,.
\end{align*}
If states are confused on average with a probability proportional to
their similarity, the effect on noise-perturbed behavior is that
behavioral dispositions gradually diffuse along a gradient of
similarity of states as well. Consequently, iterated diffusion leads
to a smoothing out and an eventual equalization of behavioral
strategies, for both the sender and the
receiver. Figure~\ref{fig:confusion-SenRec} shows an arbitrary sender
and receiver strategy after one or several rounds of iterated
application of diffusion, for different values of perceptual
imprecision $\impairment$.

The example in Figure~\ref{fig:confusion-SenRec} suggests that
iterated diffusion leads to sender and receiver behavior that is the
same at all choice points. This holds in general. After $n$ steps of
diffusion an initial sender strategy $\Sstrat$ will be $C^n \Sstrat$,
and an initial receiver strategy $\Rstrat$ will be $\Rstrat C^n$. If
we assume that $\impairment > 0$, $C$ is a positive row-stochastic
matrix (each state could in principle be confused as any other state,
albeit with possibly a very low probability). It then follows from the
Perron-Frobenius theorem that the limit $C^\infty = \lim_{n
  \rightarrow \infty} C^n$ exists (it is the Perron-projection) and
all of its rows are identical. That is why, in the limit of iterated
diffusion, $C^\infty S$ has identical rows (messages are sent with the
same probability in each state), and $R C^\infty$ likewise has
identical rows (every message is interpreted equally). All of this is
in line with the intuition that diffusion of behavior under confusion
of states iteratively equalizes behavior for similar states; if all
states can be confused for one another in principle, behavior smoothes
out entirely in the limit.

\begin{figure}
  \centering

  \includegraphics[width=\textwidth]{plots/confusion_sender.pdf}

  \includegraphics[width=\textwidth]{plots/confusion_receiver.pdf}

  \caption{Iterated diffusion for one arbitrary sender (top) and one
    receiver strategy (bottom), for a game with 50 states and 2
    messages. Each row corresponds to a different confusion level
    $\impairment$, each colum to a different number of applications of
    iterated diffusion.}
  \label{fig:confusion-SenRec}
\end{figure}

\subsection{Exploring the replicator diffusion dynamics}
\label{sec:simulations}

Now that we understand better what diffusion does, let us look at how
diffusion interacts with optimization of behavior, as described by the
replicator dynamics. The main question we should address is whether
diffusion and fitness-based replication reasonably interact, and if
so, whether the inclusion of diffusion has any noteworthy effects on
the evolving meaning of signals. Clearly, if $\impairment = 0$, the
\rdd reduces to the \rd. If $\impairment \rightarrow \infty$, the
diffusion component takes over and the \rdd reduces to the trivial
iterated diffusion that we looked at in the previous section. We will
show presently that for reasonable in-between levels of imprecision
$\impairment$, the \rdd leads to communicatively successful, yet vague
signal meaning. Moreover, suitable levels of imprecision can have
further accelerating and, therefore, unifying effects on meaning
evolution.

To explore the \rdd, we looked at numerical simulations in sim-max
games with two signals whose states were evenly spaced elements of the
unit interval that always included 0 and 1. We assumed that priors
were flat: $\Pr(\state) = \Pr(\state')$ for all $\state, \state'$. As
for utility functions, we implemented another Nosofsky-style function:
\begin{align}
  \label{eq:NosofskyUtil}
  \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
      \begin{cases}
    1 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \toler^2} \right ) \,,& \textrm{otherwise} \\
    \end{cases}
\end{align}
The tolerance parameter $\toler > 0$ models the amount of tolerable
pragmatic slack or communicative imprecision. This choice of utility
function is governed partly by convenience in parallel to the
similarity function, but also because it has, as we believe, the right
general properties for a communicative payoff function. Unlike
utilities that are, say, linearly or quadratically decreasing in
physical distance \citep[as used
by][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa},
utilities that are exponentially decreasing in negative quadratic
distance can model situations where a small amount of imprecision in
communication is tolerable, whereas similarly small differences in
intolerably far away interpretations matter very little, with a smooth
transition between these regimes
\citep[c.f.][]{OConnor2013:The-Evolution-o}.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_NS-90_tol-01_imp0_ind3098.pdf}
    \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0$}
    \label{fig:example_stratsA}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_NS-90_tol-01_imp01_ind3452.pdf}
    \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0.1$}
    \label{fig:example_stratsB}
  \end{subfigure}

  % \begin{subfigure}[]{0.45\textwidth}
  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp0_ind1001.pdf}
  %   \caption{$\ns = 10$, $\toler = 0.1$, $\impairment = 0$}
  %   \label{fig:example_stratsC}
  % \end{subfigure}
  % \hfill
  % \begin{subfigure}[]{0.45\textwidth}
  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp005_ind1201.pdf}
  %   \caption{$\ns = 10$, $\toler = 0.1$, $\impairment = 0.05$}
  %   \label{fig:example_stratsD}
  % \end{subfigure}

  \caption{Example strategies under \rdd at stopping time of our simulations.}
  \label{fig:example_strats}
\end{figure}

We ran 50 trials of the \rdd, starting with randomly sampled sender
and receiver strategies, for each triplet of independent parameter
values: $\ns = \card{T} \in \set{6, 10, 50, 90}$, $\impairment \in
\set{0, 0.05, 0.1, 0.2, 0.3}$, $\toler \in \set{0.05, 0.1, 0.2,
  0.3}$. Each trial ran for a maximum of 200 update steps of the
\rdd. This may not be enough to guarantee convergence to the eventual
attracting state, but we were interested especially in whether the
dynamics could lead to reasonable outcomes in reasonable time. A trial
was considered converged and stopped before the maximum of 200 rounds
if the total amount of change between strategies before and after the
current \rdd step was smaller than 0.001, i.e., if:
\begin{align*}
  \sum_{\state \in \States} \sum_{\messg \in \Messgs} \left|
    \rdd(\Sstrat)(\state,\messg) - \Sstrat(\state,\messg) \right| <
  0.001\,, & \text{\ \ and} \\
  \sum_{\messg \in \Messgs} \sum_{\state \in \States} \left|
    \rdd(\Rstrat)(\messg,\state) - \Rstrat(\messg,\state) \right| <
  0.001 & \,.
\end{align*}

Representative examples for resulting strategy pairs at stopping time
are given in
Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA}
shows a strategy pair at stopping time with 90 states, tolerance
$\toler = 0.1$ and imprecision $\impairment = 0$. The latter means
that the trial was effectively a plain application of the
\rd. Noteworthily, the given sender strategy approximates a pure
sender strategy that crisply partitions the state space into
non-convex sets. The irregular shape of the receiver strategy shows
clearly that the pictured strategy pair has not yet reached a
dynamically stable state under the \rd. Indeed, the trial did not
converge in our technical sense, but was stopped after 200 rounds of
iterated \rdd. In contrast, the outcome of a trial with identical
parameters, except that imprecision was $\impairment = 0.1$, is shown
in Figure~\ref{fig:example_stratsB} had converged (in our technical
sense) after 98 rounds of iterated \rdd. The sender strategy shows a
smooth blending from one ``category'' to the other, and the receiver's
interpretations are rather extended curves, peaking at a central point
in the relevant ``categories''. What these examples show is that (i)
(speaker) strategies can approach non-convex strategies and that (ii)
inclusion of imprecision can lead to seemingly well-behaved, yet vague
strategies in the sense that we are after.

To explore further what happened in all of our 50 trials for each
parameter triple, we defined a small number of metrics that aim to
capture numerically how vague, communicatively efficient and generally
well-structured the resulting strategy pairs were. Entropy and
informativity capture (slightly different aspects of) the amount of
systematicity or regularity in a language. Voronoiness and convexity
are, respectively, gradient and categorical measures for the coherence
of linguistically formed categories.

\paragraph{Entropy.} This classic information-theoretic metric
captures the amount of uncertainty in a signaling strategy. Roughly
put, entropy of a strategy captures inverse distance from a pure
strategy in the present context. Vague languages will have higher
entropy than crisp ones.

The most natural way to define it is in terms of mixed strategies
rather than behavioral strategies.  Let us recall that mixed sender
(receiver) strategies are functions $\Smixed \in
\Delta(\Messgs^\States)$ ($\Rmixed \in (\Acts^\Messgs)$).  We can
define the entropy of a sender and receiver strategy as
\begin{align*}
  E(\Smixed) & = \sum_{\Spure \in \Messgs^\States} \Smixed(\Spure) \cdot
  \log(\Smixed(\Spure)) & 
  E(\Rmixed) & = \sum_{\Rpure \in \Acts^\Messgs} \Rmixed(\Rpure) \cdot \log(\Rmixed(\Rpure)) \,.
\end{align*} 
These measures are computationally expensive to calculate, since the
size of the domain over which the sum is computed grows exponentially
with the number of choice points.  Therefore, we
converted\footnote{See Appendix~\ref{sec:proofs}.}\todo{include
  this!?} the above definitions into equivalent metrics defined in
terms of behavioral strategies.  These are as follows:
\begin{align*}
  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs}
  \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar
  \state)) 
\end{align*} 
\begin{align*}
  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\act \in \Acts}
  \Rstrat(\act \probbar \messg) \cdot \log(\Rstrat(\act \probbar
  \messg)) \,. 
\end{align*}
Values obtained by these definitions are lower bounded by $0$ and
upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States|
\cdot \log(|\Messgs|)$ and $\log(|\Acts^\Messgs|) = |\Messgs| \cdot
\log(|\Acts|)$. We work with normalized values.

The speaker strategies in Figures~\ref{fig:example_stratsA} and
\ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.21$,
respectively. The receiver strategies have respective entropies $0.43$
and $0.84$.

\paragraph{Informativity.} The quantity of information in a signal is
an old notion that also goes back to the start of information theory.
Skyrms~\citet[ch.~3]{Skyrms2010:Signals} discusses its use in the
context of signaling games. Intuitively speaking, informativity in
Skyrms' sense measures how revealing the message use is about which
state is actual. It therefore measures, in our given set-up with flat
priors, a sense of well-behavedness or regularity of a sender
strategy. Vague languages will have lower informativity than crisp
ones.

The main idea behind Skyrm's notion is that we can quantify the amount
of information about a state $\state$ in a message $\messg$ via the
relation between the probability of the state given the message
$\Pr(\state \,|\, \messg)$ and the unconditional probability of the
state $\Pr(\state)$.  Following Bayes' theorem, we can express
$\Pr(\state \,|\, \messg)$ as $\frac{\Pr(\state) \cdot \Pr(\messg
  \,|\, \state)}{\Pr(\messg)}$.  We then have $\Pr(\messg \,|\,
\state) = \Sstrat(\state,\messg)$ and $\Pr(\messg) =
\sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot
\Sstrat(\state^\prime,\messg)$.  Finally, we equate sender
infomativity with the average overall information about states in each
signal.  Following \citet[p.~36]{Skyrms2010:Signals}, and the
considerations above, the informativity about states conveyed by a
sender strategy is:
\begin{align*}
  I(\Sstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs}
  \sum_{\state \in \States} \frac{\Pr(\state) \cdot
    \Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States}
    \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \cdot \log
  \left(\frac{\Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States}
      \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \right)
  \,. 
\end{align*}
This measure is bounded between $0$ and $1$. The speaker strategies in
Figures~\ref{fig:example_stratsA} and \ref{fig:example_stratsB} have
informativity $0.69$ and $0.54$, respectively.

\citet[p.~39]{Skyrms2010:Signals} also offers a notion of
informativity about the acts performed by the receiver, but it turns
out that this is perfectly correlated with information about the
states in the results of our simulations. We therefore only look at
informativity about states.

% Conversely, we can quantify the amount of information about an act in a message.
% We equate receiver informativity with the average overall information about acts in each signal.
% Based on Skyrms' definition~\citet[p.~39]{Skyrms2010:Signals}, we define receiver informativity as follows:
% \begin{align*}
%   I(\Sstrat,\Rstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs} \sum_{\act \in \Acts} \Rstrat(\messg,\act) \cdot \log \left(\frac{\Rstrat(\messg,\act)}{\sum_{\state \in \States} \Pr(\state) \cdot \sum_{\messg^\prime \in \Messgs} \Sstrat(\state,\messg^\prime) \cdot \Rstrat(\messg^\prime,\act)} \right) \,.
% \end{align*}

\paragraph{Voronoiness.}
This metric aims to quantify how close a strategy is to being a part
of a Voronoi tessallation of the state space. Bearing in mind that
this metric is specifically geared towards the sim-max games in our
set-up, we define Voronoiness as follows. Let $\pi(\Rstrat, \messg) =
\argmax_{\state \in \States}\Rstrat(\messg,\state)$ be the prototype
of a message $\messg \in \Messgs$, then
\begin{multline*}
  V(\Sstrat, \Rstrat) = \sum_{\state \in \States} \sum_{\messg_1 \in
    \Messgs} \sum_{\messg_2 \in \Messgs} \textrm{if } s(\state,
  \pi(\Rstrat, \messg_1)) > s(\state, \pi(\Rstrat, \messg_2)) \bicond
  \Sstrat(\state, \messg_1) > \Sstrat(\state, \messg_2) \textrm{ then
  } 1 \textrm{ else } 0 
\end{multline*}
is the Voronoiness of the sender strategy $\Sstrat$ given the receiver
strategy $\Rstrat$ and
\begin{align*}
  V(\Rstrat) = \sum_{\messg \in \Messgs} \sum_{\state_1 \in \States}
  \sum_{\state_2 \in \States} \textrm{if } s(\state_1, \pi(\Rstrat,
  \messg)) > s(\state_2, \pi(\Rstrat, \messg)) \bicond \Rstrat(\messg,
  \state_1) > \Rstrat(\messg, \state_2) \textrm{ then } 1 \textrm{
    else } 0
\end{align*}
is the Voronoiness of the receiver strategy $\Rstrat$.  The metrics
are lower bounded by $0$ and upper bounded by, respectively,
$|\States| \times |\Messgs|^2$ and $|\Messgs| \times |\States|^2$,
thus we can normalize them by dividing by these values.

Despite being vague, a language can have maximal Voronoiness according
to this gradient measure. For instance, the sender strategies in
Figures~\ref{fig:example_stratsA} and \ref{fig:example_stratsB} have
Voronoiness $0.91$ and $1$, respectively. The receiver strategies have
Voronoiness $0.78$ and $1$, respectively.

\paragraph{Convexity.} At least for speaker strategies, which converge
faster, it also makes sense to define a categorical measure of
convexity, that compensates for potential vagueness. To determine
whether a sender strategy $\Sstrat$ was convex despite possibly being
vague, we looked at the derived pure strategy $\Spure$ for which
$\Spure(\state) = \messg$ only if $\Sstrat(\state,\messg) =
\max_{\messg' \in \Messgs} \Sstrat(\state,\messg')$. If that $\Spure$
was convex, we also counted $\Sstrat$ as convex.

\paragraph{Expected utility.} We also recorded the expected utility of
a strategy pair:
\begin{align*}
  \EU(\Sstrat,\Rstrat \myts ; \myts \toler) = \sum_{\state \in
    \States} \sum_{\messg \in \Messgs} \sum_{\state' \in \States}
  \Pr(\state) \cdot \Sstrat(\state, \messg) \cdot \Rstrat(\messg,
  \state') \cdot \Util(\state,\state' \myts ; \myts \toler)\,.
\end{align*}
To make direct comparisons across different parameter settings, we
normalized expected utility by the maximal amount of expected utility
obtainable in the relevant game.

\subsubsection{Results}
Figure~\ref{fig:MeanMetrics} shows the means of the recorded metrics,
together with their estimated 95\% confidence intervals, for each
triple of independent variables. Most trends are quite clear. Entropy
of sender and receiver strategies are increasing in imprecision
$\impairment$, as can be expected. Entropy of speaker strategies
hardly seems independent of tolerance levels and the number of states,
while there seems to be an effect of both parameters on the entropy of
the hearer strategy.

\begin{figure}
  \centering

  \includegraphics[width=\textwidth]{plots/MeanMetrics.pdf}

  \caption{Means of recorded metrics for each triple of independent variables.}
  \label{fig:MeanMetrics}
\end{figure}

Speaker informativity and expected utility are decreasing under
growing imprecision, again, as should be expected. There seem to be
mild effects of the number of states in the game, but hardly any in
terms of pragmatic tolerance.

Interestingly, voronoiness of both sender and receiver strategies are
appear to be non-decreasing under increasing impairment and peaking at
the maximal value. The only exception is the 6 states games and the
transition from $\impairment = 0$ to $\impairment = 0.05$. This
suggests that impairment can help evolve more regular, well-behaved
languages.

This impression is also supported by the proportion of converged and
convex trials, pictured in
Figure~\ref{fig:CategoricalMeasures}. Non-converged and non-convex
speaker strategies showed for several values of parameters in the
absence of impairment. Increasing impairment always assured
convergence and convexity.

\begin{figure}
  \centering

  \includegraphics[width=\textwidth]{plots/CategoricalMeasures.pdf}

  \caption{Proportion of converged and convex trials for each triple
    of independent variables.}
  \label{fig:CategoricalMeasures}
\end{figure}

Summarizing so far, it seems that diffusion increases entropy and
vagueness, decreases informativity and expected utility, but it also
speeds up convergence to a convex and well-behaved outcome.

Diffusion of strategies from confusion of states has another
interesting regularizing effect on the evolution of signaling. If we
look at the confidence intervals in Figure~\ref{fig:MeanMetrics}, we
see that there is very little variation in the recorded metrics for
evolved strategies, at least for higher values of impairment. On
closer inspection, it turns out that variability in low-impairment
conditions is not only due to non-convergence and/or associated
non-convexity. Figure~\ref{fig:MoreExample_strats} gives two more
examples of strategy pairs at stopping time. Both are obtained for the
same triple of parameters, both converged and are convex. However,
they are not equally efficient. In fact, the pair in
Figure~\ref{fig:example_stratsC} has a normalized expected utility of
$0.99$ while the one in Figure~\ref{fig:example_stratsD} only has
$0.89$.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_NS-6_tol-03_imp005_ind41.pdf}
    \caption{$\ns = 6$, $\toler = 0.2$, $\impairment = 0$}
    \label{fig:example_stratsC}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_NS-6_tol-03_imp005_ind23.pdf}
    \caption{$\ns = 6$, $\toler = 0.2$, $\impairment = 0$}
    \label{fig:example_stratsD}
  \end{subfigure}

  \caption{More example strategies under \rdd at stopping time of our simulations.}
  \label{fig:MoreExample_strats}
\end{figure}

This type of variability in evolutionary outcomes is also weeded out
by mild increases in impairment.

\begin{figure}
  \centering

    \includegraphics[width=\textwidth]{plots/AverageEUinGroups.pdf}

   \caption{Average normalized expected utility between evolved strategies within each triplet of independent parameter values.}
  \label{fig:AverageEUinGroups}
\end{figure}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% TeX-PDF-mode: t
%%% End: