### 2014-10-02 paper/01_intro.tex
intro; sec com

\section{Introduction}

Many of our concepts and words are vague. A vague category knows clear
cases that fall under it, clear cases that do not, and also so-called
borderline cases in-between that neither clearly apply, nor clearly
not apply, but might themselves not be equal in terms of how well they
represent the category in question. Vagueness does not seem to
dramatically affect the success of everyday communication, but it is
troublesome for some of our best theories of language. Especially the
logico-semantic tradition of Frege, Russell and the young Wittgenstein
is challenged by vagueness and the paradoxes that it gives rise to. As
can be expected, many proponents of this tradition have risen to the
challenge \citep{Williamson1994:Vagueness}.

There are other intriguing aspects about vagueness. Linguists, for
example, wonder about the reasons why some expressions are less vague
than others \citep[e.g.][]{Kennedy2007:Vagueness-and-G}. But the
puzzle that we are concerned with here is how vagueness could arise
and be maintained in the first place. This may not seem a particularly
deep issue at first glance, but on closer inspection it is a serious
worry to functionalist accounts that maintain that our concepts and
language use evolved in order to be efficient. Since the existence of
unclear borderline cases seems to entail inefficiency of
categorization or communication, the challenge, succinctly put forward
by \citet{Lipman2009:Why-is-Language}, is to explain how vagueness can
exist despite its obvious and demonstrable functional deficiency under
evolutionary pressure to be optimally expressive.

This problem, call it Lipman's problem, has a conceptual and a
technical side to it. Conceptually, it asks for reasons and general
mechanisms by which we could plausibly conceive of vagueness as
persisting the pressure of evolutionary optimization towards
precision. Such reasons and general mechanisms are, arguably, not hard
to imagine. We will address some presently. But each putative
explanation, no matter how plausible intuitively, must also be checked
for internal consistency and its potential to account for the fact
that we see vagueness as a part of a by-and-large efficient system of
categorization and communication. In other words, not all first-shot
rebuttals of Lipman's problem will do. Denying that there is any
functional pressure on efficient communication, for instance, is
dubious because it leaves the relative efficiency of our communication
with vague words entirely unexplained. In sum, the technical answer to
Lipman's problem is to give, ideally, a model of evolution of
categorization or language use that shows how competing forces that
realize both by-and-large efficient categories and vagueness.

c

\bigskip

\begin{itemize}
\item introduce sim-max games
\item problem of vagueness \citep{Lipman2009:Why-is-Language}
\item \dots
\end{itemize}

\subsection{Situating our approach}

A number of authors have recently tried to explain why vagueness
evolved, based on considerations why a vague language might be useful
\citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,Jaegherde-JaegherRooijvan-Rooij2010:Strategic-Vague,BlumeBoard2013:Intentional-Vag}. In
contrast to these, others have argued that vagueness if a natural
byproduct of limitations in information processing
\citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o}. This
paper makes a contribution to the latter line of thought.

The view that vagueness is a natural concomitant of cognitive
limitations of language users has been formalized in a number of
ways. As in other applications of evolutionary game theory, we can
distinguish micro- and macro-level approaches. Micro-level approaches
look at adaptive behavior of individual agents. Usually, changes in
the behavioral dispositions of agents occur after every single
interaction. In contrast, macro-level approaches outline more
abstract, aggregate population-based dynamics. Usually, a macro-level
dynamic captures changes of frequencies of behavioral types in the
population over time. Previous accounts have suggested both micro- and
macro-level dynamics to show how vague signal use can arise in sim-max
games.

As for micro-dynamics, \citet{FrankeJager2010:Vagueness-Signa} show
how limited memory of past interactions can lead to vague signal use,
when averaging over a single agent's behavior over time or over the
momentary behavior of a population of several language
users. \citet{OConnor2013:The-Evolution-o} introduces a variant of
reinforcement learning that includes a low-level form of stimulus
generalization. Agents update their behavior after each round of play
in such a way that similar states to the ones that actually occurred
are subject to behavioral adjustment as well. (Alternative approaches
are discussed in more detail in Section~XYZ, where we can compare them
with the system introduced in this paper.)

\citet{FrankeJager2010:Vagueness-Signa} also consider a macro-level
approach, using the notion of a \emph{quantal response}. A quantal
response function is a probabilistic choice rule that formalizes the
idea that agents make small mistakes when calculating the expected
utility of choice options. In aggregation, these probabilistic
mistakes lead to systematic ``trembles'' that produce vague signal
use. The approach we take here is superficially similar, but there are
crucial divergences. For one, we adopt a dynamic perspective by
looking at the stable states of a suitable population dynamics. For
another, we demonstrate in Section~XYZ that quantal responses can give
rise to counterintuitive predictions. These counterintuitive examples
suggest that vagueness in signaling use is not convincingly explained
by appeal to mistakes in calculating expected utility, but rather, as
we assume here, as the result of confusing similar states of the word.

Against this background, we introduce a new macro-level approach
to the evolution of vague signal use in sim-max games, where the
source of vagueness is the agents' natural disability to sharply
distinguish perceptually distinct states. This applies to both
production and comprehension, i.e., both to the sender as well as to
the receiver role.

We call our dynamic \textbf{replicator diffusion dynamic} (\rdd),
because it is a special case of the replicator mutator dynamic
(\rmd). It is a special case in the sense that the \rdd incorporates a
special kind of mutation that is based on the similarity of states (as
defined in the sim-max game). More concretely, we start from the idea
that similar states are more easily confused than less similar
states. To formalize this idea, we assume that there is a confusion
matrix $C$, where $C_{ij}$ is the probability that a state \mystate{i}
is realized as a possibly different state \mystate{j}. (There is a
slightly different sense of ``realization'' in the sender and the
receiver side, and we will enlarge on this below.) Based on this, we
show two things. For one, we show how the discrete-time formulation of
the replicator dynamic can be easily adapted to integrate diffusion of
behavior due to confusion of states. This yields the \rdd in its basic
discrete-time formulation. For another, we also show that there exists
a translation of the confusion matrix $C$ into a suitable mutation
matrix $M_C$ such that the effect of $C$ and $M_C$ are ``equivalent''
in a sense to be made precise. This supports the conclusion that the
\rdd can indeed be thought of as a special case of the \rmd.

To understand the workings of the \rdd, we study numerical simulations
\dots

The paper is structured as follows:
\begin{enumerate}
\item replicator diffusion dynamics
\item relation to replicator mutator dynamics
\item simulations of \rdd
\item comparison with other approaches
\item conclusions
\end{enumerate}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% TeX-PDF-mode: t
%%% End: