### 2014-10-02 paper/paper.tex

\documentclass[fleqn,reqno,10pt]{article}

%========================================
% Packages
%========================================

\usepackage[]{../helpers/mypackages}
%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
\bibliography{MyRefGlobal}
% \bibliography{paper}
\usepackage{../helpers/myenvironments}
\usepackage{../helpers/mycommands}
\usepackage{todonotes}
\usepackage{subcaption}

%========================================
% Standard Layout
%========================================

% Itemize
\renewcommand{\labelitemi}{\large{$\mathbf{\cdot}$}}    % itemize symbols
\renewcommand{\labelitemii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiv}{\large{$\mathbf{\cdot}$}}
% Description
\renewcommand{\descriptionlabel}[1]{\hspace\labelsep\textsc{#1}}

% Figure Captions
\usepackage{caption} % use corresponding myfiguresize!
\setlength{\captionmargin}{20pt}
\renewcommand{\captionfont}{\small}
\setlength{\belowcaptionskip}{7pt} % standard is 0pt

%========================================
% Additional layout & commands
%========================================

\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}

% Annotations
\newcommand{\mytodo}[2]{\todo[inline,color=yellow,author=#1]{#2}}
\newcommand{\question}[2]{\todo[inline,color=blue,author=#1]{#2}}
\newcommand{\answer}[2]{\todo[inline,color=green,author=#1]{#2}}

\newcommand{\rd}{\acro{rd}} % replicator dynamic
\newcommand{\rmd}{\acro{rmd}} % replicator mutator dynamic
\newcommand{\rdd}{\acro{rdd}} % replicator diffusion dynamic
\newcommand{\RD}{\ensuremath{\mathrm{RD}}} % replicator dynamic
\newcommand{\RDD}{\ensuremath{\mathrm{RDD}}} % replicator diffusion dynamic
\newcommand{\RMD}{\ensuremath{\mathrm{RMD}}} % replicator mutator

\newcommand{\Diff}{\ensuremath{\mathrm{D}}} % Difusion 
\newcommand{\Mutate}{\ensuremath{\mathrm{M}}} % Mutation

\newcommand{\impairment}{\ensuremath{\alpha_i}} % impairment
\newcommand{\toler}{\ensuremath{\alpha_t}} % tolerance
\newcommand{\ns}{\ensuremath{n_s}} % number of states

\newcommand{\similarity}{\ensuremath{\mathrm{Sim}}} % similarity function

\title{Vagueness, Noise, and Communication}
% \author{Jos\'e Pedro Correia and Michael Franke}
\date{}

\begin{document}
\maketitle

\begin{abstract}
  Signaling games have attracted a lot of attention as models for
  studying the evolution of meaning. One important aspect that is
  nevertheless typically missing from most models in the literature is
  the possibility for vagueness as a feature of successful signaling.
  Complementing recent models that have been proposed to explicitly
  address this limitation, we introduce the replicator diffusion
  dynamic, provably a special case of the replicator mutator dynamic,
  which implements a versatile and natural integration of stochastic
  noise in the form of probabilistic confusion of similar stimuli, on
  the one hand, and evolutionary pressure on optimal signal use, on
  the other. Applying the replicator diffusion dynamic to a
  generalization of David Lewis' signaling games, so-called
  similarity-maximizing games, we show that stochastic noise can not
  only lead to vague, yet communicative efficient signal use, but can
  also unify evolutionary outcomes and help avoid inefficient
  categorization.
\end{abstract}

\input{01_intro}

\input{02_noisyRMD}

\input{03_simulations}

\input{04_discussion}

\input{05_conclusions}

\input{06_appendix}

\printbibliography[heading=bibintoc]

\end{document}