### 2014-10-02 paper/01_intro.tex
introduction and background; sec com

\section{Introduction}

Many of our concepts and words are vague. A vague category knows clear
cases that fall under it, clear cases that do not, and also so-called
borderline cases in-between that neither clearly apply, nor clearly
not apply, but might themselves not be equal in terms of how well they
represent the category in question. Vagueness does not seem to
dramatically affect the success of everyday communication, but it is
troublesome for some of our best theories of language. Especially the
logico-semantic tradition of Frege, Russell and the young Wittgenstein
is challenged by vagueness and the paradoxes that it gives rise to. As
can be expected, many proponents of this tradition have risen to the
challenge
\citep[e.g.][]{Williamson1994:Vagueness,CobrerosEgre2012:Tolerant-Classi}.\todo{more
  references}

Yet there are other intriguing aspects about vagueness. The
puzzle that we are concerned with here is how vagueness could arise
and be maintained in the first place. This may not seem a particularly
deep issue at first glance, but on closer inspection it is a serious
worry to functionalist accounts that maintain that our concepts and
language use evolved in order to be efficient. Since the existence of
unclear borderline cases seems to entail inefficiency of
categorization or communication, the challenge, succinctly put forward
by \citet{Lipman2009:Why-is-Language}, is to explain how vagueness can
exist despite its obvious and demonstrable functional deficiency under
evolutionary pressure to be optimally expressive.

This problem, call it Lipman's problem, has a conceptual and a
technical side to it. Conceptually, it asks for reasons and general
mechanisms by which we could plausibly conceive of vagueness as
resisting the pressure of evolutionary selection for precision. Such
reasons and general mechanisms are, arguably, not hard to imagine. We
will address some presently. But each putative explanation, no matter
how plausible intuitively, must also be checked for internal
consistency and its potential to account for the fact that we see
vagueness as a part of a by-and-large efficient system of
categorization and communication. In other words, not all first-shot
rebuttals of Lipman's problem will do. Denying that there is any
functional pressure on efficient communication, for instance, is a
non-starter, because it leaves the relative efficiency of our
communication with vague words entirely unexplained. In sum, the
technical answer to Lipman's problem is to give, ideally, a model of
evolution of categorization or language use that shows how competing
forces that realize both by-and-large efficient categories and
vagueness.

A number of authors have recently tried to explain why vagueness
evolved, based on considerations why a vague language might be useful
\citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,Jaegherde-JaegherRooijvan-Rooij2010:Strategic-Vague,BlumeBoard2013:Intentional-Vag}. In
contrast to these, others have argued that vagueness is a natural
byproduct of limitations in information processing
\citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o,QingFranke2014:Gradable-Adject}. This
paper makes a contribution to the latter line of thought.  Concretely,
we introduce a novel variant of the replicator mutator dynamic that
integrates stochastic noise on the differential confusability of
similar stimuli. Our main contribution, therefore, is a technical
answer to Lipman's problem in the sense introduced above. The dynamic
proposed here nicely improves on and complements previous like-minded
accounts. We show that that stochastic noise can not only lead to
vague, yet communicative efficient signal use, but can also unify
evolutionary outcomes and help avoid inefficient categorization.

The next section introduces the background against which the work
presented here can be appreciated. Section~\ref{sec:repl-diff-dynam},
then, introduces the replicator diffusion dynamic, and elaborates its
relation with the replicator mutator
dynamic. Section~\ref{sec:exploring-rdd} explores the replicator
diffusion dynamic on the relevant class of generalized signaling games
introduced in
Section~\ref{sec:background}. Section~\ref{sec:discussion} reflects on
the results and compares what has been achieved to related accounts in
more detail.

\section{Background}
\label{sec:background}

The view that vagueness is a natural concomitant of cognitive
limitations of language users has been formalized in a number of ways,
using evolutionary game theory and certain generalizations of
signaling games, so called similarity-maximizing games or sim-max
games, for short
\citep{FrankeJager2010:Vagueness-Signa,OConnor2013:Evolving-Percep,OConnor2013:The-Evolution-o}. Our
contribution is best seen in relation to these accounts, as it also
relies on sim-max games. Let's introduce these first, and then zoom in
on the problem of vagueness.

\subsection{Sim-max games \& conceptual spaces}

Signaling games, as introduced by \citet{Lewis_1969:Convention}, have
a sender and a receiver. The sender knows the true state of the world,
but the receiver does not. The sender can select a signal, or message,
to reveal to the receiver, who then chooses an act. In Lewis' games,
states are maximally distinct from each other and the receiver's acts
are related to them one-to-one. If the receiver chooses the act that
corresponds to the actual state, the play is a success, otherwise a
failure. Certain regular combinations of sender signaling and receiver
reaction make messages meaningful, in the sense that their use is
correlated systematically to certain states or acts. To investigate
the conditions under which such meaning-generating behavior can evolve
is a highly interesting topic that we are only beginning to fully
understand
\citep[e.g.][]{Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,Zollman2005:Talking-to-Neig,Huttegger2007:Evolution-and-t,Pawlowitsch2008:Why-Evolution-d,Barrett2009:The-Evolution-o,Wagner2009:Communication-a,HutteggerSkyrms2010:Evolutionary-Dy,Skyrms2010:Signals,HutteggerZollman2011:Signaling-Games}.

Similarity-maximizing games are generalizations of Lewis' games in
that they allow different states to be more or less similar to one
another, and, roughly put, make success of communication a function of
that similarity. Intuitively speaking, while Lewis' games treated
communicative success as a matter of black and white, sim-max games
allow for a gradient notion of communicative success: the more similar
the receiver's interpretation is to the actual state, the more
successful the play was. Signaling games with utility-relevant
similarities in the state space are fairly standard in economics
\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor},
but have received particular attention in a more philosophical and
linguistic context by the more recent work for reasons that will
become clear presently
\citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag}.

More concretely, a sim-max game, in the relevant sense here, consists
of a set of states $\States$, a set of messages $\Messgs$ with much
fever messages than states, a probability distribution $\Pr \in
\Delta(\States)$ that gives the occurrence probabilities of states, a
similarity metric on states $\similarity \mycolon \States \times
\States \rightarrow \mathds{R}$, and a utility function $\Utils
\mycolon \States \times \States \rightarrow \mathds{R}$. We identify
the receiver's acts with the states of the world, so that the game is
one of guessing the actual state, so to speak. We also assume that
sender's and receiver's interests are alike, so we only have one
utility function. We do not consider message costs, so utilities only
depend on the actual state, and the receiver's response. The
similarity function should satisfy the usual requirements for a
metric:
\begin{align*}
  & \similarity(\mystate{1}, \mystate{2}) \ge 0 &&
  \similarity(\mystate{1}, \mystate{2}) = 0 \Leftrightarrow
  \mystate{1} = \mystate{2} \\
  & \similarity(\mystate{1}, \mystate{2}) = \similarity(\mystate{2},
  \mystate{1}) && \similarity(\mystate{1}, \mystate{2}) \le
  \similarity(\mystate{1}, \mystate{3}) + \similarity(\mystate{3}, \mystate{2})\,.
\end{align*}
The utility function should be a monotonically decreasing function of
similarity:
\begin{align*}
  \similarity(\mystate{1},\mystate{2}) \ge
  \similarity(\mystate{1},\mystate{3}) \ \ \Rightarrow \ \ 
  \Utils(\mystate{1},\mystate{2}) \ge
  \Utils(\mystate{1},\mystate{3})\,.
\end{align*}
To keep matters simple, we only look at cases where $\States$ contains
finitely many points from the unit interval, all of which have equal
probability, and with similarity given by the Euclidean distance in
this one-dimensional space.\todo{introduce utilities here?}

\citet{JagerMetzger2011:Voronoi-Languag} showed that the
evolutionarily stable states of sim-max games are remarkably
systematic. Their results were obtained for games with infinitely many
states in $n$-dimensional Euclidean space $\States \subseteq
\mathds{R}^n$ and a quadratic loss function for utilities
$\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} -
\mystate{2})^2$. For these games, the evolutionarily stable states are
demonstrably so-called Voronoi-languages. Intuitively speaking, a
Voronoi language is a pair of sender and receiver strategies, such
that the sender strategy (quasi-)partitions the state space into
convex categories, while the receiver's interpretations are the
central spots in each category. Figure~\ref{fig:VoronoiLanguage} gives
an example for such a Voronoi-language for a sim-max game with
$\States = [0,1]$, two messages, and a flat prior distribution that
selects all states equally likely. The sender's strategy (a function
from states to messages; more below) uses one message exclusively for
all points in the lower interval and another for all points in the
upper interval. The receiver's interpretations of messages are the
central points in the respective intervals.

\begin{figure}
  \centering

  \begin{tikzpicture}[scale=3]

        \begin{scope}

          \draw[-,thick] (0,0) -- (1.5,0);

          \draw (0,0) node {\color{red}{\big{[}}};

          \draw (.75,0) node {\color{red}{\big{)}}};

          \draw (.76,0) node {\color{red}{\big{(}}};

          \draw (1.5,0) node {\color{red}{\big{]}}};

            \draw (.325,.15) node
            {\color{red}{$\Spure(\mymessg{1}  \probbar \cdot)$}};

            \draw (1.075,.15) node
            {\color{red}{$\Spure(\mymessg{2} \probbar \cdot)$}};

            \node at (.375,0) (node1) {\color{blue}{$\bullet$}};

            \node at (.375,-.3) (label1)
            {\color{blue}{${\Rpure}(\cdot \probbar \mymessg{1})$}};

            \draw[-,color=blue] (node1) -- (label1);

            \node at (1.125,0) (node2) {\color{blue}{$\bullet$}};

            \node at (1.125,-.3) (label2)
            {\color{blue}{${\Rpure}(\cdot \probbar \mymessg{2})$}};

            \draw[-] (node2) -- (label2);

            \draw (0,-.2) node {$0$};

            \draw (.75,-.2) node {$0.5$};

            \draw (1.5,-.2) node {$1$};

          \end{scope}

        \end{tikzpicture}

  \caption{Example of a Voronoi-language on $\States = [0,1]$. The
    sender strategy $\Spure$ uses one signal for the lower half, and
    another for the upper half of the unit interval. The receiver
    strategy $\Rpure$ interprets messages as the central elements in the
    respective intervals.}
  \label{fig:VoronoiLanguage}
\end{figure}

This result is particularly interesting, because it demonstrates that
signaling can impose orderly categories on a metric space, without
that being the ulterior purpose of it all. Sender and receiver can be
distinct entities, whose purpose it is to communicate effectively
about the actual state. In that case, evolving Voronoi-languages would
explain why linguistic categories are well-behaved and orderly in the
way they appear to be. More abstractly, sender and receiver can also
be thought of as distinct modules in a single system, where the first
module that must discretize the information that it is fed into a
small sample of, suggestively, category labels to pass to a second
system that tries to decode the original information. In this case,
evolving Voronoi-languages would explain why conceptual categories are
well-behaved and orderly in the way that they appear to be
\citep[see][for more on this latter
interpretation]{OConnor2013:Evolving-Percep}.

Seen in this light, sim-max games appear to have the potential of
providing a foundation to approaches in cognitive semantics that rely
on the notion of conceptual spaces.
\citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has
prominently argued that natural categories are convex regions in a
suitable conceptual space. If the conceptual space has a suitable
metric, then it is possible to think of these convex categories as
derived from a set of prototypes. Fixing a set of prototypes, we
consider the category corresponding to each prototype $p$ as the set
of points that are more similar to $p$ than to any other
\citep[e.g.][]{OkabeBoots2000:Spatial-Tessell}. In this way,
\citet{Gardenfors2000:Conceptual-Spac} argues, an efficient
categorization system can be obtained: storing the prototypes lets us
recover the categories without having to store each categories
extension. However, what is left unexplained so far, is where the
prototypes come from, and why we would not see just any distribution
of prototypes as an equally efficient classification system. This is
where sim-max games, might chip in a principled approach to deriving
in an independent way, not only convex categories, but also
prototypical exemplars belonging to them.

\subsection{Vague signaling in sim-max games}

The brief outline of an approach to conceptual categorization using
sim-max games leaves many problems unaddressed. One of them is that
natural categories for continuously variable stimuli, like shades of
color, pitch heights, spatial dimension and the like, do not have
unique prototypes and clear category boundaries. We would like to
account for the possibility of vagueness, in particular the following
criteria
\citep[e.g.][]{Sainsbury1991:Is-There-Higher,KeefeSmith1997:Vagueness:-A-Re,Smith2008:Vagueness-and-D}:
(i) clear positive examples of a vague category should show a
gradient, perhaps smooth, transition to negative examples to
accommodate also ``higher-order vagueness'' ; (ii) prototypes should
likewise be gradient regions, peaking at the center of the vague
category they represent.

\citet{DouvenDecock2011:Vagueness:-A-Co} show that
\citeauthor{Gardenfors2000:Conceptual-Spac}'s conceptual spaces
approach can be extended to account for the existence of borderline
cases. From the assumption that prototypes are not unitary points, but
extended, yet convex regions in conceptual space,
\citeauthor{DouvenDecock2011:Vagueness:-A-Co} give a construction
algorithm that yields ``collated Voronoi diagrams'' with thick,
extended boundaries representing borderline regions. Building on this
work, \citet{DecockDouven2012:What-is-Graded-} show further how it is
possible to arrive at ``higher-order vagueness'' and degrees of
category membership, by, essentially, weighing in the distance of
different borderline examples to various prototypical regions. This
accounts for the first of the two desiderata mentioned above, but
still assumes that crisp non-gradient prototype regions must be
pre-given.

An alternative approach is taken by
\citet{FrankeJager2010:Vagueness-Signa} and
\citet{OConnor2013:The-Evolution-o} who show how the above desiderata
can be met by evolving strategies in sim-max games for certain types
of solution concepts, that can be divided into micro- and macro-level
approaches. Micro-level approaches look at adaptive behavior of
individual agents. Usually, changes in the behavioral dispositions of
agents occur after every single interaction. In contrast, macro-level
approaches outline more abstract, aggregate dynamics, happening in a
huge population of agents, or otherwise abstracting from seemingly
irrelevant detail. Usually, a macro-level dynamic captures changes of
frequencies of behavioral types in the population over time.

As for micro-dynamics, \citet{FrankeJager2010:Vagueness-Signa} show
how limited memory of past interactions can lead to vague signal use,
when averaging over a single agent's behavior over time or over the
momentary behavior of a population of several language
users. \citet{OConnor2013:The-Evolution-o} introduces a variant of
reinforcement learning that includes a low-level form of stimulus
generalization. Agents update their behavior after each round of play
in such a way that similar states to the ones that actually occurred
are subject to behavioral adjustment as well. (Alternative approaches
are discussed in more detail in Section~\ref{sec:discussion}, where we
can compare them with the system introduced in this paper.)

\citet{FrankeJager2010:Vagueness-Signa} also consider a macro-level
approach, using the notion of a \emph{quantal response}. A quantal
response function is a probabilistic choice rule that formalizes the
idea that agents make small mistakes when calculating the expected
utility of choice options. In aggregation, these probabilistic
mistakes lead to systematic ``trembles'' that produce vague signal
use. The approach we take here is superficially similar, but there are
crucial differences. For one, we adopt a dynamic perspective by
looking at limited-time outcomes of a dynamic process. For another, we
demonstrate in Section~\ref{sec:discussion} that quantal responses can
give rise to counterintuitive predictions. These counterintuitive
examples suggest that vagueness in signaling use is not convincingly
explained by appeal to mistakes in calculating expected utility, but
rather, as we assume here, as the result of confusing similar states
of the word.

Against this background, we introduce a new macro-level approach to
the evolution of vague signal use in sim-max games, where the source
of vagueness is the agents' natural inability to sharply distinguish
perceptually distinct states. This applies to the sender as well as to
the receiver role (albeit in slightly different manner). We call our
dynamic \emph{replicator diffusion dynamic} (\rdd), because it is a
special case of the replicator mutator dynamic (\rmd). It is a special
case in the sense that the \rdd incorporates a particular kind of
mutation that is based on the similarity of states (as defined in the
sim-max game). More concretely, we start from the idea that similar
states are more easily confused than less similar states. To formalize
this idea, we assume that there is a confusion matrix $C$, where
$C_{ij}$ is the probability that a state \mystate{i} is realized as a
possibly different state \mystate{j}. (There is a slightly different
sense of ``realization'' in the sender and the receiver side, and we
will enlarge on this below.) Based on this, we show two things. For
one, we show how the discrete-time formulation of the replicator
dynamic can be easily adapted to integrate diffusion of behavior due
to confusion of states. This yields the \rdd in its basic
discrete-time formulation. For another, we also show that there exists
a translation of the confusion matrix $C$ into a suitable mutation
matrix $M_C$ such that the effect of $C$ and $M_C$ are ``equivalent''
in a sense to be made precise. This supports the conclusion that the
\rdd can indeed be thought of as a special case of the \rmd.

Based on data from numerical simulations, we show that the inclusion
of mild levels of such stochastic noise in the replicator dynamics
does not disrupt the possibility of evolving communicative efficient
signaling strategies at all. On the contrary, there might even be a
higher-order benefit to the presence of perceptual imprecision, in
that it accelerates converges, but also unifies and regularizes
evolutionary outcomes in such a way that inefficient categorization is
avoided. These results complement similar results by
\citet{OConnor2013:The-Evolution-o}, obtained for a micro-dynamic
extension of reinforcement learning. The \rdd adds to this a more
general, abstract framework that is not tied to the specific
assumptions of turn-based reinforcement.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% TeX-PDF-mode: t
%%% End: