### 2014-10-16 paper/02_noisyRMD.tex

\section{Replicator diffusion dynamic}
\label{sec:repl-diff-dynam}

Against this background, we introduce a new macro-level approach to
the evolution of vague signal use in sim-max games, where the source
of vagueness is the agents' natural inability to sharply distinguish
similar states. This applies to the sender as well as to
the receiver role (albeit in slightly different manner). We call our
dynamic \emph{replicator diffusion dynamic} (\rdd), because it is a
special case of the replicator mutator dynamic (\rmd).

The \rdd was first introduced by \citet{Correia2013:The-Bivalent-Tr}
as a noise-perturbed variant of the replicator dynamic (\rd) in
behavioral strategies. We begin by recapitulating this formulation of
the \rd in Section~\ref{sec:repl-dynam-behav}. Then we introduce the
\rdd in Section~\ref{sec:repl-diff-dynam-1}.
Section~\ref{sec:repl-mutat-dynam} introduces the \rmd, and
Section~\ref{sec:diffusion-as-special} shows in what sense the \rdd is
a special case of the \rdd.

\subsection{Replicator dynamic in behavioral strategies}
\label{sec:repl-dynam-behav}

Fix a signaling game with finite states $\States$, messages $\Messgs$
and acts $\Acts$. Let $\Pr(\cdot) \in \Delta{\States}$ be the prior
distribution over states and $\Util_{\sen,\rec} \mycolon \States
\times \Messgs \times \Acts \rightarrow \mathds{R}$ the senders's and
receiver's utility functions. The sender's behavioral strategies are
functions $\Sstrat \in \Delta(\Messgs)^\States$; the receiver's are
functions $\Rstrat \in \Delta(\Acts)^\Messgs$. Define the expected
utility of choices in each choice point as usual:
\begin{align*}
  \EU(\messg \myts ; \myts \state, \Rstrat) & = \sum_{\act \in \Acts}
  \Rstrat(\act \probbar \messg) \cdot U_\sen(\state, \messg, \act) \\
  \EU(\act \myts ; \myts \messg, \Sstrat) & = \sum_{\state \in
    \States} \Pr(\state) \cdot \Sstrat(\messg \probbar \state) \cdot
  U_\rec(\state, \messg, \act) \,.
\end{align*}
The \emph{fitness} of a behavioral strategy at a choice point is the
frequency-weighted average of expected utilities of each choice, given
the opponent's strategy:
\begin{align*}
  \Phi(\state,\Sstrat, \Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
\EU(\messg \myts ; \myts\state,\Rstrat) \\
\Phi(\messg,\Sstrat, \Rstrat) & = \sum_{\act} \Rstrat(\act \probbar \messg)
\cdot \EU(\act \myts ; \myts \messg,\Sstrat)\,.
\end{align*}
The discrete-time replicator dynamic maps current strategies $\Sstrat$
and $\Rstrat$ to future strategies $\RD(\Sstrat)$ and $\RD(\Rstrat)$
in such a way that changes in frequency are proportional to expected
utilities. For behavioral strategies, the changes take place locally
at each of the agents' choice points:
\begin{align*}
  \RD(\Sstrat)(\messg \probbar \state) & = \frac{\Sstrat(\messg \probbar \state) \cdot
    \EU(\messg \myts ; \myts \state,\Rstrat)} {\Phi(\state,\Sstrat, \Rstrat)} \\
    \RD(\Rstrat)(\act \probbar \messg) & = \frac{\Rstrat(\act \probbar \messg) \cdot
    \EU(\act \myts ; \myts \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
\end{align*}

\subsection{Replicator diffusion dynamic in behavioral strategies}
\label{sec:repl-diff-dynam-1}

The replicator diffusion dynamic adds probabilistic confusion of
similar states to the replicator dynamic. Fix a sim-max game with
$\States = \Acts$ and a confusion matrix $C : \States \times
\States$. $C$ is a row-stochastic matrix whose elements $C_{ij}$ give
the probability that $\state_i$ is realized as $\state_j$. The
confusability of states affects senders and receivers alike, but in
slightly different ways (see
Figure~\ref{fig:noise-perturbation-of-strategies}). For the sender,
$C_{ij}$ is the probability that the actual state \mystate{i} is
perceived as \mystate{j}. For the receiver, $C_{ij}$ is the
probability that \mystate{j} is the interpretation that is actually
formed when \mystate{i} is the intended interpretation.

\begin{figure}
  \centering

    \begin{tikzpicture}[node distance = 2cm, thick]

      \begin{scope}

      \node[rectangle, draw=mycol, fill=mycol!25, thick] (actual)
      {actual state $\mystate{i}$};

      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
      actual] (perceived) {perceived state $\mystate{j}$};

      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
      perceived] (output) {chosen message $\messg$};

      \node[rectangle, thick, below of = output, node distance=1cm]
      (sender) {sender};

      \draw[->] (actual) -> (perceived) node[midway,left] {noise
        $C_{ij}$};

      \draw[->] (perceived) -> (output) node[midway,left] (label)
      {strategy $\Sstrat(\messg \probbar \mystate{j})$};

      \begin{pgfonlayer}{background}
        \node [draw=black!50, fill=black!20,fit=(actual) (label)
        (output)] {};
      \end{pgfonlayer}
    \end{scope}

      \begin{scope}[xshift=5cm]

      \node[rectangle, draw=mycol, fill=mycol!25, thick] (message)
      {observed message $\messg$};

      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
      message] (target) {target interpretation $\mystate{i}$};

      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
      target] (realized) {realized interpretation $\mystate{j}$};

      \node[rectangle, thick, below of = realized, node distance=1cm]
      (sender) {receiver};

      \draw[->] (message) -> (target) node[midway,right] (bla) {strategy
        $\Rstrat(\mystate{i} \probbar \messg)$};

      \draw[->] (target) -> (realized) node[midway,right]  {noise $C_{ij}$};

      \begin{pgfonlayer}{background}
        \node [draw=black!50, fill=black!20,fit = (message) 
        (realized) (bla)] {};
      \end{pgfonlayer}
    \end{scope}

  \end{tikzpicture}

  \caption{Effect of confusion of states on sender and receiver
    choices.}
  \label{fig:noise-perturbation-of-strategies}
\end{figure}

The aggregate effect of confusion of states on behavioral strategies
can be captured in a function that maps behavioral strategies
$\Sstrat$ and $\Rstrat$ to their diffusions $D(\Sstrat)$ and
$D(\Rstrat)$. The idea is that $\Sstrat$ and $\Rstrat$ are what, on
average, the idealized, noise-free behavior would be, while
$D(\Sstrat)$ and $D(\Rstrat)$ are behavioral strategies that describe
the agents' actual noise-perturbed probabilistic behavior. Since the
effect of confusion of states is that behavior at one choice point
percolates to behavior at similar choice points, we speak of
\emph{diffusion of behavior under confusion of states}. If we
conceive of behavioral strategies $\Sstrat$ and $\Rstrat$ as
row-stochastic matrices, the diffusion effect of state confusability
is easily captured by matrix multiplication as a diffusion function
$\Diff_C$ that, based on confusion matrix $C$, maps behavioral
strategies onto their diffused realizations:
\begin{align}
  \label{eq:confusion-function}
  \Diff_C(\Sstrat) & = C \Sstrat &    \Diff_C(\Rstrat) & = \Rstrat C\,.
\end{align}

The discrete-time replicator diffusion dynamic takes the replicator
dynamic as basic, but factors in the confusion of states at each
update step in a sequential update:
\begin{align*}
  \RDD(\Sstrat) & = \Diff_C(\RD(\Sstrat)) &   \RDD(\Rstrat) & = \Diff_C(\RD(\Rstrat)) \,.
\end{align*}
This is equivalent to the following, perhaps more transparent
formulation, given by \citet{Correia2013:The-Bivalent-Tr}:
\begin{align*}
  \RDD(\Sstrat)(\messg \probbar \state_i) & = \sum_{j} C_{ij} \cdot
  \frac{\Sstrat(\messg \probbar \state_j) \cdot
    \EU(\messg \myts ; \myts \state_j,\Rstrat)}
  {\Phi(\state_j,\Sstrat, \Rstrat)} \\
    \RDD(\Rstrat)(\state_i \probbar \messg) & = \sum_{j} C_{ji} \cdot
  \frac{\Rstrat(\state_j \probbar \messg) \cdot
    \EU(\state_j \myts ; \myts \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
\end{align*}

The idea behind the sequential definition of the \rdd is that, at each
time step, strategies are gradually optimized along the current
fitness landscape, as described by the \rd, but the realization of
optimized strategies is noisy, due to similarity of states. It is
obvious that other (discrete-time) evolutionary dynamics can be
subjected to state-confusability in an analogous fashion (see also
Section~\ref{sec:model-interpretation}). In the case of the \rd,
however, perturbation by state-confusion has a prominent close
relative in the replicator mutator dynamic.

\subsection{The replicator mutator dynamic}
\label{sec:repl-mutat-dynam}

The \rmd has been proposed first in the context of signaling game
models for the evolution of grammar
\citep[e.g.][]{KomarovaNiyogi2001:The-Evolutionar,NowakKomarova2001:Evolution-of-Un,Nowak2006:Evolutionary-Dy}. It
adds to the replicator dynamic stochastic mutation that captures, in
the context of grammar evolution, the differential learning success of
different dispositions in first language acquisition. The relation of
the \rmd to other prominent evolutionary dynamics is well understood
\citep{PageNowak2002:Unifying-Evolut}. The \rmd has seen further
fruitful applications in the context of signaling games
\citep[e.g.][]{HutteggerSkyrms2010:Evolutionary-Dy,Deo2014:The-Semantic-an}. It
is therefore desirable to relate the \rdd that we propose here to the
\rmd.

A direct comparison between \rdd and the \rmd is not possible, because
the usual formulation of the latter is in its continuous-time form,
and based on mixed strategies, not behavioral strategies. We therefore
give a discrete-time formulation of the \rmd in behavioral strategies,
which is independently useful. The relation between \rdd and \rmd will
then be plain to see. We start with a delineation of mixed and
behavioral strategies.

\paragraph{Mixed strategies.} Pure sender (receiver) strategies are
functions $\Spure \in \Messgs^\States$ ($\Rpure \in
\Acts^\Messgs$). Mixed sender (receiver) strategies are functions
$\Smixed \in \Delta(\Messgs^\States)$ ($\Rmixed \in
(\Acts^\Messgs)$). The latter give the relative population frequencies
of the former. We write $\Smixed_i$ for the frequency
$\Smixed(\Spure_i)$ of pure strategy $\Spure_i$. Likewise for the
receiver.

Every mixed strategy $\Smixed$ converts to a unique behavioral strategy defined by:
\begin{align*}
  \Sstrat(\messg \probbar \state) = \sum_{\Spure(\state) = \messg} \Smixed(\Spure)\,.
\end{align*} 
Let $G$ be this mapping from mixed to behavioral strategies. This $G$
is \emph{not} an injection, as many mixed strategies map onto the same
behavioral strategy. In this sense, mixed strategies hold more
information than behavioral strategies about the distribution of pure strategies in a population.

In the context of evolutionary dynamics, we can think of behavioral
strategies as treating each choice point as an independent update site
\citep[e.g.][]{Cressman2003:Evolutionary-Dy,Sandholm2013:Population-Game}. Dynamics
for mixed strategies conceive of agents as adjusting their whole pure
strategy globally. In this way, dependencies between different choices
at different choice points can matter to the evolutionary path. In
contrast, dynamics for behavioral strategies conceive of agents as
adjusting their strategies locally at each choice point, regardless of
what they are disposed to do at other choice points. The latter
perspective is then more coarse-grained, and not necessarily
equivalent to the former. But an advantage of dynamics in behavioral
strategies is that they reduce complexity, which is especially helpful
for computational simulations of larger games, like the ones we are
interested in here.

\paragraph{Replicator dynamic in mixed strategies.} To understand the
\rmd in mixed strategies, we first introduce the \rd in mixed
strategies. Let $F_i^{\Rmixed}$ be $\Spure_i$'s fitness given
$\Rmixed$ and $F_i^{\Smixed}$ be $\Rpure_i$'s fitness given
$\Smixed$. Then $\Phi(\Smixed,\Rmixed) = \sum_{k} \Smixed_k \cdot
F_k^{\Rmixed}$ is the average fitness in the sender population and
$\Phi(\Rmixed,\Smixed) = \sum_{k} \Rmixed_k \cdot F_k^{\Smixed}$ the
average fitness in the receiver population.\todo{give actual
  definition of fitness etc.?}

The continuous-time replicator dynamic defines the change of frequency
of mixed strategies. Its so-called two-population version, that takes
sender and receiver roles as separate update sites:
\begin{align*}
  \dot{\Smixed_i} & = \Smixed_i \cdot \left ( F_i^{\Rmixed} -
  \Phi(\Smixed,\Rmixed) \right ) &   \dot{\Rmixed_i} &  = \Rmixed_i \cdot \left ( F_i^{\Smixed} -
  \Phi(\Rmixed,\Smixed) \right ) \,.
\end{align*}
Similar to the version for behavioral strategies, a discrete-time
formulation of the \rd for mixed strategies can be conceptualized as an
update function that maps mixed strategies $\Smixed$ and $\Rmixed$
onto mixed strategies $\RD(\Smixed)$ and $\RD(\Rmixed)$ respectively,
such that:
\begin{align*}
  \RD(\Smixed)_i & = \frac{\Smixed_i \cdot F_i^{\Rmixed}}{
    \Phi(\Smixed,\Rmixed)} & \RD(\Rmixed)_i & = \frac{\Rmixed_i \cdot
    F_i^{\Smixed}}{ \Phi(\Rmixed,\Smixed)} \,.
\end{align*}

\paragraph{Replicator mutator dynamic.} The replicator mutator dynamic
extends the \rd by adding probabilistic mutation. Let $Q$ be a
row-stochastic mutation matrix where $Q_{ji}$ gives the probability
that pure sender strategy $\Spure_j$ mutates into
$\Spure_i$. Similarly, let $R$ be a row-stochastic mutation matrix
where $R_{ji}$ gives the probability that pure receiver strategy
$\Rpure_j$ mutates into $\Rpure_i$.

The \rmd is usually given only in its continuous-time form. A
two-population (non-payoff adjusted) formulation is then:
\begin{align*}
  \dot{\Smixed_i} & = \sum_{j}  Q_{ji} \cdot \Smixed_j
    \cdot F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed) &
    \dot{\Rmixed_i} & = \sum_{j}  R_{ji} \cdot \Rmixed_j
    \cdot F_j^{\Smixed} - \Rmixed_i \cdot \Phi(\Rmixed,\Smixed) \,.
\end{align*}
A discrete-time, two-population formulation of the \rmd is this 
\citep[c.f.][97]{PageNowak2002:Unifying-Evolut}:

\begin{align*}
  \RMD(\Smixed)_i & = \sum_{j} Q_{ji} \frac{\Smixed_j \cdot
    F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} & \RDD(\Rmixed)_i & =
  \sum_{j} R_{ji} \frac{\Rmixed_j \cdot F_j^{\Smixed}}{
    \Phi(\Rmixed,\Smixed)}\,.
\end{align*}

% \begin{proof}
%   It suffices to check either sender or receiver part. Focusing on the
%   former, we need to show that the continuous version is derivable
%   from the discrete version if the discrete update steps get
%   infinitely small so that:
%   \begin{align*}
%     \dot{\Smixed_i} & = \RMD(\Smixed)_i - \Smixed_i = \sum_{j} Q_{ji}
%     \frac{\Smixed_j \cdot F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} -
%     \Smixed_i = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
%         F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} - \Smixed_i \right ) \\
%     & = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
%         F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)}{ \Phi(\Smixed,\Rmixed)}  \right )\,.
%   \end{align*}
%   As $\Phi(\Smixed,\Rmixed)$ is a constant rate of change, we can drop
%   it to derive the non-payoff adjusted continuous version above,
%   since:
%   \begin{align*}
%     & \sum_{j} \Smixed_j \left ( Q_{ji} \cdot
%         F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)  \right
%       ) =     \sum_{j} \Smixed_j  Q_{ji} \cdot
%         F_j^{\Rmixed} - \sum_{j} \Smixed_j \cdot \Smixed_i \cdot
%         \Phi(\Smixed,\Rmixed) \\
%        = &    \sum_{j} \Smixed_j  Q_{ji} \cdot
%         F_j^{\Rmixed} - \Smixed_i \cdot
%         \Phi(\Smixed,\Rmixed) 
%   \end{align*}
% \end{proof}

The discrete-time \rmd has the same sequential nature as the \rdd:
first we compute the fitness-driven change according to the standard
replicator dynamic; then we compute the perturbation from
mutation. This becomes clear if we define independent mutation
functions $\Mutate_{Q,R}$ that map mixed strategies onto the outcome
of (one round of) mutation:
\begin{align*}
  \Mutate_Q(\Smixed)_i & =  \sum_j  \Smixed_j \cdot
  Q_{ji} &   \Mutate_R(\Rmixed)_i & =  \sum_{j}  \Rmixed_j \cdot
  R_{ji} \,.
\end{align*}
The discrete-time \rmd given above can then be rewritten as:
\begin{align*}
  \RDD(\Smixed) &= \Mutate_Q(\RD(\Smixed)) &   \RDD(\Rmixed) &= \Mutate_R(\RD(\Rmixed))\,. 
\end{align*}

\subsection{Diffusion as a special kind of mutation}
\label{sec:diffusion-as-special}

The sequential reformulation of the \rmd in terms of a sequence of
updates already closely resembles our initial formulation of the
\rdd. However, there is a major difference between these. While the
\rdd is a dynamic for behavioral strategies, the \rmd, as taken from
the literature, applies to mixed strategies. Similarly, the
confusability of states is an operation that is straightforwardly
definable for behavioral strategies (see
Equations~(\ref{eq:confusion-function})), while mutation in the \rmd
is defined as mutation from one pure strategy to another. Nonetheless,
it is possible to map each confusion matrix to a mutation matrix in a
natural way so that the correspondence between mixed and behavioral
strategies are conserved. This is the main result presented in this
section, and it is our justification for regarding the \rdd as the
behavioral-strategy analogue of the replicator-mutator dynamic when
the only source of mutation is confusability of states.

\begin{figure}
  \centering

  \begin{tikzpicture}[node distance = 2cm]

    \begin{scope}
      \node[] (mixed) {$\Smixed$};

      \node[below of = mixed] (mutated) {$Q^C(\Smixed)$};

      \node[left of = mixed, node distance = 1.5cm] (dummyl) {};

      \node[right of = mixed, node distance = 1.5cm] (dummyr) {};

      \draw[->, dotted, thick] (mixed) -> (mutated);

      \begin{pgfonlayer}{background}
        \node [draw=black!50, fill=black!20,fit = (dummyl) 
        (mutated) (dummyr)] (mixedSpace) {};
      \end{pgfonlayer}

      \node[below of = mixedSpace, node distance = 1.7cm] {space of
        mixed strategies};
    \end{scope}

    \begin{scope}[xshift=6cm]
      \node[] (behavioral) {$\Sstrat$};

      \node[below of = behavioral] (mutatedBeh) {$C(\Sstrat)$};

      \node[left of = behavioral, node distance = 1.5cm] (dummylB) {};

      \node[right of = behavioral, node distance = 1.5cm] (dummyrB) {};

      \draw[->, dotted, thick] (behavioral) -> (mutatedBeh);

      \begin{pgfonlayer}{background}
        \node [draw=black!50, fill=black!20,
              fit = (dummylB) (mutatedBeh) (dummyrB)] (behSpace) {};
      \end{pgfonlayer}

      \node[below of = behSpace, node distance = 1.7cm] {space of
        behavioral strategies};
    \end{scope}

    \draw[->,thick,dotted] (mixed) -> (behavioral) node[midway,
    above]{$G(\Smixed)$};

    \draw[->,thick,dotted] (mutated) -> (mutatedBeh) node[midway, below]{$G(Q^C(\Smixed))$};

  \end{tikzpicture}

  \caption{Correspondence between state-confusion and mutation.}
  \label{fig:correspondence-result}
\end{figure}

More concretely, since there is a non-injective mapping $G$ from
behavioral to mixed strategies, we would like to show that there is a
systematic translation from a confusion matrix $C$ to a mutation
matrices $Q^C$ such that for each mixed sender strategy $\Smixed$, its
unique corresponding behavioral strategy $\Sstrat = G(\Smixed)$ has a
confusion-perturbation $C(\Sstrat)$ that corresponds, via $G$, to the
mutation-perturbation $Q^C(\Smixed)$ (see
Figure~\ref{fig:correspondence-result}). In other words, we
hypothesize that a confusion matrix $C$ should give rise to a unique
mutation matrix $Q^C$ so that whenever $G(\Smixed) = \Sstrat$ we also
have $G(Q^C(\Smixed)) = C(\Sstrat)$. Similarly, for the receiver. This
is the content of Theorem~\ref{thm:Correspondence} given below, a
proof for which is in Appendix~\ref{sec:proofs}. The remainder of this
section introduces the relevant construction that maps arbitrary $C$
to suitable $Q^C$.

\paragraph{Confusion-based mutations.} There are natural conversions
of $C$ into $Q^C$ and $R^C$. The case for the receiver is easier, so
we start with that.

The probability that $\Rpure_i$ is realized as $\Rpure_j$ under
diffusion with $C$ is the product of the probabilities, for each
$\messg$, that the state $\Rpure_i(\messg)$ is perceived as state
$\Rpure_j(\messg)$. Abusing notation by referring to the indices of
states $\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ with
$\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ directly, we define:
\begin{align}
  \label{eq:construction-rec}
  R^C_{ij} = \prod_{\messg} C_{\Rpure_i(\messg)\Rpure_j(\messg)}\,.
\end{align}

Now look at the sender. The probability that $\Spure_j$ is realized as
$\Spure_i$ under diffusion with $C$ is the product of the
probabilities, over all states $\mystate{k}$, that the message
$\Spure_i(\mystate{k})$, that $\Spure_i$ would produce at state
$\mystate{k}$ in the absence of noise, is produced by a noisy
realization of $\Spure_j$, which is the probability $\sum_{\state_l
  \in \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}$ that the state
$\mystate{k}$ is realized as a state $\mystate{l}$ which $\Spure_j$
would map unto $\Spure_i(\mystate{k})$. So, define:
\begin{align}
  \label{eq:construction-sen}
  Q^C_{ji} = \prod_{\state_k} \sum_{\state_l \in
    \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}\,.
\end{align}

For example, consider a signaling game with two states and two
messages. Let the confusion matrix be:
\begin{align*}
  C=
  \begin{pmatrix}
    .8 & .2 \\
    .2 & .8 
  \end{pmatrix}\,.
\end{align*}
The resulting mutation matrices are:
\begin{align*}
Q^C & = \bordermatrix{ ~ & 11 & 12 & 21 & 22 \cr
                      11 & 1 & 0 & 0 & 0 \cr
                      12 & .16 & .64 & .04 & .16 \cr
                      21 & .16 & .04 & .64 & .16 \cr
                      22 & 0 & 0 & 0 & 1 \cr}
&                    
  R^C & = \bordermatrix{ ~ & 11 & 12 & 21 & 22 \cr
                      11 & .64 & .16 & .16 & .04 \cr
                      12 & .16 & .64 & .04 & .16 \cr
                      21 & .16 & .04 & .64 & .16 \cr
                      22 & .04 & .12 & .12 & .64 \cr}\,.
\end{align*}
Here, a pair like $21$, for example, refers to a pure sender strategy
with $\Spure(\state_1)=\messg_2$ and $\Spure(\state_2) =
\messg_1$. Similarly for the receiver.

Based on the translations from confusion matrices $C$ to mutation
matrices $Q^C$ and $R^C$ in Equations~(\ref{eq:construction-rec}) and
(\ref{eq:construction-sen}), we can finally formulate the
desired correspondence result, that reveals the \rdd as a special case
of the \rmd in behavioral strategies where the only source of mutation is
imperfect discriminability of states.

\begin{theorem}
  \label{thm:Correspondence}
  (i) If $G(\Smixed) = \Sstrat$, then $G(\Mutate_{Q^C}(\Smixed)) =
  \Diff_C(\Sstrat)$. And, (ii) if $G(\Rmixed) = \Rstrat$, then
  $G(\Mutate_{R^C}(\Rmixed)) = \Diff_C(\Rstrat)$.
\end{theorem}

\noindent A proof is given in Appendix~\ref{sec:proofs}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% TeX-PDF-mode: t
%%% End: