### 2014-10-16 paper/04_discussion.tex

\section{Discussion}
\label{sec:discussion}

We have introduced the \rdd as a new variant of the \rmd that
implements stochastic noise in the form of confusability of similar
states. The probabilities of confusing similar states is easily
implemented for behavioral strategies, but, as seen in
Section~\ref{sec:repl-diff-dynam}, can also be translated into a
mutation matrix for pure strategies. This allows, in principle,
several conceptual interpretations of the \rdd, on which we will
briefly elaborate below in Section~\ref{sec:model-interpretation}.

Next to this technical contribution, the results of our numerical
simulations also advance our understanding of the possibility of
evolving regular and efficient categories despite their (higher-order)
vagueness. Section~\ref{sec:relat-with-prev} zooms in on the relation
with closely related accounts once more, to delineate the present
approach more precisely.

\subsection{Model interpretation}
\label{sec:model-interpretation}

We mentioned in passing in Section~\ref{sec:repl-diff-dynam-1} that
adding diffusion to other discrete-time evolutionary dynamics that
operate on behavioral strategies is entirely straightforward. We
could, for instance, easily diffuse the outcome of a best-response
dynamic at each time step. That would make sense if we thought of
agents as prone to confusing similar states, but otherwise rationally
optimizing behavior. The reason that we chose the \rd to combine with
diffusion is twofold. A minor reason is that it makes for a
conceptually interesting link with the \rmd. A more important reason
is that the \rd is especially versatile and non-committal about what
the exact process of adaption is that is being modelled.

Originally the \rd was introduced as mathematical model of evolution
under asexual reproduction, motivated by concepts from the theory of
natural selection \citep{TaylorJonker1978:Evolutionary-St}. The most
conservative interpretation of the \rdd in our present context is thus
a biological one: we can imagine signaling strategies as innate or
fixed behaviorial tendencies of organisms, steps in the evolutionary
process as successive generations, and selection as capturing the
reproductive advantage of fitter individuals. This interpretation,
strictly speaking, requires formalization in terms of mixed strategies
via the \rmd, and possibly also a symmetrizing of the game, so that
every agent is assumed to have a unique sender and receiver role at
the same time, the pair of which is bequeathed onto the next
generation. Diffusion, in this context, could be either of two
things. It could be differential mutation probabilities in line with
other interpretations of mutation in the \rmd
\citep[e.g.][]{NowakKomarova2001:Evolution-of-Un,KomarovaNiyogi2001:The-Evolutionar}:
pure strategies are not faithfully reproduced, e.g. by imperfections
of learning or underdetermination of input, and mistakes in
inheriting pure strategies are more likely if they result from the
confusion of similar states. Another possible biological
interpretation of the \rdd, is that inheritance is faithful, but
strategies are noisily realized. Strategies are not necessarily
selected for what they are, but rather for how they are realized, once
noise is factored in.

But the replicator dynamic is not only a model of biological
evolution. It can also be interpreted as a high-level description of
the likely development of other behavioral adaptation processes, like
differential imitation \citep[see][for various derivations of the
\rd]{Sandholm2013:Population-Game}. Under this interpretation, the \rd
is consistent with the idea that what is subject to evolutionary
optimization are not organisms but behavior: agents or systems can
adapt their behavior to the perceived environment or agents can adopt
strategies from others. Fitness captures the success of behavioral
patterns, which in the case of language can be thought of in terms of
communicative success.  Differential reproduction represents the
tendency of more successful strategies to be more likely realized, be
it through imitation or some solitary learning or optimization
mechanism.

Under this non-biological interpretation, we have again two options of
picturing what diffusion is, similar to the biological cases
before. One possibility is that diffusion is noise in the adoption of
strategies, e.g., by conditional imitation, of strategies by other
signalers. Another possibility is that diffusion is again noise on the
realization of strategies: while behavior is optimized to be efficient
(be it due to learning, introspection or imitation), realization of
strategies is bound to be noisy due to confusability of similar
states.

We believe that all of the four mentioned interpretations are, on
first approximations, feasible conceptualizations of the \rdd, and
that it is a good thing to know of a working account of vague
signaling that sketches where fitness-based selection under
state-confusability will take is, abstracting away from the details of
actually playing the game, inheriting, imitating or otherwise
optimizing behavior in whatever particular way. It is a good thing to
know this on the macro-level, especially since there are also
micro-level accounts that nicely complement the picture. We turn to
one such next.

\subsection{Relation with previous accounts}
\label{sec:relat-with-prev}

\paragraph{Generalized reinforcement.}
% Repeated below...
%\citet{OConnor2013:The-Evolution-o} introduced a generalization of
%Herrnstein reinforcement learning for sim-max games and showed that
%this not only leads to vague signaling patterns, but can speed-up
%evolution of efficient signaling strategies, especially in games with
%many states.

Under \emph{Herrnstein reinforcement learning}, sender and
receiver play the game repeatedly and adjust their dispositions to act
after each round of play, in such a way that the actual (non-negative)
payoff gained in the current interaction is added to the
non-normalized propensities for repeating the same behavior. For
sim-max games, this means that when the sender chose $\messg$ in state
$\state$, and this resulted in some non-negative payoff (which is
guaranteed by our choice of utility function), the probability that
the sender chooses $\messg$ again in $\state$ is increased, but
nothing else changes. In particular, the sender's behavior in other
choice points does not change. \emph{Generalized reinforcement
  learning} is different here. When the use of $\messg$ in $\state$
gave positive payoff, then not only will its future use probability be
promoted at $\state$ but also at other states, proportional to how
similar these are to $\state$. Similar amendments take care of the way
that the receiver updates his choice dispositions.

\citet{OConnor2013:The-Evolution-o} shows that this extension not only
leads to vague signaling of the appropriate kind, but also speeds up
learning in such a way that, especially for sim-max games with higher
numbers of states, higher levels of communicative success are reached
in shorter learning periods than is possible without stimulus
generalization. Technically, this result is partly due to the fact
that signalers make bigger changes to their behavioral strategies
after each round of play under generalized reinforcement than under
the Herrnstein variety. But that only explains the speed of
adaptation, not necessarily that generalization also leads to
regularity and communicative efficiency.

Diffusion of strategies in the \rdd can also be conceived of as a form
of generalization, and works in large part quite analogous to stimulus
generalization in \citeauthor{OConnor2013:The-Evolution-o}'s
approach. But there are still
differences. \citeauthor{OConnor2013:The-Evolution-o} showed that
generalized reinforcement learning can speed up the development of
efficient signaling, especially for higher numbers of
states. Complementing this, we showed that diffusion regularizes
evolutionary outcomes and prevents meandering around sub-optimal and
non-convex strategies, also for low numbers of states.

Conceptually, the \rdd is a more abstract framework than generalized
reinforcement learning. The latter is foremost motivated as a learning
dynamic that has two players adapt their individual strategies after
each concrete round of play. In contrast, the \rdd describes a more
abstract, average dynamical change in behavioral
dispositions. Although the behavior of (some forms of) reinforcement
learning mirror those of the \rd (at some stage in time)
(\cite{BorgersSarin997:Learning-Throug,HopkinsPosch2005:Attainability-o,Beggs2005:On-the-Converge}),
this does not mean that \emph{generalized} reinforcement learning is
also necessarily a plausible high-level description of, say,
generalized learning in a population of agents. Seen in this light,
generalized reinforcement learning and the \rdd nicely complement each
other, as similarly-minded accounts operating at different levels of
abstraction.

\paragraph{Quantal response equilibria.}
\citet{FrankeJager2010:Vagueness-Signa} suggested a number of ways in
which information processing limitations of signaling agents could
lead to vague strategies. The model that is most clearly related to
the present approach uses the notion of a \emph{quantal response},
also known as a \emph{logit response} or a \emph{soft-max function}
\citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,McKelveyPalfrey1995:Quantal-Respons,McKelveyPalfrey1998:Quantal-Respons,GoereeHolt2008:Quantal-Respons}. A
quantal response function is a paramterized generalization of the
classic best response function. For example, if $U \mycolon \Acts
\rightarrow \mathds{R}$ is the measure of expected utility over
choices $\Acts$ of an agent, then a best response function would have
the agent choose $\act$ only if $U(\act) = \max_{\act' \in \Acts}
U(a)$. A quantal response function rather assumes that agents would
choose $\act$ with a probability proportional to $\expo(\lambda \cdot
U(\act))$, where $\lambda$ is a rationality parameter. If $\lambda
\rightarrow \infty$ we retrieve the behavior of the best response
function, but if it is positive but finite, any choice $\act$ will
receive a positive probability, but acts with higher expected utility
will be more likely. The underlying motivation for this choice rule is
the assumption that there is noise in the computation of expected
utilities and/or in maximization of expected utilities. Consequently,
choices with almost equal expected utilities will be chosen with
almost equal probability (for moderate values of $\lambda$).

\citet{FrankeJager2010:Vagueness-Signa} show that quantal response
equilibria of sim-max games, i.e., pairs of sender and receiver
strategies such that the sender strategy is the quantal response to
the expected utilities under the receiver strategy and vice versa, can
show the desired marks of vague
signaling. Figure~\ref{fig:exampleQRE_stratsA} gives an example of a
quantal response equilibrium for a sim-max game, as used in our set-up
but with higher tolerance $\toler = 0.5$. Sender and receiver
strategies look very much like what evolves under \rdd with modest
values of perceptual imprecision.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/exampleStratQRE_tolerance05.pdf}
    \caption{$\ns = 90$, $\lambda = 15$, $\toler = 0.5$}
    \label{fig:exampleQRE_stratsA}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/exampleStratQRE_tolerance005.pdf}
    \caption{$\ns = 90$, $\lambda = 15$, $\toler = 0.05$}
    \label{fig:exampleQRE_stratsB}
  \end{subfigure}

  \caption{Examples of vague quantal response equilibria.}
  \label{fig:exampleQREs}
\end{figure}

But not all quantal response equilibria are equally plausible, and the
ones that are not suggest that it is less natural to think of
vagueness as arising from of errors in the computation or maximization
of expected utility than that it arises from confusion of similar
states. To see what the problem is, we can look at cases like given in
Figure~\ref{fig:exampleQRE_stratsB}, which is the quantal response
equilibrium for a game with lower tolerance $\toler = 0.05$. Unlike
what evolves under \rdd in this case, sender strategies have vague
boundaries also towards the end of the unit intervals. Technically,
this is because quantal responses equalize message use far away from
the ``prototypical'' interpretation, not just in-between categories,
so to speak. This, in turn, is because quantal responses introduce
noise into the decision making at the level of computing or maximizing
expected utility of choices.

That this is conceptually odd shows even more clearly in a case where
the state space is intuitively unbounded, as for instance for the
property ``tall''. If the usual interpretation of a ``tall man'' peaks
at around, say, 195cm then when meeting a giant of $n$ meters, senders
would, according to the quantal response approach, be ever more
inclined to describe the giant indifferently as either ``tall'' or
``short'' the larger $n$ gets. This is because, as the distance from
the prototype increases for larger $n$, the difference between the
expected utilities of saying ``short'' or ``tall'' will converge to
zero. 
%Whence that the quantal response approach would predict that
%senders would grow indifferent between choice of antonyms as $n$
%grows, which seems weird.

Admittedly, this argument hinges on the choice of utility
function. Still, to the extent that the chosen lower bounded utility
functions are reasonable ---and we think they are very reasonable---,
the case suggests that quantal responses are not a good model,
intuitively speaking, for why linguistic categories are
vague. Vagueness is more plausibly an effect of perceptual confusion
of similar states, than of computational errors in maximizing expected
utility.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% TeX-PDF-mode: t
%%% End: