### 2014-11-04 paper/paper_short.tex
paper short version

\documentclass[fleqn,reqno,10pt]{article}

%========================================
% Packages
%========================================

\usepackage[]{../helpers/mypackages}
%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
%\bibliography{MyRefGlobal}
\bibliography{../helpers/MyRefGlobal}
\bibliography{paper} 
\usepackage{../helpers/myenvironments}
\usepackage{../helpers/mycommands}
\usepackage{todonotes}
\usepackage{subcaption}

%========================================
% Standard Layout
%========================================

% Itemize
\renewcommand{\labelitemi}{\large{$\mathbf{\cdot}$}}    % itemize symbols
\renewcommand{\labelitemii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiv}{\large{$\mathbf{\cdot}$}}
% Description
\renewcommand{\descriptionlabel}[1]{\hspace\labelsep\textsc{#1}}

% Figure Captions
\usepackage{caption} % use corresponding myfiguresize!
\setlength{\captionmargin}{20pt}
\renewcommand{\captionfont}{\small}
\setlength{\belowcaptionskip}{7pt} % standard is 0pt

%========================================
% Additional layout & commands
%========================================

\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}

% Annotations
\newcommand{\mytodo}[2]{\todo[inline,color=yellow,author=#1]{#2}}
\newcommand{\question}[2]{\todo[inline,color=blue,author=#1]{#2}}
\newcommand{\answer}[2]{\todo[inline,color=green,author=#1]{#2}}

\newcommand{\rd}{\acro{rd}} % replicator dynamic
\newcommand{\rmd}{\acro{rmd}} % replicator mutator dynamic
\newcommand{\rdd}{\acro{rdd}} % replicator diffusion dynamic
\newcommand{\RD}{\ensuremath{\mathrm{RD}}} % replicator dynamic
\newcommand{\RDD}{\ensuremath{\mathrm{RDD}}} % replicator diffusion dynamic
\newcommand{\RMD}{\ensuremath{\mathrm{RMD}}} % replicator mutator

\newcommand{\Diff}{\ensuremath{\mathrm{D}}} % Difusion 
\newcommand{\Mutate}{\ensuremath{\mathrm{M}}} % Mutation

\newcommand{\impairment}{\ensuremath{\alpha}} % impairment
\newcommand{\toler}{\ensuremath{\beta}} % tolerance
\newcommand{\ns}{\ensuremath{n_s}} % number of states

\newcommand{\similarity}{\ensuremath{\mathrm{Sim}}} % similarity function

% \linespread{1.5}

\title{Vagueness, Noise, and Signaling}
% \author{Michael Franke \and Jos\'e Pedro Correia}
\date{}

\begin{document}

\maketitle

\begin{abstract}
  Signaling games are appealing models for studying the evolution of
  meaning, but typically do not explain vagueness as a feature of
  successful signaling.  Complementing recent like-minded models, we
  introduce the replicator diffusion dynamic as a special case of the
  replicator mutator dynamic that combines evolutionary pressure on
  optimal signal use with stochastic noise in the processing of
  similar stimuli. Applying this new dynamic to a generalization of
  Lewis' signaling games, we show that stochastic imprecision leads to
  vague, yet communicative efficient signal use, and, moreover, also
  unifies evolutionary outcomes and helps avoid suboptimal
  categorization.
\end{abstract}

Many of our concepts and words are vague. A vague category knows clear
cases that fall under it, clear cases that do not, and also so-called
borderline cases. Borderline cases do not clearly apply, nor do they
clearly not apply, and there may be differences between borderline
cases in terms of how well they represent the category in
question. Vagueness does not seem to dramatically affect the success
of everyday communication, but it is troublesome for some of the most
prominent theories of language and meaning. This is especially the
case for the logico-semantic tradition of Frege, Russell and the young
Wittgenstein which is challenged by the paradoxes vagueness gives rise
to. But there are other intriguing aspects about vagueness.

The puzzle that we are concerned with here is how vagueness could
arise and be maintained in the first place. This is a serious worry to
functionalist accounts that maintain that our concepts and language
use evolved in order to be efficient. Since the existence of unclear
borderline cases seems to entail inefficiency of categorization or
communication, the challenge, succinctly put forward by
\citet{Lipman2009:Why-is-Language}, is to explain how vagueness can
exist despite its obvious and demonstrable functional deficiency under
evolutionary pressure to be optimally expressive. We address
\emph{Lipman's problem} as a technical challenge,
namely to find a conceptually sound and mathematically coherent formal
model of the evolution of categorization and language use that yields
by-and-large efficient categories \emph{and} vagueness.

A number of authors have recently tried to explain why vagueness
evolved as something that is itself useful
\citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,BlumeBoard2013:Intentional-Vag}.
Others have argued that vagueness is a natural byproduct of
limitations in information processing
\citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o}. This
paper contributes to the latter line of thought. Concretely, we
introduce the replicator diffusion dynamics ---a novel variant of the
replicator mutator dynamic--- that integrates stochastic noise on the
differential confusability of similar stimuli. The dynamic proposed
here generalizes and complements previous like-minded accounts. We
show that stochastic noise can not only lead to vague, yet
communicative efficient signal use, but can also unify evolutionary
outcomes and help avoid suboptimal categorization.

The next section introduces the background against which the work
presented here can be appreciated. Section~\ref{sec:repl-diff-dynam}
introduces the replicator diffusion dynamic and elaborates on its
relation with the replicator mutator
dynamic. Section~\ref{sec:exploring-rdd} explores the replicator
diffusion dynamic. Section~\ref{sec:discussion} reflects and compares
our approach to others.

\section{Background}
\label{sec:background}

% The view that vagueness is a natural concomitant of cognitive
% limitations of language users has been formalized in a number of ways,
% using evolutionary game theory and certain generalizations of Lewis'
% signaling games \citep{Lewis_1969:Convention}, so called
% \emph{similarity-maximizing games}, or \emph{sim-max games}, for short
% \citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct}.
% %\citep{FrankeJager2010:Vagueness-Signa,OConnor2013:Evolving-Percep,OConnor2013:The-Evolution-o}.
% Our contribution is best seen in relation to these accounts, as it also
% relies on sim-max games. Let's introduce these first, and then zoom in
% on the problem of vagueness.

\subsection{Sim-max games \& conceptual spaces}

Signaling games, as introduced by \citet{Lewis_1969:Convention}, have
a sender and a receiver. The sender knows the true state of the world,
but the receiver does not. The sender can select a signal, or message,
to reveal to the receiver, who then chooses an act. In Lewis' games,
if the receiver chooses the act that corresponds to the actual state,
the play is a success, otherwise a failure. Certain regular
combinations of sender signaling and receiver reaction make messages
meaningful, in the sense that their use is correlated systematically
to certain states or acts. To investigate the conditions under which
such meaning-generating behavior can evolve is a highly interesting
topic that we are only beginning to fully understand
\citep[e.g.][]{Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,Huttegger2007:Evolution-and-t,Pawlowitsch2008:Why-Evolution-d,Barrett2009:The-Evolution-o,HutteggerSkyrms2010:Evolutionary-Dy,Skyrms2010:Signals}.

Similarity-maximizing games, or \emph{sim-max games} for short, are
generalizations of Lewis' games where different states are allowed to
be more or less similar to one another. While Lewis' games treated
communicative success as a matter of black and white, sim-max games
allow for a gradient notion of communicative success: the more similar
the receiver's interpretation is to the actual state, the more
successful the play is taken to be. Signaling games with
utility-relevant similarities in the state space are fairly standard
in economics
\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor},
but have received particular attention in a more philosophical and
linguistic context for reasons that will become clear presently
\citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag}.

A sim-max game, in the relevant sense here, consists of a set of
states $\States$, a set of messages $\Messgs$ with much fever messages
than states, a prior probability distribution $\Pr \in
\Delta(\States)$ that gives the occurrence probabilities of states, a
similarity metric on states $\similarity \mycolon \States \times
\States \rightarrow \mathds{R}$, and a utility function $\Utils
\mycolon \States \times \States \rightarrow \mathds{R}$. We identify
the receiver's acts with the states of the world, so that the game is
one of guessing the actual state, so to speak. We also assume that
sender's and receiver's interests are alike, so we only have one
utility function. We do not consider message costs, so utilities only
depend on the actual state, and the receiver's response. The
similarity function should satisfy the usual requirements for a metric
and the utility function should be a monotonically decreasing function
of similarity.

\citet{JagerMetzger2011:Voronoi-Languag} showed that the
evolutionarily stable states of sim-max games are remarkably
systematic. Their results were obtained for games with infinitely many
states in $n$-dimensional Euclidean space $\States \subseteq
\mathds{R}^n$ and a quadratic loss function for utilities
$\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} -
\mystate{2})^2$. For these games, the evolutionarily stable states are
demonstrably so-called Voronoi languages. Roughly put, a Voronoi
language is a pair of sender and receiver strategies, such that the
sender strategy partitions the state space into convex categories,
while the receiver's interpretations are the central spots in each
category. For example, if \States is the unit interval and all states
are equiprobable, an \ESS of a sim-max game with two messages would
have the sender use one message exclusively for all points in the
lower half of the unit interval and another for all points in the upper
half; the receiver's interpretations of messages are the central
points in the respective intervals.

This result is interesting, because it demonstrates that signaling can
impose orderly categories on a metric space, without that being the
ulterior purpose of it all. Sender and receiver can be distinct
entities, whose purpose is to communicate effectively about the actual
state. In that case, evolving Voronoi languages would explain why
\emph{linguistic categories} are well-behaved and orderly in the way
they appear to be. More abstractly, sender and receiver can also be
thought of as distinct modules in a single system, where the first
module must discretize the information by selecting a small sample of,
suggestively, category labels. These are passed to a second module
that tries to decode the original information. In this case, evolving
Voronoi languages would explain why \emph{conceptual categories} are
well-behaved and orderly in the way that they appear to be.

Seen in this light, sim-max games may provide a foundation to
approaches in cognitive semantics that rely on the notion of
conceptual spaces.  \citet[][70--77]{Gardenfors2000:Conceptual-Spac},
for example, has prominently argued that natural categories are convex
regions in conceptual space. If the conceptual space has a suitable
metric, convex categories can be derived from a set of prototypes. The
category corresponding to prototype $p$ is the set of points that are
more similar to $p$ than to any other
\citep[e.g.][]{OkabeBoots2000:Spatial-Tessell}. In this way,
\citet{Gardenfors2000:Conceptual-Spac} argues, an efficient
categorization system can be obtained: storing the prototypes lets us
recover the categories without having to store each category's
extension. However, what is left unexplained, is where the prototypes
come from, and why we would not see just any distribution of
prototypes as an equally efficient classification system. This is
where sim-max games can contribute a principled approach to deriving,
in an independent way, not only convex categories but also
prototypical exemplars belonging to them.

\subsection{Vague signaling in sim-max games}

This outline of an approach to categorization using sim-max games
leaves many problems unaddressed. One of them is that natural
categories for continuously variable stimuli usually do not have
unique, point-valued prototypes and clear category boundaries. We
would like to account for the possibility of such vagueness, in
particular for the following criteria: (i) clear positive examples of
a vague category should show a gradient transition to clear negative
examples to (so-called \emph{higher-order vagueness}); (ii) prototypes
should likewise be gradient regions, peaking at the center of the
vague category they represent
\citep[e.g.][]{Sainsbury1991:Is-There-Higher,KeefeSmith1997:Vagueness:-A-Re}.

\citet{DouvenDecock2011:Vagueness:-A-Co} show that
\citeauthor{Gardenfors2000:Conceptual-Spac}'s conceptual spaces
approach can be extended to account for the existence of borderline
cases. From the assumption that prototypes are extended, yet convex
regions in conceptual space, they give a construction algorithm that
yields ``collated Voronoi diagrams'' with thick boundaries
representing borderline
regions. \citet{DecockDouven2012:What-is-Graded-} show further how it
is possible to arrive at higher-order vagueness, by weighing in the
distance of different borderline cases to various prototypical
regions. This accounts for the first of the two desiderata mentioned
above, but still assumes that crisp prototype regions must be
given.

An alternative approach is taken by
\citet{FrankeJager2010:Vagueness-Signa} and
\citet{OConnor2013:The-Evolution-o} who show how the above desiderata
can be met by evolving strategies in sim-max games. We should
distinguish micro- and macro-level approaches. Micro-level approaches
look at adaptive behavior of individual agents. Usually, changes in
the behavioral dispositions of agents occur after every single
interaction. In contrast, macro-level approaches outline more
abstract, aggregate dynamics, happening in a population of agents, or
otherwise abstracting from seemingly irrelevant detail. Usually, a
macro-level dynamic captures changes of frequencies of behavioral
types in the population over time.

As for micro-level approaches, \citet{FrankeJager2010:Vagueness-Signa}
show how limited memory of past interactions can lead to vague signal
use, when averaging over a single agent's behavior over time or over
the momentary behavior of a population of several language
users. \citet{OConnor2013:The-Evolution-o} introduces a variant of
reinforcement learning that entails a low-level form of stimulus
generalization. Agents update their behavior after each round of play
in such a way that states similar to the one that actually occurred
are subject to behavioral adjustment as well.

\citet{FrankeJager2010:Vagueness-Signa} also consider a macro-level
approach, using the notion of a \emph{quantal response}. A quantal
response function is a probabilistic choice rule that formalizes the
idea that agents make small mistakes when calculating the expected
utility of choice options. In aggregation, these probabilistic
mistakes lead to systematic ``trembles'' that produce vague signal
use.

The approach we take here is superficially similar to the latter, but
there are crucial differences. For one, we adopt a dynamic perspective
by looking at limited-time outcomes of a dynamic process. For another,
we demonstrate in Section~\ref{sec:discussion} that quantal responses
can give rise to counterintuitive predictions. These counterintuitive
examples suggest that vagueness in sim-max games is not convincingly
explained by appeal to mistakes in calculating expected utility, but
rather, as we assume here, as the result of confusing similar
states. A more in-depth discussion of alternative approaches is
deferred until Section~\ref{sec:discussion}.  Let us first look into
our own proposal in more detail.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Replicator diffusion dynamic}
\label{sec:repl-diff-dynam}

Against this background, we introduce a new macro-level approach,
where the source of vagueness is the agents' natural inability to
sharply distinguish similar states. We call our dynamic
\emph{replicator diffusion dynamic} (\rdd), because it is an extension
of the replicator dynamic (\rd) that adds diffusion of behavior, and a
special case of the replicator mutator dynamic (\rmd).

We begin by recapitulating the formulation of the \rd in
Section~\ref{sec:repl-dynam-behav}, and introduce the \rdd in Section~\ref{sec:repl-diff-dynam-1}.
Section~\ref{sec:repl-mutat-dynam} introduces the \rmd, and
Section~\ref{sec:diffusion-as-special} shows in what sense the \rdd is
a special case of the \rmd.

\subsection{Replicator dynamic in behavioral strategies}
\label{sec:repl-dynam-behav}

Fix a signaling game with finite states $\States$, messages $\Messgs$
and acts $\Acts$. Let $\Pr(\cdot) \in \Delta{\States}$ be the prior
distribution over states and $\Util_{\sen,\rec} \mycolon \States
\times \Messgs \times \Acts \rightarrow \mathds{R}$ the senders's and
receiver's utility functions. The sender's behavioral strategies are
functions $\Sstrat \in \Delta(\Messgs)^\States$; the receiver's are
functions $\Rstrat \in \Delta(\Acts)^\Messgs$. Define the expected
utility of choices in each choice point as usual:
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\act \in \Acts}
  \Rstrat(\act \probbar \messg) \cdot U_\sen(\state, \messg, \act) \\
  \EU(\act, \messg, \Sstrat) & = \sum_{\state \in
    \States} \Pr(\state) \cdot \Sstrat(\messg \probbar \state) \cdot
  U_\rec(\state, \messg, \act) \,.
\end{align*}
The \emph{fitness} of a behavioral strategy at a choice point is the
frequency-weighted average of expected utilities of each choice, given
the opponent's strategy:
\begin{align*}
  \Phi(\state,\Sstrat, \Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
\EU(\messg, \state,\Rstrat) \\
\Phi(\messg,\Sstrat, \Rstrat) & = \sum_{\act} \Rstrat(\act \probbar \messg)
\cdot \EU(\act, \messg,\Sstrat)\,.
\end{align*}
The discrete-time replicator dynamic maps current strategies $\Sstrat$
and $\Rstrat$ to future strategies $\RD(\Sstrat)$ and $\RD(\Rstrat)$
in such a way that changes in frequency are proportional to expected
utilities. For behavioral strategies, the changes take place locally
at each of the agents' choice points:
\begin{align*}
  \RD(\Sstrat)(\messg \probbar \state) & = \frac{\Sstrat(\messg \probbar \state) \cdot
    \EU(\messg, \state,\Rstrat)} {\Phi(\state,\Sstrat, \Rstrat)} \\
    \RD(\Rstrat)(\act \probbar \messg) & = \frac{\Rstrat(\act \probbar \messg) \cdot
    \EU(\act, \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
\end{align*}

\subsection{Replicator diffusion dynamic in behavioral strategies}
\label{sec:repl-diff-dynam-1}

The replicator diffusion dynamic was first introduced by \citet{Correia2013:The-Bivalent-Tr}
as a noise-perturbed variant of the replicator dynamic (\rd) in
behavioral strategies. Fix a sim-max game with
$\States = \Acts$ and a confusion matrix $C : \States \times
\States$. $C$ is a row-stochastic matrix whose elements $C_{ij}$ give
the probability that $\state_i$ is realized as $\state_j$. The
confusability of states affects senders and receivers alike, but in
slightly different ways (see
Figure~\ref{fig:noise-perturbation-of-strategies}). For the sender,
$C_{ij}$ is the probability that the actual state \mystate{i} is
perceived as \mystate{j}. For the receiver, $C_{ij}$ is the
probability that \mystate{j} is the interpretation that is actually
formed when \mystate{i} is the intended interpretation.

\begin{figure}
  \centering

    \begin{tikzpicture}[node distance = 2cm, thick]

      \begin{scope}

      \node[rectangle, draw=mycol, fill=mycol!25, thick] (actual)
      {actual state $\mystate{i}$};

      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
      actual] (perceived) {perceived state $\mystate{j}$};

      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
      perceived] (output) {chosen message $\messg$};

      \node[rectangle, thick, below of = output, node distance=1cm]
      (sender) {sender};

      \draw[->] (actual) -> (perceived) node[midway,left] {noise
        $C_{ij}$};

      \draw[->] (perceived) -> (output) node[midway,left] (label)
      {strategy $\Sstrat(\messg \probbar \mystate{j})$};

%      \begin{pgfonlayer}{background}
%        \node [draw=black!50, fill=black!20,fit=(actual) (label)
%        (output)] {};
%      \end{pgfonlayer}
    \end{scope}

      \begin{scope}[xshift=5cm]

      \node[rectangle, draw=mycol, fill=mycol!25, thick] (message)
      {observed message $\messg$};

      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
      message] (target) {target interpretation $\mystate{i}$};

      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
      target] (realized) {realized interpretation $\mystate{j}$};

      \node[rectangle, thick, below of = realized, node distance=1cm]
      (sender) {receiver};

      \draw[->] (message) -> (target) node[midway,right] (bla) {strategy
        $\Rstrat(\mystate{i} \probbar \messg)$};

      \draw[->] (target) -> (realized) node[midway,right]  {noise $C_{ij}$};

%      \begin{pgfonlayer}{background}
%        \node [draw=black!50, fill=black!20,fit = (message) 
%        (realized) (bla)] {};
%      \end{pgfonlayer}
    \end{scope}

  \end{tikzpicture}

  \caption{Effect of confusion of states on sender and receiver
    choices.}
  \label{fig:noise-perturbation-of-strategies}
\end{figure}

The aggregate effect of confusion of states on behavioral strategies
can be captured in a function that maps behavioral strategies
$\Sstrat$ and $\Rstrat$ to their diffusions $D(\Sstrat)$ and
$D(\Rstrat)$. The idea is that $\Sstrat$ and $\Rstrat$ are what, on
average, the idealized, noise-free behavior would be, while
$D(\Sstrat)$ and $D(\Rstrat)$ are behavioral strategies that describe
the agents' actual noise-perturbed probabilistic behavior. Since the
effect of confusion of states is that behavior at one choice point
percolates to behavior at similar choice points, we speak of
\emph{diffusion of behavior under confusion of states}. If we
conceive of behavioral strategies $\Sstrat$ and $\Rstrat$ as
row-stochastic matrices, the diffusion effect of state confusability
is easily captured by matrix multiplication as a diffusion function
$\Diff_C$ that, based on confusion matrix $C$, maps behavioral
strategies onto their diffused realizations:
\begin{align}
  \label{eq:confusion-function}
  \Diff_C(\Sstrat) & = C \Sstrat &    \Diff_C(\Rstrat) & = \Rstrat C\,.
\end{align}

The discrete-time replicator diffusion dynamic takes the replicator
dynamic as basic, but factors in the confusion of states at each
update step in a sequential update:
\begin{align*}
  \RDD(\Sstrat) & = \Diff_C(\RD(\Sstrat)) &   \RDD(\Rstrat) & = \Diff_C(\RD(\Rstrat)) \,.
\end{align*}
This is equivalent to the following, perhaps more transparent
formulation, given by \citet{Correia2013:The-Bivalent-Tr}:
\begin{align*}
  \RDD(\Sstrat)(\messg \probbar \state_i) & = \sum_{j} C_{ij} \cdot
  \frac{\Sstrat(\messg \probbar \state_j) \cdot
    \EU(\messg, \state_j,\Rstrat)}
  {\Phi(\state_j,\Sstrat, \Rstrat)} \\
    \RDD(\Rstrat)(\state_i \probbar \messg) & = \sum_{j} C_{ji} \cdot
  \frac{\Rstrat(\state_j \probbar \messg) \cdot
    \EU(\state_j, \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
\end{align*}

The idea behind the sequential definition of the \rdd is that, at each
time step, strategies are gradually optimized along the current
fitness landscape, as described by the \rd, but the realization of
optimized strategies is noisy, due to similarity of states. It is
obvious that other (discrete-time) evolutionary dynamics can be
subjected to state-confusability in an analogous fashion (see also
Section~\ref{sec:model-interpretation}). In the case of the \rd,
however, perturbation by state-confusion has a prominent close
relative in the replicator mutator dynamic.

\subsection{The replicator mutator dynamic}
\label{sec:repl-mutat-dynam}

The \rmd has been proposed first in the context of signaling game
models for the evolution of grammar
\citep[e.g.][]{KomarovaNiyogi2001:The-Evolutionar,NowakKomarova2001:Evolution-of-Un}. It
adds stochastic mutation to the replicator dynamic in order to capture, in
the context of grammar evolution, the differential learning success of
different dispositions in first language acquisition. The relation of
the \rmd to other prominent evolutionary dynamics is well understood
\citep{PageNowak2002:Unifying-Evolut}. The \rmd has seen further
fruitful applications in the context of signaling games
\citep[e.g.][]{HutteggerSkyrms2010:Evolutionary-Dy}. It
is therefore desirable to relate the \rdd that we propose here to the
\rmd.

A direct comparison between \rdd and the \rmd is not possible, because
the usual formulation of the latter is in its continuous-time form,
and based on mixed strategies, not behavioral strategies. We therefore
give a discrete-time formulation of the \rmd in behavioral strategies,
which is independently useful. The relation between \rdd and \rmd will
then be plain to see. We start with a delineation of mixed and
behavioral strategies.

\paragraph{Mixed strategies.} Pure sender (receiver) strategies are
functions $\Spure \in \Messgs^\States$ ($\Rpure \in
\Acts^\Messgs$). Mixed sender (receiver) strategies are functions
$\Smixed \in \Delta(\Messgs^\States)$ ($\Rmixed \in
(\Acts^\Messgs)$). The latter give the relative population frequencies
of the former. We write $\Smixed_i$ for the frequency
$\Smixed(\Spure_i)$ of pure strategy $\Spure_i$. Likewise for the
receiver.

Every mixed strategy $\Smixed$ converts to a unique behavioral strategy defined by:
\begin{align}
  \label{eq:CorrespondenceBehavioralMixed}
  \Sstrat(\messg \probbar \state) = \sum_{\Spure(\state) = \messg} \Smixed(\Spure)\,.
\end{align} 
Let $G$ be this mapping from mixed to behavioral strategies. Note that $G$
is \emph{not} an injection, as many mixed strategies map onto the same
behavioral strategy. In this sense, mixed strategies hold more
information than behavioral strategies about the distribution of pure strategies in a population.

In the context of evolutionary dynamics, we can think of behavioral
strategies as treating each choice point as an independent update site
\citep[e.g.][]{Cressman2003:Evolutionary-Dy,Sandholm2013:Population-Game}. Dynamics
for mixed strategies conceive of agents as adjusting their whole pure
strategy globally. In this way, dependencies between different choices
at different choice points can matter to the evolutionary path. In
contrast, dynamics for behavioral strategies conceive of agents as
adjusting their strategies locally at each choice point, regardless of
what they are disposed to do at other choice points. The latter
perspective is then more coarse-grained, and not necessarily
equivalent to the former. But an advantage of dynamics in behavioral
strategies is that they reduce complexity, which is especially helpful
for computational simulations of larger games, like the ones we are
interested in here.

\paragraph{Replicator dynamic in mixed strategies.} To understand the
\rmd in mixed strategies, we first introduce the \rd in mixed
strategies. Let $F_i^{\Rmixed}$ be $\Spure_i$'s fitness given
$\Rmixed$ and $F_i^{\Smixed}$ be $\Rpure_i$'s fitness given
$\Smixed$. Then $\Phi(\Smixed,\Rmixed) = \sum_{k} \Smixed_k \cdot
F_k^{\Rmixed}$ is the average fitness in the sender population and
$\Phi(\Rmixed,\Smixed) = \sum_{k} \Rmixed_k \cdot F_k^{\Smixed}$ the
average fitness in the receiver population.

The continuous-time replicator dynamic defines the change of frequency
of mixed strategies. Its so-called two-population version, that takes
sender and receiver roles as separate update sites:
\begin{align*}
  \dot{\Smixed_i} & = \Smixed_i \cdot \left ( F_i^{\Rmixed} -
  \Phi(\Smixed,\Rmixed) \right ) &   \dot{\Rmixed_i} &  = \Rmixed_i \cdot \left ( F_i^{\Smixed} -
  \Phi(\Rmixed,\Smixed) \right ) \,.
\end{align*}
Similar to the version for behavioral strategies, a discrete-time
formulation of the \rd for mixed strategies can be conceptualized as an
update function that maps mixed strategies $\Smixed$ and $\Rmixed$
onto mixed strategies $\RD(\Smixed)$ and $\RD(\Rmixed)$ respectively,
such that:
\begin{align*}
  \RD(\Smixed)_i & = \frac{\Smixed_i \cdot F_i^{\Rmixed}}{
    \Phi(\Smixed,\Rmixed)} & \RD(\Rmixed)_i & = \frac{\Rmixed_i \cdot
    F_i^{\Smixed}}{ \Phi(\Rmixed,\Smixed)} \,.
\end{align*}

\paragraph{Replicator mutator dynamic.} The replicator mutator dynamic
extends the \rd by adding probabilistic mutation. Let $Q$ be a
row-stochastic mutation matrix where $Q_{ji}$ gives the probability
that pure sender strategy $\Spure_j$ mutates into
$\Spure_i$. Similarly, let $R$ be a row-stochastic mutation matrix
where $R_{ji}$ gives the probability that pure receiver strategy
$\Rpure_j$ mutates into $\Rpure_i$.

The \rmd is usually given only in its continuous-time form. A
two-population (non-payoff adjusted) formulation is then:
\begin{align*}
  \dot{\Smixed_i} & = \sum_{j}  Q_{ji} \cdot \Smixed_j
    \cdot F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed) &
    \dot{\Rmixed_i} & = \sum_{j}  R_{ji} \cdot \Rmixed_j
    \cdot F_j^{\Smixed} - \Rmixed_i \cdot \Phi(\Rmixed,\Smixed) \,.
\end{align*}
A discrete-time, two-population formulation of the \rmd can be defined as follows 
\citep[c.f.][97]{PageNowak2002:Unifying-Evolut}:

\begin{align*}
  \RMD(\Smixed)_i & = \sum_{j} Q_{ji} \frac{\Smixed_j \cdot
    F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} & \RDD(\Rmixed)_i & =
  \sum_{j} R_{ji} \frac{\Rmixed_j \cdot F_j^{\Smixed}}{
    \Phi(\Rmixed,\Smixed)}\,.
\end{align*}

% \begin{proof}
%   It suffices to check either sender or receiver part. Focusing on the
%   former, we need to show that the continuous version is derivable
%   from the discrete version if the discrete update steps get
%   infinitely small so that:
%   \begin{align*}
%     \dot{\Smixed_i} & = \RMD(\Smixed)_i - \Smixed_i = \sum_{j} Q_{ji}
%     \frac{\Smixed_j \cdot F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} -
%     \Smixed_i = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
%         F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} - \Smixed_i \right ) \\
%     & = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
%         F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)}{ \Phi(\Smixed,\Rmixed)}  \right )\,.
%   \end{align*}
%   As $\Phi(\Smixed,\Rmixed)$ is a constant rate of change, we can drop
%   it to derive the non-payoff adjusted continuous version above,
%   since:
%   \begin{align*}
%     & \sum_{j} \Smixed_j \left ( Q_{ji} \cdot
%         F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)  \right
%       ) =     \sum_{j} \Smixed_j  Q_{ji} \cdot
%         F_j^{\Rmixed} - \sum_{j} \Smixed_j \cdot \Smixed_i \cdot
%         \Phi(\Smixed,\Rmixed) \\
%        = &    \sum_{j} \Smixed_j  Q_{ji} \cdot
%         F_j^{\Rmixed} - \Smixed_i \cdot
%         \Phi(\Smixed,\Rmixed) 
%   \end{align*}
% \end{proof}

The discrete-time \rmd has the same sequential nature as the \rdd:
first we compute the fitness-driven change according to the standard
replicator dynamic; then we compute the perturbation from
mutation. This becomes clear if we define independent mutation
functions $\Mutate_{Q,R}$ that map mixed strategies onto the outcome
of (one round of) mutation:
\begin{align}
  \label{eq:Mutation}
  \Mutate_Q(\Smixed)_i & =  \sum_j  \Smixed_j \cdot
  Q_{ji} &   \Mutate_R(\Rmixed)_i & =  \sum_{j}  \Rmixed_j \cdot
  R_{ji} \,.
\end{align}
The discrete-time \rmd given above can then be rewritten as:
\begin{align*}
  \RDD(\Smixed) &= \Mutate_Q(\RD(\Smixed)) &   \RDD(\Rmixed) &= \Mutate_R(\RD(\Rmixed))\,. 
\end{align*}

\subsection{Diffusion as mutation}
\label{sec:diffusion-as-special}

The sequential reformulation of the \rmd in terms of a sequence of
updates already closely resembles our initial formulation of the
\rdd. However, there is a major difference between these. While the
\rdd is a dynamic for behavioral strategies, the \rmd, as taken from
the literature, applies to mixed strategies. Similarly, the
confusability of states is an operation that is straightforwardly
definable for behavioral strategies (see
Equations~(\ref{eq:confusion-function})), while mutation in the \rmd
is defined as mutation from one pure strategy to another. Nonetheless,
it is possible to map each confusion matrix to a mutation matrix in a
natural way so that the correspondence between mixed and behavioral
strategies are conserved. This is the main result presented in this
section, and it is our justification for regarding the \rdd as the
behavioral-strategy analogue of the replicator-mutator dynamic when
the only source of mutation is confusability of states.

\begin{figure}
  \centering

  \begin{tikzpicture}[node distance = 1cm]

    \begin{scope}
      \node[] (mixed) {$\Smixed$};

      \node[below of = mixed] (mutated) {$Q^C(\Smixed)$};

      \node[left of = mixed, node distance = 1.5cm] (dummyl) {};

      \node[right of = mixed, node distance = 1.5cm] (dummyr) {};

      \draw[->, dotted, thick] (mixed) -> (mutated);

      \begin{pgfonlayer}{background}
        \node [draw=black!50, fill=black!10,fit = (dummyl) 
        (mutated) (dummyr)] (mixedSpace) {};
      \end{pgfonlayer}

      \node[below of = mixedSpace, node distance = 1.7cm] {space of
        mixed strategies};
    \end{scope}

    \begin{scope}[xshift=6cm]
      \node[] (behavioral) {$\Sstrat$};

      \node[below of = behavioral] (mutatedBeh) {$C(\Sstrat)$};

      \node[left of = behavioral, node distance = 1.5cm] (dummylB) {};

      \node[right of = behavioral, node distance = 1.5cm] (dummyrB) {};

      \draw[->, dotted, thick] (behavioral) -> (mutatedBeh);

      \begin{pgfonlayer}{background}
        \node [draw=black!50, fill=black!10,
              fit = (dummylB) (mutatedBeh) (dummyrB)] (behSpace) {};
      \end{pgfonlayer}

      \node[below of = behSpace, node distance = 1.7cm] {space of
        behavioral strategies};
    \end{scope}

    \draw[->,thick,dotted] (mixed) -> (behavioral) node[midway,
    above]{$G(\Smixed)$};

    \draw[->,thick,dotted] (mutated) -> (mutatedBeh) node[midway, below]{$G(Q^C(\Smixed))$};

  \end{tikzpicture}

  \caption{Correspondence between state-confusion and mutation.}
  \label{fig:correspondence-result}
\end{figure}

More concretely, since there is a non-injective mapping $G$ from
behavioral to mixed strategies, we would like to show that there is a
systematic translation from a confusion matrix $C$ to a mutation
matrices $Q^C$ such that for each mixed sender strategy $\Smixed$, its
unique corresponding behavioral strategy $\Sstrat = G(\Smixed)$ has a
confusion-perturbation $C(\Sstrat)$ that corresponds, via $G$, to the
mutation-perturbation $Q^C(\Smixed)$ (see
Figure~\ref{fig:correspondence-result}). In other words, we
hypothesize that a confusion matrix $C$ should give rise to a unique
mutation matrix $Q^C$ so that whenever $G(\Smixed) = \Sstrat$ we also
have $G(Q^C(\Smixed)) = C(\Sstrat)$. Similarly, for the receiver.
%This is the content of Theorem~\ref{thm:Correspondence} given below, a
%proof for which is in Appendix~\ref{sec:proofs}. The remainder of this
%section introduces the relevant construction that maps arbitrary $C$
%to suitable $Q^C$.

\paragraph{Confusion-based mutations.} There are natural conversions
of $C$ into $Q^C$ and $R^C$. The case for the receiver is easier, so
we start with that.

The probability that $\Rpure_i$ is realized as $\Rpure_j$ under
diffusion with $C$ is the product of the probabilities, for each
$\messg$, that the state $\Rpure_i(\messg)$ is perceived as state
$\Rpure_j(\messg)$. Abusing notation by referring to the indices of
states $\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ with
$\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ directly, we define:
\begin{align}
  \label{eq:construction-rec}
  R^C_{ij} = \prod_{\messg} C_{\Rpure_i(\messg)\Rpure_j(\messg)}\,.
\end{align}

Now look at the sender. The probability that $\Spure_j$ is realized as
$\Spure_i$ under diffusion with $C$ is the product of the
probabilities, over all states $\mystate{k}$, that the message
$\Spure_i(\mystate{k})$, that $\Spure_i$ would produce at state
$\mystate{k}$ in the absence of noise, is produced by a noisy
realization of $\Spure_j$, which is the probability $\sum_{\state_l
  \in \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}$ that the state
$\mystate{k}$ is realized as a state $\mystate{l}$ which $\Spure_j$
would map unto $\Spure_i(\mystate{k})$. So, define:
\begin{align}
  \label{eq:construction-sen}
  Q^C_{ji} = \prod_{\state_k} \sum_{\state_l \in
    \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}\,.
\end{align}

For example, consider a signaling game with two states and two
messages. Let the confusion matrix be:
\begin{align*}
  C=
  \begin{pmatrix}
    .8 & .2 \\
    .2 & .8 
  \end{pmatrix}\,.
\end{align*}
The resulting mutation matrices are:
\begin{align*}
Q^C & = \bordermatrix{ ~ & 11 & 12 & 21 & 22 \cr
                      11 & 1 & 0 & 0 & 0 \cr
                      12 & .16 & .64 & .04 & .16 \cr
                      21 & .16 & .04 & .64 & .16 \cr
                      22 & 0 & 0 & 0 & 1 \cr}
&                    
  R^C & = \bordermatrix{ ~ & 11 & 12 & 21 & 22 \cr
                      11 & .64 & .16 & .16 & .04 \cr
                      12 & .16 & .64 & .04 & .16 \cr
                      21 & .16 & .04 & .64 & .16 \cr
                      22 & .04 & .12 & .12 & .64 \cr}\,.
\end{align*}
Here, a pair like $21$, for example, refers to a pure sender strategy
with $\Spure(\state_1)=\messg_2$ and $\Spure(\state_2) =
\messg_1$. Similarly for the receiver.

Based on the translations from confusion matrices $C$ to mutation
matrices $Q^C$ and $R^C$ in Equations~(\ref{eq:construction-rec}) and
(\ref{eq:construction-sen}), we can finally formulate the
desired correspondence result, that reveals the \rdd as a special case
of the \rmd in behavioral strategies where the only source of mutation is
imperfect discriminability of states.

\begin{theorem}
  \label{thm:Correspondence}
  (i) If $G(\Smixed) = \Sstrat$, then $G(\Mutate_{Q^C}(\Smixed)) =
  \Diff_C(\Sstrat)$. And, (ii) if $G(\Rmixed) = \Rstrat$, then
  $G(\Mutate_{R^C}(\Rmixed)) = \Diff_C(\Rstrat)$.
\end{theorem}

\noindent A proof is given in Appendix~\ref{sec:proofs}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Exploring the RDD}
\label{sec:exploring-rdd}

The previous section showed that the \rdd can be conceptualized as a
two-step procedure: we first calculate where the replicator dynamic
would take us, then apply diffusion. To understand the \rdd
better, in Section~\ref{sec:iterated-diffusion} we look closer at the novel
diffusion part, before exploring in Section~\ref{sec:simulations} how
diffusion and fitness-based selection interact.

\subsection{(Iterated) Diffusion}
\label{sec:iterated-diffusion}

Confusion of states should intuitively be a function of their
perceptual similarity. To make this concrete, let us assume that the
state space of the sim-max game consists of $n \ge 2$ states that are
equally spaced across the unit interval, including $0$ and $1$. The
distance $\card{\mystate{i} - \mystate{j}}$ is the objective, physical
similarity between two states $\mystate{i}$ and
$\mystate{j}$. Distance in physical space feeds into a perceptual
similarity function, as described by
\citet{Nosofsky1986:Attention-Simil}:
\begin{align*}
  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \impairment) =
      \begin{cases}
    1 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \impairment^2} \right ) & \textrm{otherwise,} \\
    \end{cases}
\end{align*}
where $\impairment >= 0$ is an imprecision or indiscriminability
parameter. When $\impairment=0$ agents perfectly discriminate between
states; when $\impairment \rightarrow \infty$ agents cannot
discriminate states at all. The top part of
Figure~\ref{fig:NosofskySim} gives an impression of
Nosofsky-similarity for different parameter values. Other
formalizations of perceptual similarity are possible, including ones
that allow for different discriminability in different areas of the
state space, but we stick with Nosofsky's similarity function for the
time being, because it is mathematically simple, yet an established
notion in mathematical psychology.

\begin{figure}
  \centering

  \includegraphics[width=0.85\textwidth]{plots/NosofskySim.pdf}

  \caption{Examples of Nosofsky-similarity (non-normalized) and
    derived probability of state confusion (normalized) for
    $\card{\States} = 50$ and different
    values of imprecision $\impairment$.}
  \label{fig:NosofskySim}
\end{figure}

We further assume that the probability of confusing any two states
$\mystate{i}$ and $\mystate{j}$ is proportional to their perceived
similarity and therefore obtained by normalization:
\begin{align*}
  C_{ij} = \frac{\similarity(\mystate{i}, \mystate{j} \myts ; \myts
  \impairment)} {\sum_j \similarity(\mystate{i}, \mystate{j} \myts ; \myts
  \impairment)} \,.
\end{align*}
Confusion of states is then also a function of imprecision parameter
$\impairment$ (see Figure~\ref{fig:NosofskySim}, bottom part). For $\impairment
= 0$ the confusion matrix has $C_{ii} = 1$; for $\impairment > 0$, $C$ is
positive, i.e., $C_{ij} >0$ for all $i$ and $j$; for $\impairment
\rightarrow \infty$ we find $C_{ij} = \nicefrac{1}{\card{\States}}$ for all
$i$ and $j$.

If states are confused on average with a probability proportional to
their similarity, repeated applications of diffusion of behavior as
defined in Equation~(\ref{eq:confusion-function}) causes behavioral
dispositions gradually to diffuse along a gradient of similarity of
states as well. Consequently, iterated diffusion leads to a smoothing
out and an eventual equalization of behavioral strategies, for both
the sender and the receiver. After $n$ steps of iterated diffusion an
initial sender strategy $\Sstrat$ will be $C^n \Sstrat$, and an
initial receiver strategy $\Rstrat$ will be $\Rstrat C^n$. If we
assume that $\impairment > 0$, $C$ is a positive row-stochastic matrix
(each state could in principle be confused as any other state, albeit
with possibly a very low probability). It then follows from the
Perron-Frobenius theorem that the limit $C^\infty = \lim_{n
  \rightarrow \infty} C^n$ exists (it is the Perron-projection) and
all of its rows are identical. That is why, in the limit of iterated
diffusion, $C^\infty S$ has identical rows (messages are sent with the
same probability in each state), and $R C^\infty$ likewise has
identical rows (every message is interpreted equally). All of this is
in line with the intuition that diffusion of behavior under confusion
of states iteratively equalizes behavior for similar states; if all
states can be confused for one another in principle, behavior smoothes
out entirely in the limit.

\subsection{Exploring the replicator diffusion dynamics}
\label{sec:simulations}

Now that we understand better what diffusion does, let us look at how
diffusion interacts with optimization of behavior, as described by the
replicator dynamics. The main question we should address is whether
diffusion and fitness-based replication reasonably interact, and if
so, whether the inclusion of diffusion has any noteworthy effects on
the evolving meaning of signals. Clearly, if $\impairment = 0$, the
\rdd reduces to the \rd. If $\impairment \rightarrow \infty$, the
diffusion component takes over and the \rdd reduces to the trivial
iterated diffusion process that we looked at in the previous
section. We will show presently that for reasonable in-between levels
of imprecision $\impairment$, the \rdd leads to communicatively
successful, yet vague signal meaning. Suitable levels of imprecision
can have further accelerating and unifying effects on meaning
evolution.

\subsubsection{Experimental set-up}

To explore the \rdd, we turn to numerical simulations. Let states be
evenly spaced elements of the unit interval that always include 0 and
1, and let priors be flat: $\Pr(\state) = \Pr(\state')$ for all
$\state, \state'$. As for utility functions, we take another
Nosofsky-style function:
\begin{align*}
  \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
      \begin{cases}
    1 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \toler^2} \right ) & \textrm{otherwise.} \\
    \end{cases}
\end{align*}
The tolerance parameter $\toler \ge 0$ models the amount of tolerable
pragmatic slack or communicative imprecision. This choice of utility
function is governed partly by convenience in parallel to the
similarity function, but also because it has, as we believe, the right
general properties for a communicative payoff function. Unlike
utilities that are, say, linearly or quadratically decreasing in
physical distance \citep[c.f.][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa},
utilities that are exponentially decreasing in negative quadratic
distance can model situations where a small amount of imprecision in
communication is tolerable, whereas similarly small differences in
intolerably far away interpretations matter very little, with a smooth
transition between these regimes
\citep[c.f.][]{OConnor2013:The-Evolution-o}.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_NS-90_tol-01_imp0_ind3098.pdf}
    \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0$}
    \label{fig:example_stratsA}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_NS-90_tol-01_imp01_ind3452.pdf}
    \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0.1$}
    \label{fig:example_stratsB}
  \end{subfigure}

  % \begin{subfigure}[]{0.45\textwidth}
  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp0_ind1001.pdf}
  %   \caption{$\ns = 10$, $\toler = 0.1$, $\impairment = 0$}
  %   \label{fig:example_stratsC}
  % \end{subfigure}
  % \hfill
  % \begin{subfigure}[]{0.45\textwidth}
  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp005_ind1201.pdf}
  %   \caption{$\ns = 10$, $\toler = 0.1$, $\impairment = 0.05$}
  %   \label{fig:example_stratsD}
  % \end{subfigure}

  \caption{Example strategies under \rdd at stopping time of our simulations.}
  \label{fig:example_strats}
\end{figure}

We ran 50 trials of the \rdd, starting with randomly sampled sender
and receiver strategies, for each triplet of independent parameter
values: $\ns = \card{T} \in \set{6, 10, 50, 90}$, $\impairment \in
\set{0, 0.05, 0.1, 0.2, 0.3}$, $\toler \in \set{0.05, 0.1, 0.2,
  0.3}$. Each trial ran for a maximum of 200 update steps of the
\rdd. This may not be enough to guarantee convergence to the eventual
attracting state, but we were interested especially in whether the
dynamics could lead to reasonable outcomes in reasonable time. A trial
was considered converged and stopped before the maximum of 200 rounds
if the total amount of change between strategies before and after the
current \rdd step was smaller than 0.001, i.e., if:
\begin{align*}
  \sum_{\state \in \States} \sum_{\messg \in \Messgs} \left|
    \rdd(\Sstrat)(\state,\messg) - \Sstrat(\state,\messg) \right| <
  0.001\,, & \text{\ \ and} \\
  \sum_{\messg \in \Messgs} \sum_{\state \in \States} \left|
    \rdd(\Rstrat)(\messg,\state) - \Rstrat(\messg,\state) \right| <
  0.001 & \,.
\end{align*}

Representative examples for resulting strategy pairs at stopping time
are given in
Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA}
shows a strategy pair at stopping time with 90 states, tolerance
$\toler = 0.1$ and imprecision $\impairment = 0$. Zero imprecision
means that the trial was effectively an application of the plain
\rd. Noteworthily, the given sender strategy approximates a pure
sender strategy that crisply partitions the state space into
non-convex sets. The irregular shape of the receiver strategy shows
clearly that the pictured strategy pair has not yet reached a
dynamically stable state under the \rd. Indeed, the trial did not
converge in our technical sense, but was stopped after the maximum of
200 rounds. In contrast, the outcome of a trial with identical
parameters, except with imprecision $\impairment = 0.1$, which is
shown in Figure~\ref{fig:example_stratsB}, had converged (in our
technical sense) after 99 rounds of iterated \rdd. The sender strategy
shows a smooth blending from one ``category'' to the other, and the
receiver's interpretations are rather extended curves, peaking at a
central point in the relevant ``categories.''

These examples already show two interesting things. Firstly, inclusion
of imprecision can lead to seemingly well-behaved, yet vague
strategies in the sense that we are after. The sender strategy in
Figure~\ref{fig:example_stratsB} identifies clear positive and clear
negative cases for each signal, with a smooth transition
in-between. The receiver's interpretations of signals can be seen as
smoothed-out prototype regions. Secondly, (sender) strategies can
approach non-convex pure strategies under the replicator dynamic. We
see this in our limited-time simulations, but this also holds, for
some types of utility functions, for the limiting case. This was first
observed by Elliott Wagner in unpublished work
\citep[c.f.][]{OConnor2013:Evolving-Percep}. A full analysis of the
dynamics of sim-max games is beyond the scope of this paper, but we
will see shortly that diffusion from confusability of states clearly
prevents evolutionary paths that meander for a long time in the
vicinity of non-convex strategies.

\subsubsection{Dependent measures}

To further explore our simulation results, we defined a small number
of metrics that aim to capture numerically how vague, communicatively
efficient and generally well-structured the resulting strategy pairs
were. \emph{Entropy} and \emph{informativity} capture (slightly
different aspects of) the amount of systematicity or regularity in
signal use. \emph{Voronoiness} and \emph{convexity} are, respectively,
gradient and categorical measures for the coherence of signal-induced
categories. \emph{Expected utility} measures the communicative
efficiency of evolved strategy pairs.

\paragraph{Entropy.} This classic information-theoretic notion
captures the amount of uncertainty in a probability
distribution. Roughly put, entropy of a signaling strategy captures
inverse distance from a pure strategy. The usual definition of entropy
applies directly to mixed strategies:
\begin{align*}
  E(\Smixed) & = \sum_{\Spure \in \Messgs^\States} \Smixed(\Spure) \cdot
  \log(\Smixed(\Spure)) & 
  E(\Rmixed) & = \sum_{\Rpure \in \States^\Messgs} \Rmixed(\Rpure) \cdot \log(\Rmixed(\Rpure)) \,.
\end{align*} 
These measures are computationally expensive to calculate, since the
size of the domain over which the sum is computed grows exponentially
with the number of choice points. But provably equivalent metrics for behavioral
strategies are ready to hand:
\begin{align*}
  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs}
  \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar
  \state)) 
\end{align*} 
\begin{align*}
  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\state \in \States}
  \Rstrat(\state \probbar \messg) \cdot \log(\Rstrat(\state \probbar
  \messg)) \,. 
\end{align*}
Values obtained by these definitions are lower bounded by $0$ and
upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States|
\cdot \log(|\Messgs|)$ and $\log(|\States^\Messgs|) = |\Messgs| \cdot
\log(|\States|)$. We work with normalized values.

The sender strategies in Figures~\ref{fig:example_stratsA} and
\ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.21$,
respectively. The receiver strategies have respective entropies $0.43$
and $0.84$. In general, we expect that vague languages will have
higher entropy than crisp ones and that increasing imprecision will
also increase entropy, all else being equal.

\paragraph{Informativity.} The quantity of information in a signal is
an old notion that also goes back to the start of information theory.
\citet[ch.~3]{Skyrms2010:Signals} discusses its use in the context of
signaling games. Intuitively speaking, informativity in Skyrms' sense
measures how revealing the sender's signal use is about which state is
actual. It therefore measures, in our given set-up with flat priors, a
sense of well-behavedness or regularity of a sender strategy.

The main idea behind Skyrms' notion is that we can quantify the amount
of information about a state $\state$ in a message $\messg$ via the
relation between the probability of the state given the message
$\Pr(\state \,|\, \messg)$ and the unconditional probability of the
state $\Pr(\state)$.  Following Bayes' theorem, we can express
$\Pr(\state \,|\, \messg)$ as $\frac{\Pr(\state) \cdot \Pr(\messg
  \,|\, \state)}{\Pr(\messg)}$.  We then have $\Pr(\messg \,|\,
\state) = \Sstrat(\state,\messg)$ and $\Pr(\messg) =
\sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot
\Sstrat(\state^\prime,\messg)$.  Finally, we equate sender
infomativity with the average overall information about states in each
signal.  Following \citet[p.~36]{Skyrms2010:Signals}, and the
considerations above, the informativity about states conveyed by a
sender strategy is:
\begin{align*}
  I(\Sstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs}
  \sum_{\state \in \States} \frac{\Pr(\state) \cdot
    \Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States}
    \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \cdot \log
  \left(\frac{\Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States}
      \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \right)
  \,. 
\end{align*}
This measure is bounded between $0$ and $1$. The sender strategies in
Figures~\ref{fig:example_stratsA} and \ref{fig:example_stratsB} have
informativity $0.69$ and $0.54$, respectively. In general, we expect
that vague languages will have lower informativity than crisp
ones and that increasing imprecision will decrease informativity.

\citet[p.~39]{Skyrms2010:Signals} also offers a notion of
informativity about the acts performed by the receiver, but it turns
out that this is perfectly correlated with information about the
states in the results of our simulations. We therefore only look at
informativity about states.

% Conversely, we can quantify the amount of information about an act in a message.
% We equate receiver informativity with the average overall information about acts in each signal.
% Based on Skyrms' definition~\citet[p.~39]{Skyrms2010:Signals}, we define receiver informativity as follows:
% \begin{align*}
%   I(\Sstrat,\Rstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs} \sum_{\act \in \Acts} \Rstrat(\messg,\act) \cdot \log \left(\frac{\Rstrat(\messg,\act)}{\sum_{\state \in \States} \Pr(\state) \cdot \sum_{\messg^\prime \in \Messgs} \Sstrat(\state,\messg^\prime) \cdot \Rstrat(\messg^\prime,\act)} \right) \,.
% \end{align*}

\paragraph{Voronoiness.}
This metric aims to quantify how close a strategy is to being a part
of a Voronoi tessallation of the state space. Bearing in mind that
this metric is specifically geared towards the sim-max games in our
set-up, we define Voronoiness as follows. Let $\pi(\Rstrat, \messg) =
\argmax_{\state \in \States}\Rstrat(\messg,\state)$ be the prototype
of a message $\messg \in \Messgs$, then
\begin{align*}
  V(\Sstrat, \Rstrat) & = \sum_{\state \in \States} \sum_{\messg_1 \in
    \Messgs} \sum_{\messg_2 \in \Messgs} F(\Sstrat, \Rstrat, \state,
  \messg_1, \messg_2) \text{\ \ \ with } \\
  F(\Sstrat, \Rstrat, \state,
  \messg_1, \messg_2) & =
  \begin{cases}
    1 & \text{if } \similarity(\state, \pi(\Rstrat, \messg_1)) >
    \similarity(\state, \pi(\Rstrat, \messg_2)) \bicond \\
    &  \ \ \ \ \ \Sstrat(\state, \messg_1) >
    \Sstrat(\state, \messg_2) \\
    0 & \text{otherwise.}
  \end{cases}
\end{align*}
is the Voronoiness of the sender strategy $\Sstrat$ given the receiver
strategy $\Rstrat$ and
\begin{align*}
  V(\Rstrat) & = \sum_{\messg \in
    \Messgs} \sum_{\state_1 \in \States}  \sum_{\state_2 \in \States}
  F(\Rstrat, \messg, \state_1, \state_2) \text{\ \ \ with } \\
  F(\Rstrat, \messg, \state_1, \state_2) & =
  \begin{cases}
    1 & \text{if } \similarity(\state_1, \pi(\Rstrat,
  \messg)) > \similarity(\state_2, \pi(\Rstrat, \messg)) \\
  & \ \ \ \ \ \bicond \Rstrat(\messg,
  \state_1) > \Rstrat(\messg, \state_2) \\
    0 & \text{otherwise.}
  \end{cases}
\end{align*}
is the Voronoiness of the receiver strategy $\Rstrat$.  The metrics
are lower bound by $0$ and upper bound by, respectively, $|\States|
\times |\Messgs|^2$ and $|\Messgs| \times |\States|^2$, thus we will
normalize them by dividing by these values.

Despite being vague, a language can have maximal Voronoiness according
to this gradient measure. For instance, the sender strategies in
Figures~\ref{fig:example_stratsA} and \ref{fig:example_stratsB} have
Voronoiness $0.91$ and $1$, respectively. The receiver strategies have
Voronoiness $0.78$ and $1$, respectively. We do not \emph{prima facie}
expect a conceptual connection between vagueness and Voronoiness, and
so do not predict a clear effect of imprecision on Voronoiness either.

\paragraph{Convexity.} At least for sender strategies, which converge
faster, it also makes sense to define a categorical measure of
convexity, that compensates for potential vagueness. To determine
whether a sender strategy $\Sstrat$ is convex despite possibly being
vague, we look at the derived pure strategy $\Spure$ for which
$\Spure(\state) = \argmax_{\messg' \in \Messgs}
\Sstrat(\state,\messg')$. If that $\Spure$ is convex, we also count
$\Sstrat$ as convex. The sender strategy in
Figure~\ref{fig:example_stratsA} is not convex, while the one in
Figure~\ref{fig:example_stratsB} is. Again, we do not expect a
conceptual relation between vagueness and imprecision, on the one
hand, and convexity on the other.

\paragraph{Expected utility.} We also record the expected utility of
a strategy pair:
\begin{align*}
  \EU(\Sstrat,\Rstrat \myts ; \myts \toler) = \sum_{\state \in
    \States} \sum_{\messg \in \Messgs} \sum_{\state' \in \States}
  \Pr(\state) \cdot \Sstrat(\state, \messg) \cdot \Rstrat(\messg,
  \state') \cdot \Util(\state,\state' \myts ; \myts \toler)\,.
\end{align*}
To make direct comparisons across different parameter settings, we
normalize expected utility by the maximal amount of expected utility
obtainable in the relevant game.

The strategy pair in Figure~\ref{fig:example_stratsA} has a normalized
expected utility of $0.992$, the pair in
Figure~\ref{fig:example_stratsB} has $0.901$. Generally, vagueness and
imprecision can be expected to decrease expected utility
\citep[c.f.][]{Lipman2009:Why-is-Language}. The crucial question is whether
communicative success drops unacceptably fast with moderate levels of
vagueness and imprecision.

\subsubsection{Results}

\begin{figure}
  \centering

  \includegraphics[width=\textwidth]{plots/MeanMetrics.pdf}

  \caption{Means of recorded metrics for each triple of independent
    variables. Notice that the plots also show 95\% confidence
    intervals, but these are most often invisible, owing to the
    uniformity of outcomes across trials.}
  \label{fig:MeanMetrics}
\end{figure}

Figure~\ref{fig:MeanMetrics} shows the means of the recorded metrics,
together with their estimated 95\% confidence intervals, for each
triple of independent variables. Most trends are quite clear, and not
unexpected. Entropy of sender and receiver strategies are increasing
with imprecision. Entropy of sender strategies seems independent of
tolerance levels and the number of states, while there seems to be an
effect of both parameters on the entropy of the receiver
strategy. Sender informativity and expected utility are decreasing
under growing imprecision. There seem to be mild effects of the number
of states in the game, but hardly any in terms of pragmatic
tolerance. What was not expected \emph{a priori} was that Voronoiness
of both sender and receiver strategies appear to be non-decreasing
under increasing impairment, with the only exception being the 6 states
games and the transition from $\impairment = 0$ to $\impairment =
0.05$. Expected utility of evolved languages mostly decreases in
increasing imprecision, with the sole exception of a rise for $\toler
= 0.2$ and $n_s = 6$ at $\impairment = 0.1$. Still, normalized expected
utility of evolved strategy pairs under mild values of imprecision are
not necessarily much worse than their respective perfect precision
\rd-analogues. This is especially so for lower numbers of states
(which increase the chance of non-convergence and non-convexity; see
Figure~\ref{fig:CategoricalMeasures} and the discussion below).

Taken together, these findings suggest the following effect of
confusability of states and diffusion of behavior. Mild forms of
perceptual imprecision lead to slightly less communicative efficiency,
more vagueness and more regular, well-behaved languages. This
impression is also supported by the proportion of converged and convex
trials, pictured in
Figure~\ref{fig:CategoricalMeasures}. Non-converged and non-convex
sender strategies showed for several values of parameters in the
absence of impairment. Increasing impairment always assured
convergence and convexity: diffusion speeds up convergence to a convex
and regular language.

\begin{figure}
  \centering

  \includegraphics[width=0.8\textwidth]{plots/CategoricalMeasures.pdf}

  \caption{Proportion of converged and convex trials.}
  \label{fig:CategoricalMeasures}
\end{figure}

Diffusion also has another interesting regularizing effect on the
evolution of signaling. If we look at the confidence intervals in
Figure~\ref{fig:MeanMetrics}, we see that there is very little
variation in the recorded metrics for evolved strategies, at least for
higher values of impairment. On closer inspection, it turns out that
variability in low-impairment conditions is not only due to
non-convergence or non-convexity. Figure~\ref{fig:MoreExample_strats}
gives two more examples of strategy pairs at stopping time. Both are
obtained for the same triple of parameters, both converged and are
convex. However, they are not equally efficient. In fact, the pair in
Figure~\ref{fig:example_stratsC} has a normalized expected utility of
$0.99$ while the one in Figure~\ref{fig:example_stratsD} only has
$0.89$.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_NS-6_tol-03_imp005_ind41.pdf}
    \caption{$\ns = 6$, $\toler = 0.2$, $\impairment = 0$}
    \label{fig:example_stratsC}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_NS-6_tol-03_imp005_ind23.pdf}
    \caption{$\ns = 6$, $\toler = 0.2$, $\impairment = 0$}
    \label{fig:example_stratsD}
  \end{subfigure}

  \caption{More example strategies under \rdd at stopping time of our simulations.}
  \label{fig:MoreExample_strats}
\end{figure}

Interestingly, this type of variability in evolutionary outcomes is
also weeded out by mild increases in impairment. To see this, we
calculated the average distance between evolved sender strategies
within each group of trials that had identical independent
parameter values. We determined the distance between sender strategies
$\Sstrat$ and $\Sstrat'$ as the average Hellinger distance between
probability distributions $\Sstrat(\state)$ and $\Sstrat'(\state)$
over messages at each choice point $\state$:
\begin{align*}
  \text{HD}(\Sstrat,\Sstrat') & = \frac{1}{{\card{\States} \cdot
     \sqrt{2}}} \cdot  \sum_{\state \in \States} 
 \sqrt{\sum_{\messg \in  \Messgs}
         \left ( \sqrt{\Sstrat(\state,\messg)} -
         \sqrt{\Sstrat'(\state,\messg)} \right )^2 }\,.
\end{align*}
To compensate for the arbitrariness of message use, we set the
distance between strategies $\Sstrat$ and $\Sstrat'$ to be the maximum
of $\text{HD}(\Sstrat,\Sstrat')$ and $\text{HD}(\Sstrat^*,\Sstrat')$
where $\Sstrat^*$ is $\Sstrat$ with reversed message indices. The
\emph{within group distances}, i.e., the average distances between all
sender strategies in each parameter group, are plotted in
Figure~\ref{fig:AverageEUinGroups}. This clearly shows that with
increasing imprecision, the sender strategies evolving under \rdd were
exactly alike (modulo swapping of messages). In other words,
perceptual imprecision unifies evolutionary outcomes, and in fact
guarantees sender strategies that are not only convex, but also
maximally efficient in that they induce a vague category split exactly
in the middle of the unit interval.

\begin{figure}
  \centering

    \includegraphics[width=0.8\textwidth]{plots/WithinGroupMeasures.pdf}

    \caption{For each group, i.e., triplet of independent variables,
      the plots show (i) within group distances and (ii) within group
      expected utilities, as defined in the main text.}
  \label{fig:AverageEUinGroups}
\end{figure}

This unifying property of perceptual imprecision could be considered
an evolutionary beneficial side-effect. Higher imprecision can lead to
higher \emph{within group expected utility}, i.e., the average
expected utility that each evolved language (i.e., pair of sender and
receiver strategy at stopping time) scored when playing against an
arbitrary other language of the same parameter group (see
Figure~\ref{fig:AverageEUinGroups}). What this means is that, while
imprecision might decrease the communicative efficiency of individual
languages, it increases the conceptual coherence and communicative
success between independently evolving strategies. It is almost as if
mere confusability of states imposes a regularity constraint on
evolving categories.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}
\label{sec:discussion}

\subsection{Model interpretation}
\label{sec:model-interpretation}

We mentioned in passing in Section~\ref{sec:repl-diff-dynam-1} that
adding diffusion to other discrete-time evolutionary dynamics that
operate on behavioral strategies is entirely straightforward. We
could, for instance, easily diffuse the outcome of a best-response
dynamic at each time step. That would make sense if we thought of
agents as prone to confusing similar states, but otherwise rationally
optimizing behavior. The reason that we chose the \rd to combine with
diffusion is twofold. A minor reason is that it makes for a
conceptually interesting link with the \rmd. A more important reason
is that the \rd is especially versatile and non-committal about what
the exact process of adaption is that is being modelled. Generally,
the \rd could be a biological or a cultural dynamic, and so could the
\rdd.

Under a biological interpretation of the \rd, signaling strategies
are innate behavioral dispositions of organisms, steps in the
evolutionary process are successive generations, and selection
captures the reproductive advantage of fitter individuals. This
interpretation, strictly speaking, requires formalization in terms of
mixed strategies via the \rmd, and possibly also a symmetrizing of the
game, so that every agent is assumed to have a unique sender and
receiver role at the same time, the pair of which is bequeathed onto
the next generation.

Diffusion, in this context, could be either of two things. It could be
differential mutation probabilities in line with other interpretations
of mutation in the \rmd
\citep[e.g.][]{NowakKomarova2001:Evolution-of-Un,KomarovaNiyogi2001:The-Evolutionar}:
pure strategies are not faithfully reproduced, e.g. by imperfections
of learning or underdetermination of input, and mistakes in inheriting
pure strategies are more likely if they result from the confusion of
similar states. Another possible biological interpretation of the
\rdd, is that inheritance is faithful, but strategies are noisily
realized. Strategies are not necessarily selected for what they are,
but rather for how they are realized, once noise is factored in.

As a cultural dynamic, the \rd can also be interpreted as a high-level
description of the likely development of other behavioral adaptation
processes, like differential imitation \citep[see][for various
derivations of the \rd]{Sandholm2013:Population-Game}. Under this
interpretation, what is subject to evolutionary optimization is
behavior, not genetic make-up. Fitness captures the success of
behavioral patterns.  Differential reproduction represents the
tendency of more successful ways of behavior to be more likely
realized, be it through imitation or some solitary learning or
optimization mechanism.

Under this non-biological interpretation, we have again two options of
picturing what diffusion is, similar to the biological cases
before. One possibility is that diffusion is noise in the adoption of
strategies, e.g. by conditional imitation, of strategies by other
signalers. Another possibility is, again, that diffusion is noise on
the realization of strategies: while behavior is optimized to be
efficient (be it due to learning, introspection or imitation),
realization of strategies is bound to be noisy due to confusability of
similar states.

We believe that all of the four mentioned interpretations are, on
first approximations, feasible conceptualizations of the \rdd, and
that it is a good thing to know of a working account of vague
signaling that sketches where fitness-based selection under
state-confusability will take us, abstracting away from the details of
actually playing the game, inheriting, imitating or otherwise
optimizing behavior in whatever particular way. It is a good thing to
know this on the macro-level, especially since there are also
micro-level accounts that nicely complement the picture. We turn to
one such next.

\subsection{Relation with previous accounts}
\label{sec:relat-with-prev}

\paragraph{Generalized reinforcement.}
% Repeated below...
%\citet{OConnor2013:The-Evolution-o} introduced a generalization of
%Herrnstein reinforcement learning for sim-max games and showed that
%this not only leads to vague signaling patterns, but can speed-up
%evolution of efficient signaling strategies, especially in games with
%many states.

Under \emph{Herrnstein reinforcement learning}, sender and receiver
adjust their dispositions to act after each round of play, in such a
way that the actual (non-negative) payoff gained in the previous
interaction is added to the non-normalized propensities for repeating
the same behavior. For sim-max games, this means that, when the sender
chose $\messg$ in state $\state$, and this resulted in some
non-negative payoff, the probability that the sender chooses $\messg$
again in $\state$ is increased, but nothing else changes. In
particular, the sender's behavior in other choice points does not
change. \emph{Generalized reinforcement learning} is different. When
the use of $\messg$ in $\state$ gave positive payoff, then not only
will its future use probability be promoted at $\state$ but also at
other states, proportional to how similar these are to
$\state$. Similar amendments take care of the way that the receiver
updates his choice dispositions.

\citet{OConnor2013:The-Evolution-o} shows that this extension not only
leads to vague signaling of the appropriate kind, but also speeds up
learning in such a way that, especially for sim-max games with higher
numbers of states, higher levels of communicative success are reached
in shorter learning periods. Technically, this result is partly due to
the fact that signalers make bigger changes to their behavioral
strategies after each round of play under generalized reinforcement
than under the Herrnstein variety. But that only explains the speed of
adaptation, not necessarily that generalization also leads to
regularity and communicative efficiency, but it does.

Diffusion of strategies in the \rdd can also be conceived of as a form
of generalization, and works in large part quite analogous to stimulus
generalization in \citeauthor{OConnor2013:The-Evolution-o}'s
approach. But there are still
differences. \citeauthor{OConnor2013:The-Evolution-o} showed that
generalized reinforcement learning can speed up the development of
efficient signaling, especially for higher numbers of
states. Complementing this, we showed that diffusion regularizes
evolutionary outcomes and prevents meandering around sub-optimal and
non-convex strategies, also for low numbers of states.

Conceptually, the \rdd is a more abstract framework than generalized
reinforcement learning. The latter is foremost motivated as a learning
dynamic that has two players adapt their individual strategies after
each concrete round of play. In contrast, the \rdd describes a more
abstract, average dynamical change in behavioral
dispositions. Although the behavior of (some forms of) reinforcement
learning mirror those of the \rd (at some stage in time)
(\cite{BorgersSarin997:Learning-Throug,HopkinsPosch2005:Attainability-o,Beggs2005:On-the-Converge}),
this does not mean that \emph{generalized} reinforcement learning is
also necessarily a plausible high-level description of, say,
generalized learning in a population of agents. Seen in this light,
generalized reinforcement learning and the \rdd nicely complement each
other, as similarly-minded accounts operating at different levels of
abstraction.

\paragraph{Quantal response equilibria.}
\citet{FrankeJager2010:Vagueness-Signa} suggested a number of ways in
which information-processing limitations could lead to vague
strategies. The model that is most clearly related to the present
approach uses the notion of a \emph{quantal response}, also known as a
\emph{soft-max function}
\citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,GoereeHolt2008:Quantal-Respons}. A
quantal response function is a paramterized generalization of the
classic best response function. For example, if $U \mycolon \Acts
\rightarrow \mathds{R}$ is the measure of expected utility over
choices $\Acts$ of an agent, then a best response function would have
the agent choose $\act$ only if $U(\act) = \max_{\act' \in \Acts}
U(a)$. A quantal response function rather assumes that agents would
choose $\act$ with a probability proportional to $\expo(\lambda \cdot
U(\act))$, where $\lambda$ is a rationality parameter. If $\lambda
\rightarrow \infty$ we retrieve the behavior of the best response
function, but if it is positive but finite, any choice $\act$ will
receive a positive probability, but acts with higher expected utility
will be more likely. The underlying motivation is the assumption that
there is noise in the computation of expected utilities and/or in
maximization of expected utilities. Consequently, choices with almost
equal expected utilities will be chosen with almost equal probability
(for moderate values of $\lambda$).

\citet{FrankeJager2010:Vagueness-Signa} show that quantal response
equilibria of sim-max games, i.e., pairs of sender and receiver
strategies such that the sender strategy is the quantal response to
the expected utilities under the receiver strategy and vice versa, can
show the desired marks of vague
signaling. Figure~\ref{fig:exampleQRE_stratsA} gives an example of a
quantal response equilibrium for a sim-max game, as used in our set-up
but with higher tolerance $\toler = 0.5$. Sender and receiver
strategies look very much like what evolves under \rdd with modest
values of perceptual imprecision.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/exampleStratQRE_tolerance05.pdf}
    \caption{$\ns = 90$, $\lambda = 15$, $\toler = 0.5$}
    \label{fig:exampleQRE_stratsA}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/exampleStratQRE_tolerance005.pdf}
    \caption{$\ns = 90$, $\lambda = 15$, $\toler = 0.05$}
    \label{fig:exampleQRE_stratsB}
  \end{subfigure}

  \caption{Examples of vague quantal response equilibria.}
  \label{fig:exampleQREs}
\end{figure}

But not all quantal response equilibria are equally plausible, and the
ones that are not suggest that it is less natural to think of
vagueness as arising from errors in the computation or maximization of
expected utility than that it arises from confusion of similar
states. To see what the problem is, we can look at cases like given in
Figure~\ref{fig:exampleQRE_stratsB}, which is the quantal response
equilibrium for a game with lower tolerance $\toler = 0.05$. Unlike
what evolves under \rdd in this case, sender strategies have vague
boundaries also towards the end of the unit intervals. Technically,
this is because quantal responses equalize message use far away from
the ``prototypical'' interpretation, not just in-between categories,
so to speak. This, in turn, is because quantal responses introduce
noise into the decision making at the level of computing or maximizing
expected utility of choices.

That this is conceptually odd shows even more clearly in a case where
the state space is intuitively unbounded, as for instance for the
property ``tall''. If the usual interpretation of a ``tall man'' peaks
at around, say, 195cm then when meeting a giant of $n$ meters, senders
would, according to the quantal response approach, be ever more
inclined to describe the giant indifferently as either ``tall'' or
``short'' the larger $n$ gets. This is because, as the distance from
the prototype increases for larger $n$, the difference between the
expected utilities of saying ``short'' or ``tall'' will converge to
zero. 
%Whence that the quantal response approach would predict that
%senders would grow indifferent between choice of antonyms as $n$
%grows, which seems weird.

Admittedly, this argument hinges on the choice of utility
function. Still, to the extent that the chosen lower bounded utility
functions are reasonable---and we think they are very reasonable---
the case suggests that quantal responses are not a good model,
intuitively speaking, for why linguistic categories are
vague. Vagueness is more plausibly an effect of perceptual confusion
of similar states than of computational errors in maximizing expected
utility.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{sec:conclusions}

Vagueness is a pervasive but seemingly non-disruptive feature of
natural communication and classification systems. From an evolutionary
point of view, the challenge arises to explain how vagueness can
persist under selective pressure for precision. This is foremost a
technical challenge, probing for the possibility of integrating into a
consistent model forces that lead to vagueness and forces that lead
to efficient information transfer. This paper proposed one such model
in the context of sim-max games and explored some of its consequences.

We introduced the replicator diffusion dynamic as a special case of
the replicator mutator dynamic. The \rdd incorporates a special kind
of stochastic mutation based on differential confusability of similar
states. Based on data from numerical simulations, we demonstrated that
the inclusion of mild levels of stochastic confusability of states
does not undermine the possibility of evolving communicatively
successful signaling strategies at all. On the contrary, strategies
that evolved under mild diffusion induce a highly regular and
systematic category structure that shows the signs of vagueness as
desired. There might even be a higher-order benefit to the presence of
imprecision, in that it can accelerate convergence to optimal
categorization, by preventing evolutionary paths to stay near
inefficient non-convex strategies for a long time. Diffusion thus also
unifies and regularizes evolutionary outcomes in such a way that
sub-optimal categorization is avoided.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Proof of Theorem}
\label{sec:proofs}

\begin{proof}[Part (i).]
  Fix $\Smixed$ and $\Sstrat = G(\Smixed)$. Look first at the rhs of
  the consequent:
  \begin{align*}
    \Diff_C(\Sstrat)(\messg_y \probbar \state_x) & =  \sum_{\state_l} C_{xl}
    \cdot \Sstrat(\messg_y \probbar \state_l) && \text{(by Equation~(\ref{eq:confusion-function}))} \\
    & =  \sum_{\state_l} C_{xl}
    \cdot  \sum_{\Spure_i(\state_l) = \messg_y} \Smixed_i && \text{(by Equation~(\ref{eq:CorrespondenceBehavioralMixed}))} \\
    & = \sum_{\state_l}
    \sum_{\Spure_i(\state_l) = \messg_y} \Smixed_i \cdot C_{xl}\,.
  \end{align*}
  Next consider the lhs of the consequent:
  \begin{align*}
    G(\Mutate_{Q^C}(\Smixed))(\messg_y \probbar \state_x) & =
    \sum_{\Spure_i(\state_x)=\messg_y} \Mutate_{Q^C}(\Smixed_i) &&
    \text{(by Equation~(\ref{eq:CorrespondenceBehavioralMixed}))} \\
    & = \sum_{\Spure_i(\state_x)=\messg_y} \sum_{\Spure_j}
    \Smixed_j \cdot Q^C_{ji} &&
    \text{(by Equation~(\ref{eq:Mutation}))} \\
    & = \sum_{\Spure_i(\state_x)=\messg_y} \sum_{\Spure_j}
    \Smixed_j \cdot \prod_{\state_l} \sum_{\state_m \in
      \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm} &&
    \text{(by Equation~(\ref{eq:construction-sen}))} \\
    & = \sum_{\Spure_j} \Smixed_j \cdot
    \sum_{\Spure_i(\state_x)=\messg_y} \prod_{\state_l}
    \sum_{\state_m \in \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm}
  \end{align*}
  To simplify this further we look at a fixed $\Spure_j$ and consider
  the term: 
  \begin{align}
    \label{eq:term}
    \sum_{\Spure_i(\state_x)=\messg_y} \prod_{\state_l} \sum_{\state_m
      \in \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm}\,.
  \end{align}
  Let $Y$ be the row-stochastic matrix with $Y_{kl} = \sum_{\state_m
    \in \Spure_j^{-1}(\messg_l)} C_{km}$. Every pure sender strategy
  maps each state $\state_k$ onto exactly one $Y_{kl}$. If we quantify
  over all pure strategies, we essentially look at each such
  mapping. Term (\ref{eq:term}) above sums over all pure strategies
  that map $\state_k$ onto $\messg_y$. The above term then sums over
  all products whose factors are tuples in $\times_{k>2} \set{y
    \setbar \exists l \mycolon y = Y_{kl}}$. So term (\ref{eq:term})
  expands to (where $e = \card{\States}$ and $d=\card{\Messgs}$):
  \begin{align*}
    & (Y_{11} \cdot Y_{21} \cdot Y_{31} \cdot \ldots \cdot Y_{e1}) +
    (Y_{11} \cdot Y_{21} \cdot Y_{31} \cdot \ldots \cdot Y_{e2}) + 
    \dots \\
    & + (Y_{11} \cdot Y_{2d} \cdot
    Y_{3d} \cdot \ldots \cdot
    Y_{ed}) 
  \end{align*}
  But since $Y$ is a row-stochastic matrix, this simplifies to
  $Y_{xy}$. Continuing the derivation with this:
  \begin{align*}
    G(\Mutate_{Q^C}(\Smixed))(\messg_y \probbar \state_x) 
    & = \sum_{\Spure_j} \Smixed_j \cdot
    \sum_{\state_l \in \Spure_j^{-1}(\messg_y)} C_{xl} \\
    & = \sum_{\Spure_j} \sum_{\state_l \in \Spure_j^{-1}(\messg_y)} \Smixed_j \cdot
     C_{xl} \\
     & = \sum_{\state_l}
    \sum_{\Spure_i(\state_l) = \messg_y} \Smixed_i \cdot C_{xl}\,.
  \end{align*}

\end{proof}

\begin{proof}[Part (ii).]
  Fix $\Rmixed$ and $\Rstrat = G(\Rmixed)$. The rhs of the consequent
  expands to:
  \begin{align*}
    \Diff_C(\Rstrat)(\state_x \probbar \messg_y) & = \sum_{\state_j}
    \sum_{i \in \set{k \setbar \Rpure_k(\messg_y) = \state_x}}
    \Rmixed_i \cdot C_{jx} && \text{(by Equation~(\ref{eq:confusion-function}))} \\
    & = \sum_{\Rpure_i} \sum_{j \in \set{k \setbar \Rpure_i(\messg_y) = \state_j}}
    \Rmixed_i \cdot C_{jx} \\
    & = \sum_{\Rpure_i} \Rmixed_i \cdot C_{\Rpure_i(\messg_y)x} \,.
  \end{align*}
  The rhs expands to (by Equations~(\ref{eq:CorrespondenceBehavioralMixed}),
  (\ref{eq:Mutation}) and (\ref{eq:construction-rec})):
  \begin{align*}
    G(\Mutate_{R^C}(\Rmixed))(\state_x \probbar \messg_y) & = \sum_{\Rpure_i}
    \Rmixed_i \cdot \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) =
        \state_x}} 
    \prod_{\messg} C_{\Rpure_i(\messg)\Rpure_j(\messg)} \,.
  \end{align*}
  Without loss of generality, assume that $x=y=1$, and fix
  $\card{M}=d$ and let $e$ be the number of pure receiver
  strategies. Then the last term can be rewritten as:
  \begin{align*}
    & = \sum_{i}
    \Rmixed_i \cdot ( C_{\Rpure_i(\messg_1)1} \cdot
      C_{\Rpure_i(\messg_2)\Rpure_1(\messg_2)} \cdot \ldots \cdot
      C_{\Rpure_i(\messg_d)\Rpure_1(\messg_d)} + \ldots  \\
      & \textcolor{white}{ = \sum_{i}
    \Rmixed_i  (}  +  C_{\Rpure_i(\messg_1)1} \cdot
      C_{\Rpure_i(\messg_2)\Rpure_2(\messg_2)} \cdot \ldots \cdot
      C_{\Rpure_i(\messg_d)\Rpure_2(\messg_d)} + \ldots  \\
      & \textcolor{white}{ = \sum_{i}
    \Rmixed_i  (}  + C_{\Rpure_i(\messg_1)1} \cdot
      C_{\Rpure_i(\messg_2)\Rpure_e(\messg_2)} \cdot \ldots \cdot
      C_{\Rpure_i(\messg_d)\Rpure_e(\messg_d)}) ) \,.
  \end{align*}
  For every messages $\messg_l$, $C_{\Rpure_i(\messg_l)}$ is a stochastic
  vector. For $l>1$, all elements of these vectors appear equally
  often. But that means that these cancel out. What remains is:
  \begin{align*}
    & = \sum_{\Rpure_i} \Rmixed_i \cdot C_{\Rpure_i(\messg_y)x} \,.
  \end{align*}
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\printbibliography[heading=bibintoc]

\end{document}