### 2014-11-05 snippets/data_exploration/data_exploration_tmp.Rmd
results graphs merged

Data Exploration
================

``` {r warnings=FALSE, message=FALSE, echo=FALSE}

require('ellipse')
require('ggplot2')
require('reshape2')
require('plyr')
source('~/Desktop/data/svn/programming/R/helpers/helpers.R')
source('~/Desktop/data/svn/vagueness-games/data-analysis.R')
source('~/Desktop/data/svn/vagueness-games/helper_functions.R')

```

plot all the relevant variables against each other

``` {r warning = FALSE, fig.width=12, fig.height=12}

pairs(~ Speaker.entropy + Hearer.entropy + Speaker.Voronoiness + Hearer.Voronoiness +
        Speaker.informativity + Hearer.informativity + Expected.utility +
        Speaker.Convex.Cat, data=ds)
```

some properties are very clearly positively correlated:

* speaker informativity, hearer informativity & expected utility
* impairment & speaker entropy

these two "clusters" are also strongly negatively correlated

``` {r warning = FALSE, fig.width=7, fig.height=7}

dss = ds[c("Impairment", "Speaker.informativity", "Hearer.informativity", "Expected.utility",
           "Speaker.entropy")]

ctab <- cor(dss)

plotcorr(ctab)

```

Replotting data for each combination of values of independent variables
-----------------------------------------------------------------------

We have three independent variables: impairment, tolerance & number of states. The following plots show the means of all 50 data points for each combination of values for independent variables, together with the estimated confidence intervals.

```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=10}

data$converged = data$Iterations < 201
data$convex = data$Speaker.Convex.Cat

ds = data[,c("Number.of.states", "Impairment", "Tolerance", 
             "Speaker.entropy", "Hearer.entropy",
             "Speaker.Voronoiness", "Hearer.Voronoiness",
             "Expected.utility",
             "convex", "converged"
             )]

dm = melt(ds, id.vars = c("Number.of.states", "Impairment", "Tolerance"))

dms = summarySE(dm, groupvars = c("Number.of.states", "Impairment", "Tolerance", "variable"), measurevar="value")
dsub = dms
dsub$Number.of.states = factor(dsub$Number.of.states)

myLabeller <- function(var, value){
    value <- as.character(value)
    if (var=="variable") { 
        value[value=="Speaker.entropy"] <- "entropy sen"
        value[value=="Hearer.entropy"] <- "entropy rec"
        value[value=="Speaker.Voronoiness"] <- "Voronoiness sen"
        value[value=="Hearer.Voronoiness"] <- "Voronoiness rec"
        value[value=="Expected.utility"] <- "expected utility"
    }
    return(value)
}

pd <- position_dodge(.01)
mysize = 1.65
sp = ggplot(dsub, aes(x=Impairment, y=value, color = Number.of.states, shape=Number.of.states, linetype=Number.of.states)) + 
    geom_line(color="gray", position = pd) + xlab("imprecision") +
     geom_point(position = pd, size = mysize) +
     geom_errorbar(aes(ymin=value-ci, ymax=value+ci), width=0, position=pd) +
    scale_shape_discrete(name = "# states", guide = guide_legend()) +
    scale_linetype_discrete(name = "# states", guide = guide_legend()) +
    scale_color_discrete(name = "# states", guide = guide_legend()) +
    theme(legend.position="top")

sp = sp + facet_grid(variable ~ Tolerance, scales="free", labeller=myLabeller)

sp + facet_grid(. ~ sex, labeller=mf_labeller)
show(sp)

ggsave(sp, file = "paper/plots/MeanMetrics2.pdf", width=8, height = 10)

```

Entropy
-------

```{r fig.width=5, fig.height=5}
cor.test(data$Speaker.entropy, data$Hearer.entropy)
qplot(data$Speaker.entropy, data$Hearer.entropy) + geom_smooth(method=lm)
```

Informativity
-------------

This is the notion from Skyrms. It seems very well-behaved. There is a clear positive correlation between speaker and hearer informativity.

```{r fig.width=5, fig.height=5}
cor.test(data$Speaker.informativity, data$Hearer.informativity)
qplot(data$Speaker.informativity, data$Hearer.informativity) + geom_smooth(method=lm)
```

There is a clear negative correlation between impairment and informativity. The remaining variance here must and can be accounted for in terms of Number.of.states (see main plot).

```{r fig.width=5, fig.height=5}
cor.test(data$Impairment, data$Speaker.informativity)
qplot(data$Impairment, data$Speaker.informativity) + geom_smooth(method=lm)
```

Voronoiness
-------

```{r fig.width=5, fig.height=5}
cor.test(data$Speaker.Voronoiness, data$Hearer.Voronoiness)
qplot(data$Speaker.Voronoiness, data$Hearer.Voronoiness) + geom_smooth(method=lm)
```

Percentage of Converged and Convex strategies
---------------------------------------------

```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=10}

data$Converged = data$Iterations < 201
data$Convex = data$Speaker.Convex.Cat

ds = data[,c("Number.of.states", "Impairment", "Tolerance", 
             "Converged", "Convex")]

dm = melt(ds, id.vars = c("Number.of.states", "Impairment", "Tolerance"))

dms = summarySE(dm, groupvars = c("Number.of.states", "Impairment", "Tolerance", "variable"), measurevar="value")
dsub = dms

pd <- position_dodge(.05)
sp = ggplot(dms, aes(x=factor(Impairment), y = value, group = variable, shape = variable, linetype = variable)) 
sp = sp + geom_line(position = pd, color = "grey") + geom_point(position = pd)
sp = sp + facet_grid(Number.of.states ~ Tolerance, scales="free") + ylab("proportion") + xlab("imprecision") +
  scale_shape_discrete(name = "property") +
  scale_linetype_discrete(name = "property")
show(sp)
ggsave(sp, file = "paper/plots/CategoricalMeasures.pdf", width=8, height = 5)

```