### 2015-04-03 paper/derivation_RDD/derivation_RDD.tex
derivation of RDD from conditional imitiation - first shot

\documentclass[fleqn,reqno,10pt]{article}

%========================================
% Packages
%========================================

\usepackage[]{../../helpers/mypackages}
%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
%\bibliography{MyRefGlobal}
\bibliography{../../helpers/MyRefGlobal}
\bibliography{paper} 
\usepackage{../../helpers/myenvironments}
\usepackage{../../helpers/mycommands}
\usepackage{todonotes}
\usepackage{subcaption}

%========================================
% Standard Layout
%========================================

% Itemize
\renewcommand{\labelitemi}{\large{$\mathbf{\cdot}$}}    % itemize symbols
\renewcommand{\labelitemii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiv}{\large{$\mathbf{\cdot}$}}
% Description
\renewcommand{\descriptionlabel}[1]{\hspace\labelsep\textsc{#1}}

% Figure Captions
\usepackage{caption} % use corresponding myfiguresize!
\setlength{\captionmargin}{20pt}
\renewcommand{\captionfont}{\small}
\setlength{\belowcaptionskip}{7pt} % standard is 0pt

%========================================
% Additional layout & commands
%========================================

\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}

% Annotations
\newcommand{\mytodo}[2]{\todo[inline,color=yellow,author=#1]{#2}}
\newcommand{\question}[2]{\todo[inline,color=blue,author=#1]{#2}}
\newcommand{\answer}[2]{\todo[inline,color=green,author=#1]{#2}}

\newcommand{\rd}{\acro{rd}} % replicator dynamic
\newcommand{\rmd}{\acro{rmd}} % replicator mutator dynamic
\newcommand{\rdd}{\acro{rdd}} % replicator diffusion dynamic
\newcommand{\RD}{\ensuremath{\mathrm{RD}}} % replicator dynamic
\newcommand{\RDD}{\ensuremath{\mathrm{RDD}}} % replicator diffusion dynamic
\newcommand{\RMD}{\ensuremath{\mathrm{RMD}}} % replicator mutator

\newcommand{\Diff}{\ensuremath{\mathrm{D}}} % Difusion 
\newcommand{\Mutate}{\ensuremath{\mathrm{M}}} % Mutation

\newcommand{\impairment}{\ensuremath{\alpha}} % impairment
\newcommand{\toler}{\ensuremath{\beta}} % tolerance
\newcommand{\ns}{\ensuremath{n_s}} % number of states

\newcommand{\similarity}{\ensuremath{\mathrm{Sim}}} % similarity function

\doublespacing

\begin{document}

\section*{Derivation of the RDD by imitation-based revision}

We derive the proposed \rdd as the mean field of an imitative revision
protocol, following the example 4.3.1 of
\citet{Sandholm2010:Population-Game}. Our derivation relies on the
usual assumptions about conditional imitation at the agent level, but
must add special assumptions about how noise affects revision
probabilities. In keeping with the idea that agents in evolutionary
models are largely myopic, we assume here that noise is real but that
agents are unaware of it. When agents perceive the actual expected
utility of their choice and of another agent's choice (as is usual in
the motivation of imitation-based revision protocols), we assume
additionally that agents confuse their possibly non-veridical
subjective representations of states with the actual ones (details
below).

\paragraph{Notation.} As previously in this paper, let
$\Sstrat(\mymessg{i} \probbar \mystate{o})$ be the probability that a
speaker randomly picked from the population chooses $\mymessg{i}$ if
the actual/objective state is $\mystate{o}$. This factors in the
possibility that the sampled speaker perceives the objective state
$\mystate{o}$ as some subjective state $\mystate{s}$. In contrast, let
$\Sstrat_s(\mymessg{i} \probbar \mystate{s})$ be the probability that
a randomly sampled speaker chooses $\mymessg{i}$ if she experiences
some objective state $\mystate{o}$ as $\mystate{s}$. This is then the
hypothetical behavior of the speaker population in a (possibly
counterfactual) world without perceptual confusion. It is also what
the speakers actually do from their subjective point of view, and it
is this that changes in behavior (e.g., by imitation) will change in
the first place. Changes to $\Sstrat_s$ will, of course, also imply
changes in $\Sstrat$, as the latter is the result of what actually
happens when agents would like to do behave as in $\Sstrat_s$ in a
noisy world, as described by Equation~XYZ.

Similarly, $\Rstrat(\mystate{o} \probbar \messg)$ is the probability
that a randomly sampled receiver actually realizes objective state
$\mystate{o}$. $\Rstrat_s(\mystate{s} \probbar \messg)$ is the
probability of a subjectively intended interpretation $\mystate{s}$
that a randomly sampled receiver would pick in a noise-free
world. Again, behavior changes cause changes in $\Rstrat_s$, which in
turn influence $\Rstrat$.

Let $\EU(\messg)$ be short for $\EU(\messg,\state,\Rstrat)$ whenever
$\state$ and $\Rstrat$ are implicit, similarly for $\EU(\state)$ as a
shorthand for $\EU(\state,\messg,\Sstrat)$.

\paragraph{Standard assumptions.} Let's assume that we are in an
entirely noise-free world so that $\Sstrat = \Sstrat_s$ and $\Rstrat =
\Rstrat_s$. The standard assumptions for imitation-based revisions are
these. Agents get revision opportunities for all actual choice points
with equal probability (independently of how likely these would occur
in actual play). Take the case of a sender (receiver side is
parallel). If agent $A$ is given a revision opportunity for actual
state $\state$, she consults her own choice, which is $\mymessg{i}$
with probability $\Sstrat(\mymessg{i} \probbar \state)$, and observes
the behavior of one randomly picked other agent $B$, which is
$\mymessg{j}$ with probability $\Sstrat(\mymessg{j} \probbar
\state)$. $A$ also observes the expected utility of both choices (at
$\state$) and switches to $B$'s choice with probability $r_{ij} = [
\EU(\mymessg{j}) - \EU(\mymessg{i}) ]_+$, so that the overall
switching probability of an agent like $A$ who samples $\mymessg{i}$
for $\state$ is $p_{ij} = \Sstrat(\mymessg{j} \probbar \state) \
r_{ij}$.

\paragraph{Imitation under perceptual noise.} There are many
conceivable ways in which perceptual noise might influence the above
revision protocol. Our choice here is guided by theoretical and
practical considerations alike. We would like maintain the assumption
that agents in evolutionary models are not necessarily fully aware of
their surroundings. In particular, we assume here that agents are not
themselves aware of perceptual noise. Switching probabilities are
therefore to be defined in terms of actual expected utility (and not
subjective expectations about actual expected utility that take noise
into consideration). Moreover, we assume that agents who have
(unconscious) access to expected utilities for actual states
$\mystate{o}$ will confuse whatever subjective representation
$\mystate{s}$ they may have had during their revision opportunity for
the actual state $\mystate{o}$.

Concretely, suppose agent $A$ has a revision opportunity for objective
state $\mystate{o}$, and chooses $\mymessg{i}$ with probability
$\Sstrat(\mymessg{i} \probbar \mystate{o})$. Similarly, $B$ chooses
$\mymessg{j}$ with probability $\Sstrat(\mymessg{j} \probbar
\mystate{o})$. It may be that $A$ perceived $\mystate{o}$ as
$\mystate{s_A}$ and that $B$ perceived $\mystate{s_B}$, but if $A$ is
unaware of noise he would see expect and see no difference. A observes
(as usual) the actual expected utilities of $\mymessg{i}$ and
$\mymessg{j}$ for the actual state $\mystate{o}$. Consequently, he
updates his subjective behavior for $\mystate{o}$ (even if he has
actually chosen $\mymessg{i}$ since he perceived
$\mystate{s_A}$). This myopic revision behavior gives rise to expected
changes in the average subjective/noise-free behavior $\Sstrat_s$ as
follows:
\begin{align*}
  \dot{\Sstrat}_s(\mymessg{i} \probbar \state) & = \sum_j
  \Sstrat(\mymessg{j} \probbar \state) \cdot p_{ji} -  
  \Sstrat(\mymessg{i} \probbar \state) \cdot \sum_j p_{ij} \\
  & = \sum_j  \Sstrat(\mymessg{j} \probbar \state) \cdot
    \Sstrat(\mymessg{i} \probbar \state) \cdot [\EU(\mymessg{i}) -
    \EU(\mymessg{j})]_+ \\ 
   & \ \ \ \  - \Sstrat(\mymessg{i} \probbar \state) \cdot \sum_j
   \Sstrat(\mymessg{j} \probbar \state) \cdot [\EU(\mymessg{j}) -
    \EU(\mymessg{i})]_+ \\
   & = \Sstrat(\mymessg{i} \probbar \state) \cdot \left ( \sum_j
     \Sstrat(\mymessg{j} \probbar \state) \cdot \EU(\mymessg{i}) - \sum_j
     \Sstrat(\mymessg{j} \probbar \state) \cdot \EU(\mymessg{j})
   \right ) \\
   & = \Sstrat(\mymessg{i} \probbar \state) \cdot \left (
     \EU(\mymessg{i}) - \text{average EU at $\state$} \right )
\end{align*}
This is the (continuous-time) replicator dynamics for behavioral
strategies.

Similar arguments let us derive the receiver part. Suppose that agent
$A$ has a revision opportunity for choice point $\messg$. He and a
random other agent $B$ actually choose $\mystate{o_A}$ and
$\mystate{o_B}$ with probability $\Rstrat(\mystate{o_{A,B}} \probbar
\messg)$, even though they may have intended to realize different
interpretations $\mystate{s_A}$ and $\mystate{s_B}$ coming from
$\Rstrat_s(\mystate{o_{A,B}} \probbar \messg)$, but then
noise-perturbed. Agent $A$ observes the actual expected utility for
the actually realized interpretations and, by assumptions, believes
(myopically) that he had actually wanted to realize $\mystate{o_A}$
and that $B$ had wanted to realize $\mystate{o_B}$. Under these
assumptions, the derivation of the replicator dynamic for subjectively
intended (noise-free) receiver behavior $\Rstrat_s$ proceeds in
parallel with the above sender case.

At each time step, the noise-free behavior of agents is, of course,
perturbed by noise. This is why, after each update step, the \rdd
defined in Equation~XYZ perturbes the outcome of the replicator step
with perceptual noise.

\printbibliography[heading=bibintoc]

\end{document}