### 2015-07-24 paper/derivation_deepNoise/derivation_deepNoise.tex
notes minor edits

\documentclass[fleqn,reqno,11pt]{article}

%========================================
% Packages
%========================================

\usepackage[]{../../helpers/mypackages}
%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
%\bibliography{MyRefGlobal}
\bibliography{../../helpers/MyRefGlobal}
\bibliography{paper} 
\usepackage{../../helpers/myenvironments}
\usepackage{../../helpers/mycommands}
\usepackage{todonotes}
\usepackage{subcaption}

%========================================
% Standard Layout
%========================================

% Itemize
\renewcommand{\labelitemi}{\large{$\mathbf{\cdot}$}}    % itemize symbols
\renewcommand{\labelitemii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiv}{\large{$\mathbf{\cdot}$}}
% Description
\renewcommand{\descriptionlabel}[1]{\hspace\labelsep\textsc{#1}}

% Figure Captions
\usepackage{caption} % use corresponding myfiguresize!
\setlength{\captionmargin}{20pt}
\renewcommand{\captionfont}{\small}
\setlength{\belowcaptionskip}{7pt} % standard is 0pt

%========================================
% Additional layout & commands
%========================================

\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}

% Annotations
\newcommand{\mytodo}[2]{\todo[inline,color=yellow,author=#1]{#2}}
\newcommand{\question}[2]{\todo[inline,color=blue,author=#1]{#2}}
\newcommand{\answer}[2]{\todo[inline,color=green,author=#1]{#2}}

\newcommand{\rd}{\acro{rd}} % replicator dynamic
\newcommand{\rmd}{\acro{rmd}} % replicator mutator dynamic
\newcommand{\rdd}{\acro{rdd}} % replicator diffusion dynamic
\newcommand{\RD}{\ensuremath{\mathrm{RD}}} % replicator dynamic
\newcommand{\RDD}{\ensuremath{\mathrm{RDD}}} % replicator diffusion dynamic
\newcommand{\RMD}{\ensuremath{\mathrm{RMD}}} % replicator mutator

\newcommand{\Diff}{\ensuremath{\mathrm{D}}} % Difusion 
\newcommand{\Mutate}{\ensuremath{\mathrm{M}}} % Mutation

\newcommand{\impairment}{\ensuremath{\alpha}} % impairment
\newcommand{\toler}{\ensuremath{\beta}} % tolerance
\newcommand{\ns}{\ensuremath{n_s}} % number of states

\newcommand{\similarity}{\ensuremath{\mathrm{Sim}}} % similarity function

\usepackage{fullpage}

% \doublespacing

\begin{document}

Our goal is to derive a simple and practical dynamic for behavioral adaptations in a sim-max
game given noisy perception and realization of states. The focus here in on adaptations
triggered by imitation of behavior, so that we look at adaptations of local behavior, in terms
of behavioral strategies. We want to take noise into account (i) during the realization of
strategies and (ii) in perceiving other agents' strategy choices (for imitation).

Fix a sim-max game and fix behavioral strategies $\Sstrat$ and $\Rstrat$ that represent the
actual behavior of agents without noise or perturbations. As usual, we assume that agents carry
pure strategies and that behavioral strategies express population frequencies. So:
$\Sstrat(\messg \probbar \mystate{o})$ is the probability with which a randomly drawn sender
from the population plays $\messg$ when she perceives the state $\mystate{o}$;
$\Rstrat(\mystate{i} \probbar \messg)$ is the probability that a receiver, randomly drawn from
the population will intend to play $\mystate{i}$ when hearing message $\messg$. There could be
various sources of noise that perturb: (i) what agents observe (conditional probability
distribution $P_o$, see below) and (ii) how receivers realize intended state (interpretation)
choices (conditional probability distribution $P_r$, see below). Together with priors over
states, these are the basic probabilities from which everything else will be derived. (The
replicator dynamic is the limiting case where $P_o$ and $P_r$ are trivial.)

Concretely, we think of the process of playing the sim-max game as subject to several pieces of
stochasticity:
\begin{enumerate}
\item the actual state $\mystate{a}$ occurs with $P(\mystate{a}$)
\item any randomly sampled sender would observes state $\mystate{o}$ with probability
  $P_o(\mystate{o} \probbar \mystate{a})$
  \begin{itemize}
  \item given observed $\mystate{o}$, the probability of actual state $\mystate{a}$ is
    $P_{\overline{o}}(\mystate{a} \probbar \mystate{o}) \propto P(\mystate{a}) P_o(\mystate{o}
    \probbar \mystate{a})$
  \end{itemize}
\item a randomly sampled sender will play $\messg$ when observing $\mystate{o}$ with
  probability $\Sstrat(\messg \probbar \mystate{o})$
  \begin{itemize}
  \item when the true state is $\mystate{a}$ the probability that a randomly sampled sender
    plays $\messg$ is $P_\Sstrat(\messg \probbar \mystate{a}) = \sum_{\mystate{o}}
    P_o(\mystate{o} \probbar \mystate{a}) \Sstrat(\messg \probbar \mystate{o})$
  \end{itemize}
\item given $\messg$ the probability that a randomly drawn receiver will intend to interpret
  with $\mystate{i}$ is $\Rstrat(\mystate{i} \probbar \messg)$
\item if a receiver intends to realize $\mystate{i}$, the probability that he realizes
  $\mystate{r}$ is $P_r(\mystate{r} \probbar \mystate{i})$
\item $P_\Rstrat(\mystate{r} \probbar \messg) = \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
  \mystate{i}) \Rstrat(\mystate{i} \probbar \messg)$ is the probability that a randomly sampled
  receiver realizes $\mystate{r}$ when observing $\messg$
\end{enumerate}

We want to define a protocol with which agents update their behavior locally, for each choice
point. Given that we have many aspects of noise in the system, revision protocols will be
subject to this noise as well. As a consequence, it appears that \textbf{different revision
  protocols that would all give rise to the replicator dynamic in a noise-free setting, can
  give rise to different mean field dynamics under noise}. Our aim here is to consider one
revision protocol, namely ``imitation of success'' (Sandholm) that is plausible enough and
yields a practical and elegant mean field dynamic that we can study easily through a
discrete-time implementation.

Revision by ``imitation of success'' generally proceeds as follows. Call the possibly
updating/switching agent ``learner'' and the possibly to-be-imitated agent ``teacher''. A
random learner is drawn from the population and given a chance to change her behavior at choice
point $k$. Suppose, she currently plays $c$ at $k$. She observes what a randomly sampled
teacher would do at choice point $k$, say $c'$ and switches from $c$ to $c'$ with switch
probability $\text{SWITCH}(c \rightarrow c' , k)$. For the imitation of success protocol, this
latter switching probability is just the expected utility of $c'$ at choice point $k$ (if that
is always positive, like in our case). The mean change induced by this is:
\begin{align*}
  \text{mean change of frequency $c$, at $k$} & = \text{probability of switchers $c'
    \rightarrow c$ at $k$ (inflow)} \\
& - \text{probability of switchers $c \rightarrow c'$ at $k$ (outflow)}
\end{align*}
with:
\begin{align*}
  \text{inflow} & = \sum_{c'} \text{probability learner actually plays $c'$ at perceived $k$} \\
  & \ \ \ \times \text{probability learner observes teacher play $c$ at $k$} \\
  & \ \ \ \times \text{SWITCH}(c' \rightarrow c , k) \\
  \text{outflow} & = \sum_{c'} \text{probability learner actually plays $c$ at perceived $k$} \\
  & \ \ \ \times \text{probability learner observes teacher play $c'$ at $k$} \\
  & \ \ \ \times \text{SWITCH}(c \rightarrow c' , k)
\end{align*}

Given noise in a sim-max game, as described, this update protocol must take into account that
learners may not always perfectly perceive what the teacher is doing, and that expected
utilities (and hence switching probabilities) are also subject to noisy realizations of
strategies.

Define:
\begin{itemize}
\item $P_o(\messg \probbar \mystate{o}) = \sum_{\mystate{a}} P_{\overline{o}}(\mystate{a}
  \probbar \mystate{o}) \myts P_{\Sstrat}(\messg \probbar \mystate{a})$ is the probability that
  a learner with revision opportunity for $\mystate{o}$ perceives a random teacher produce $\messg$
\item
  $P_o(\mystate{o} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
  \myts P_{\Rstrat}(\mystate{r} \probbar \messg)$
  is the probability that a learner with a revision opportunity for $\messg$ perceives a random
  teacher produce $\mystate{o}$\footnote{We assume that the noise
    $P_o(\mystate{o} \probbar \mystate{a})$ that applies to the sender's observations also
    applies to the receiver's observations when trying to imitate other agents' strategies. We
    could easily make the model more complex and introduce another measure of noise for the
    perceptual state confusability during receiver learning. But I don't think that this is
    needed.}
\item $\EU(\messg , \mystate{o}, \Rstrat) = \sum_{\mystate{a}}  P_{\overline{o}}(\mystate{a}
  \probbar \mystate{o}) \sum_{\mystate{r}}  P_\Rstrat(\mystate{r} \probbar
  \messg) \utils(\mystate{a}, \messg, \mystate{r})$ is the actual noise-aware utility of
  sending message $\messg$ when perceiving state $\mystate{o}$ given that the receiver
  population plays a noisy realization of $\Rstrat$
\item $P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \propto P(\mystate{a})
  P_{\Sstrat}(\messg \probbar \mystate{a})$ is the probability that the real state is
  $\mystate{a}$ given that a random sender produced $\messg$
\item $\EU(\mystate{i} , \messg, \Sstrat) = \sum_{\mystate{a}}
  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \sum_{\mystate{r}}
  P_{r}(\mystate{r} \probbar \mystate{i})  \utils(\mystate{a}, \messg, \mystate{r})$ is the actual noise-aware utility of
  intending to realize interpretation $\mystate{i}$ given message $\messg$ when the sender
  population plays a noisy realization of $\Sstrat$
\end{itemize}

Given this we can derive the continuous time expected change in behavioral sender strategies as
follows:
\begin{align*}
  \Sstrat(\messg \probbar \state) & = \sum_{\messg'} \Sstrat(\messg' \probbar \state) \myts
  P_o(\messg \probbar \state) \myts \EU(\messg, \state, \Rstrat) - \sum_{\messg'} \Sstrat(\messg \probbar \state) \myts
  P_o(\messg' \probbar \state) \myts \EU(\messg', \state, \Rstrat) \\
  & = P_o(\messg \probbar \state) \myts \EU(\messg, \state, \Rstrat) - \Sstrat(\messg \probbar \state) \myts \sum_{\messg'} 
  P_o(\messg' \probbar \state) \myts \EU(\messg', \state, \Rstrat)
\end{align*}
which has a discrete-time solution that is easily implementable:
\begin{align*}
  \Sstrat'(\messg \probbar \state) \propto P_o(\messg \probbar \state) \myts \EU(\messg,
  \state, \Rstrat)
\end{align*}
The case of the receiver is analogous:
\begin{align*}
  \Rstrat(\state \probbar \messg) & = \sum_{\state'} \Rstrat(\state' \probbar \messg) \myts
  P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \sum_{\state'} \Rstrat(\state \probbar \messg) \myts
  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) \\
  & = P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \Rstrat(\state \probbar \messg) \myts \sum_{\state'} 
  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat)\\
  \Rstrat'(\state \probbar \messg) & \propto P_o(\state \probbar \messg) \myts \EU(\state,
  \messg, \Sstrat)
\end{align*}

\printbibliography[heading=bibintoc]

\end{document}