### 2015-08-12 paper/version_04/paper.tex
dynamics write up

\documentclass[fleqn,reqno,10pt]{article}

%========================================
% Packages
%========================================

\usepackage[]{../../helpers/mypackages}
%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
%\bibliography{MyRefGlobal}
\bibliography{../../helpers/MyRefGlobal}
\bibliography{paper} 
\usepackage{../../helpers/myenvironments}
\usepackage{../../helpers/mycommands}
\usepackage{todonotes}
\usepackage{subcaption}

%========================================
% Standard Layout
%========================================

% Itemize
\renewcommand{\labelitemi}{\large{$\mathbf{\cdot}$}}    % itemize symbols
\renewcommand{\labelitemii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiv}{\large{$\mathbf{\cdot}$}}
% Description
\renewcommand{\descriptionlabel}[1]{\hspace\labelsep\textsc{#1}}

% Figure Captions
\usepackage{caption} % use corresponding myfiguresize!
\setlength{\captionmargin}{20pt}
\renewcommand{\captionfont}{\small}
\setlength{\belowcaptionskip}{7pt} % standard is 0pt

%========================================
% Additional layout & commands
%========================================

\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}

% Annotations
\newcommand{\mytodo}[2]{\todo[inline,color=yellow,author=#1]{#2}}
\newcommand{\question}[2]{\todo[inline,color=blue,author=#1]{#2}}
\newcommand{\answer}[2]{\todo[inline,color=green,author=#1]{#2}}

\newcommand{\rd}{\acro{rd}} % replicator dynamic
\newcommand{\rmd}{\acro{rmd}} % replicator mutator dynamic
\newcommand{\rdd}{\acro{rdd}} % replicator diffusion dynamic
\newcommand{\RD}{\ensuremath{\mathrm{RD}}} % replicator dynamic
\newcommand{\RDD}{\ensuremath{\mathrm{RDD}}} % replicator diffusion dynamic
\newcommand{\RMD}{\ensuremath{\mathrm{RMD}}} % replicator mutator
\newcommand{\icid}{\acro{icid}} % imprecise conditional imitation dynamic
\newcommand{\ICID}{\ensuremath{\mathrm{ICID}}} % imprecise conditional imitation dynamic

\newcommand{\Diff}{\ensuremath{\mathrm{D}}} % Difusion 
\newcommand{\Mutate}{\ensuremath{\mathrm{M}}} % Mutation

\newcommand{\imprecision}{\ensuremath{\alpha}} % imprecision
\newcommand{\toler}{\ensuremath{\beta}} % tolerance
\newcommand{\ns}{\ensuremath{n_s}} % number of states

\newcommand{\similarity}{\ensuremath{\mathrm{Sim}}} % similarity function

\newcommand{\myred}[1]{\textcolor{red}{#1}}

\doublespacing

\title{Vagueness, Noise, and Signaling}
% \author{Michael Franke \and Jos\'e Pedro Correia}
\date{}

\begin{document}

\maketitle

\begin{abstract}
  Signaling games are popular models for studying the evolution of meaning, but typical
  approaches do not incorporate vagueness as a feature of successful signaling.  Complementing
  recent like-minded models, we describe an aggregate population-level dynamic that describes a
  process of imitation of successful signaling behavior under imprecise perception and
  realization of similar stimuli. Applying this new dynamic to a generalization of Lewis'
  signaling games, we show that stochastic imprecision leads to vague, yet by-and-large
  efficient signal use, and, moreover, that it unifies evolutionary outcomes and helps avoid
  suboptimal categorization. \myred{The upshot of this is that generalization emerges, at an
    aggregate level, without agents actually generalizing.}
\end{abstract}

\section{Introduction}
\label{sec:introduction}

Many concepts and expressions are vague. A vague category knows clear
cases that fall under it, clear cases that do not, and also so-called
borderline cases. Borderline cases do not clearly apply, nor do they
clearly not apply, and there may be differences between borderline
cases in terms of how well they represent the category in
question. Vagueness does not seem to dramatically affect the success
of everyday communication, but it is troublesome for some of the most
prominent theories of language and meaning. This is especially so for the logico-semantic tradition of Frege, Russell and the young
Wittgenstein which is challenged by the paradoxes vagueness gives rise
to.

But there are other intriguing aspects about vagueness. The puzzle
that we are concerned with here is how vagueness could arise and be
maintained in the first place. This is a serious issue for
functionalist accounts that maintain that concepts and linguistic
meanings evolved driven towards efficiency. Since the existence of
unclear borderline cases seems to entail inefficiency of
categorization or communication, the challenge, succinctly put forward
by \citet{Lipman2009:Why-is-Language}, is to explain how vagueness can
persist under evolutionary pressure to be optimally discriminative. We
address \emph{Lipman's problem} as a technical challenge, namely to
find a conceptually sound and mathematically coherent formal model of
the evolution of categorization and language use that yields
by-and-large efficient categorization \emph{and} vagueness.

A number of authors have recently tried to explain why vagueness evolved as something that is
itself useful
\citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,BlumeBoard2013:Intentional-Vag}.
Others have argued that vagueness is a natural byproduct of limitations in information
processing \citep[e.g.][]{FrankeJager2010:Vagueness-Signa} or the result of particular
low-level learning strategies \citep[e.g.][]{OConnor2013:The-Evolution-o}. This paper
contributes to the latter line of thought. Concretely, we introduce an dynamic that describes
average behavior resulting from imitation under confusability of similar stimuli. The dynamic
proposed here complements previous like-minded accounts. We show that stochastic noise can not
only lead to vague, yet relatively efficient signal use, but can also unify evolutionary
outcomes and help avoid suboptimal categorization. Moreover, the system shows generalization
behavior without a single generalizing agent.

The next section introduces the background against which the work presented here can be
appreciated. Section~\ref{sec:repl-diff-dynam} introduces our noise-affected imitation dynamic,
which is then explored in Section~\ref{sec:exploring-rdd}.  Section~\ref{sec:discussion}
reflects and compares our approach to others. % Two appendices cover technical material that is
% not strictly necessary to appreciate the main conceptual points of the paper.
% Appendix~\ref{sec:diffusion-as-special} shows that our replicator diffusion dynamic is a
% special case of the well-known replicator mutator dynamic. Appendix~\ref{sec:deriv-rdd-imit}
% gives a derivation of our dynamic as the most-likely path of a population whose agents imitate
% successful behavior without being aware of perceptual noise themselves.

\section{Background}
\label{sec:background}

% The view that vagueness is a natural concomitant of cognitive
% limitations of language users has been formalized in a number of ways,
% using evolutionary game theory and certain generalizations of Lewis'
% signaling games \citep{Lewis_1969:Convention}, so called
% \emph{similarity-maximizing games}, or \emph{sim-max games}, for short
% \citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct}.
% %\citep{FrankeJager2010:Vagueness-Signa,OConnor2013:Evolving-Percep,OConnor2013:The-Evolution-o}.
% Our contribution is best seen in relation to these accounts, as it also
% relies on sim-max games. Let's introduce these first, and then zoom in
% on the problem of vagueness.

\subsection{Sim-max games \& conceptual spaces}

Signaling games, as introduced by \citet{Lewis_1969:Convention}, have
a sender and a receiver. The sender knows the true state of the world,
but the receiver does not. The sender can select a signal, or message,
to reveal to the receiver, who then chooses an act. In Lewis' games,
if the receiver chooses the act that corresponds to the actual state,
the play is a success, otherwise a failure. Certain regular
combinations of sender signaling and receiver reaction make messages
meaningful, in the sense that their use is correlated systematically
to certain states or acts. To investigate the conditions under which
such meaning-generating behavior can evolve is a highly interesting
topic that we are only beginning to fully understand
\citep[e.g.][]{Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,Huttegger2007:Evolution-and-t,Pawlowitsch2008:Why-Evolution-d,Barrett2009:The-Evolution-o,HutteggerSkyrms2010:Evolutionary-Dy,Skyrms2010:Signals}.

Similarity-maximizing (short: sim-max) games are a variation of
Lewis' games where the receiver's actions are equated with the state
space (one can think of the actions as choosing states) and different states are allowed to be more or less
similar to one another. While Lewis' games treated communicative
success as a matter of black and white, sim-max games allow for a
gradient notion: the more similar the receiver's interpretation is to
the actual state, the better. Signaling games with utility-relevant
similarities in the state space are fairly standard in economics
\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor},
but have received particular attention in a more philosophical context
for reasons that will become clear presently.

A sim-max game consists of a set of states $\States$, a set of messages $\Messgs$ with much
fewer messages than states, a probability distribution $\Pr \in \Delta(\States)$ such that
$\Pr(\state)$ gives the probability that state $\state$ occurs, a similarity metric\footnote{A
  metric is a function of distance between any two points in a given space. It should satisfy
  certain axioms that ensure behavior that one would intuitively expect from the words
  ``similarity'' and ``distance'' alone, but these details do not matter for the purposes of
  this paper.}
  % : similarity should always be non-negative, zero only for the distance
  % between an element and itself, symmetric, and satisfy the triangle inequality that
  % $\similarity(x,z) \le \similarity(x,y) + \similarity(y,z)$
on states $\similarity \mycolon \States \times \States \rightarrow \mathds{R}$ such that
$\similarity(\mystate{1}, \mystate{2})$ is the (perceptual) similarity between $\mystate{1}$
and $\mystate{2}$, and a utility function $\Utils \mycolon \States \times \States \rightarrow
\mathds{R}$ such that $\Utils(\mystate{1}, \mystate{2})$ is the payoff for sender and receiver
for a play with actual state $\mystate{1}$ and receiver interpretation $\mystate{2}$. We
identify the receiver's acts with the states of the world, so that the game is one of guessing
the actual state, so to speak. The utility function should be a monotonically decreasing
function of similarity.

\citet{JagerMetzger2011:Voronoi-Languag} showed that the evolutionarily stable states of
sim-max games are remarkably systematic: the evolutionarily stable states are demonstrably
so-called Voronoi languages.\footnote{Their results were obtained for games with infinitely
  many states in $n$-dimensional Euclidean space $\States \subseteq \mathds{R}^n$ and a
  quadratic loss function for utilities
  $\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} - \mystate{2})^2$.} Roughly put, a Voronoi
language is a pair of sender and receiver strategies, such that the sender strategy partitions
the state space into convex categories, while the receiver's interpretations are the central
spots in each category. A subset $X$ of $\mathds{R}^n$ is convex if, informally put, all points
in $X$ are connected via a straight line that lies entirely in $X$; $X$ has no gaps or
dents. For example, if \States is the unit interval and all states are equiprobable, a Voronoi
language with two messages would have the sender use one message exclusively for all points in
the lower half of the unit interval and another for all points in the upper half; the
receiver's interpretations of messages are the central points, .25 and .75, in the respective
intervals.  See Figure~\ref{fig:Voronoi-language-example} for an illustration.

\begin{figure}
  \centering

  \begin{tikzpicture}[scale=3]

        \begin{scope}

          \draw[-,thick] (0,0) -- (1.5,0);

          \draw (0,0) node {\color{red}{\big{[}}};

          \draw (.75,0) node {\color{red}{\big{)}}};

          \draw (.76,0) node {\color{red}{\big{(}}};

          \draw (1.5,0) node {\color{red}{\big{]}}};

            \draw (.325,.15) node
            {\color{red}{$\Spure(\mymessg{1}  \probbar \cdot)$}};

            \draw (1.075,.15) node
            {\color{red}{$\Spure(\mymessg{2} \probbar \cdot)$}};

            \node at (.375,0) (node1) {\color{blue}{$\bullet$}};

            \node at (.375,-.3) (label1)
            {\color{blue}{${\Rpure}(\cdot \probbar \mymessg{1})$}};

            \draw[-,color=blue] (node1) -- (label1);

            \node at (1.125,0) (node2) {\color{blue}{$\bullet$}};

            \node at (1.125,-.3) (label2)
            {\color{blue}{${\Rpure}(\cdot \probbar \mymessg{2})$}};

            \draw[-] (node2) -- (label2);

            \draw (0,-.2) node {$0$};

            \draw (.75,-.2) node {$0.5$};

            \draw (1.5,-.2) node {$1$};

          \end{scope}

        \end{tikzpicture}

        \caption{Example of a Voronoi language on $\States = [0,1]$. The (pure) sender strategy
          $\Spure \in \Messgs^\States$ uses one signal for the lower half, and another for the
          upper half of the unit interval. The (pure) receiver strategy $\Rpure \in
          \States^\Messgs$ selects the central elements in the respective intervals. See
          \citet{JagerMetzger2011:Voronoi-Languag} for further details.}
  \label{fig:Voronoi-language-example}
\end{figure}

This result is interesting because it demonstrates that signaling can
impose orderly categories on a metric space, without that being the
ulterior purpose of it all. Sender and receiver can be distinct
entities, whose purpose is to communicate effectively about the actual
state. In that case, evolving Voronoi languages would explain why
\emph{linguistic categories} are well-behaved and orderly in the way
they appear to be. More abstractly, sender and receiver can also be
thought of as distinct modules in a single system, where the first
module must discretize the information it is fed by selecting a small
sample of, suggestively, category labels. These are passed to a second
module that tries to decode the original information. In this case,
evolving Voronoi languages would explain why \emph{conceptual
  categories} are well-behaved and orderly in the way that they appear
to be.

Seen in this light, sim-max games may provide a foundation to approaches in cognitive semantics
that rely on the notion of conceptual spaces.
\citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued that
natural categories are convex regions in conceptual space. If the conceptual space has a
suitable metric, convex categories can be derived from a set of prototypes. The category
corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
other. In this way, \citet{Gardenfors2000:Conceptual-Spac} argues, an efficient categorization
system can be obtained: storing the prototypes lets us recover the categories without having to
store each category's extension. However, what is left unexplained, is where the prototypes
come from, and why we would not see just any distribution of prototypes as an equally efficient
classification system. This is where sim-max games can contribute a principled approach to
deriving, in an independent way, not only convex categories but also prototypical exemplars
belonging to them.  These ideas, and more, are developed further by
\citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC},
among others.

\subsection{Vague signaling in sim-max games}

This outline of an approach to categorization using sim-max games
leaves many problems unaddressed. One of them is that natural
categories for continuously variable stimuli usually do not have
unique, point-valued prototypes and clear category boundaries. We
would like to account for the possibility of such vagueness, in
particular: (i) clear positive examples of a vague category should
show a gradient transition to clear negative examples;
(ii) prototypes should likewise be gradient regions, peaking at the
center of the vague category they represent.

\citet{DouvenDecock2011:Vagueness:-A-Co} show that
\citeauthor{Gardenfors2000:Conceptual-Spac}'s conceptual spaces
approach can be extended to account for the existence of borderline
cases. From the assumption that prototypes are extended, yet convex
regions in conceptual space, a construction algorithm is available
that yields ``collated Voronoi diagrams'' with thick boundaries
representing borderline
regions. \citet{DecockDouven2012:What-is-Graded-} show further how it
is possible to arrive at a gradient transition between categories, by
weighing in the distance of different borderline cases to various
prototypical regions. This accounts for the first of the two
desiderata mentioned above, but still assumes that crisp prototype
regions must be given.

An alternative approach is taken by \citet{FrankeJager2010:Vagueness-Signa} and
\citet{OConnor2013:The-Evolution-o} who show how the above desiderata can be met by evolving
strategies in sim-max games. The approach we take here is similar in spirit, but different in
relevant detail. A more in-depth comparison is deferred until Section~\ref{sec:discussion}. Let
us first look into our own proposal in more detail.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Imprecise conditional imitation}
\label{sec:repl-diff-dynam}

Signaling agents can adapt their dispositions to act, given some feedback about their past
success, in multiple ways. Usually, we would assume that changes in behavior should, at least
on average, lean towards increasing chances of communicative success. Such behavioral
adaptations can be described at different levels of abstraction. At the level of individual
agents, we can picture a more or less idealized process of how each agent adapts dispositions
for future actions based on various pieces of information available to the agent, e.g., about
the game, the other players' behavior and the like. More abstractly, at the level of a
population of agents, we can describe how average behavioral dispositions will evolve. The
population-level perspective abstracts over petty stochastic fluctuations and zooms in on the
general tendency or direction of evolution that ensues from some (often: simplified) behavior
at the agent-level.

To better understand the interaction of confusability of stimuli and selective pressure towards
successful communication, we look at a population-level dynamic that describes the most likely
evolutionary path of a population of signaling agents who imitate other agents' behavior but
may occasionally confuse states for one another. We call our dynamic \emph{imprecise
  conditional imitation dynamic} (\icid), because it derives from an update protocol based on
conditional imitation and allows for imprecise perception and realization of states. In the
special limiting case where imprecision vanishes, the process is just the well-known replicator
dynamic \citep{TaylorJonker1978:Evolutionary-St}. To appreciate the \icid and its motivation,
we first briefly review the replicator dynamic and its derivation from conditional imitation of
successes.

\subsection{Replicator dynamic in behavioral strategies}
\label{sec:repl-dynam-behav}

Fix a signaling game with finite states $\States$, messages $\Messgs$ and acts $\Acts$. Let
$\Pr(\cdot) \in \Delta(\States)$ be the prior distribution over states and
$\Util_{\sen,\rec} \mycolon \States \times \Messgs \times \Acts \rightarrow \mathds{R}$ the
sender's and receiver's utility functions. A \emph{behavioral strategy} is a function that maps
an agent's choice points to a probability distribution over available choices.\footnote{Our
  focus is on behavioral strategies, not \emph{mixed strategies}, i.e., probability
  distributions over functions from each choice point to an act, such as
  $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioral strategies assume that agents can
  adapt their behavior locally, so to speak, i.e., independently at each choice point. Our
  focus on behavioral strategies greatly reduces the complexity of the dynamic and simplifies
  numerical simulations. But it also seems the more plausible choice for imitation-based update
  protocols: agents only observe how, on some occasion, some other agent behaved in \emph{one}
  particular situation, not how that agent would behave in \emph{all} relevant choice
  situations; they imitate the use of a single word, so to speak, not a whole lexicon. (See
  \citet{Cressman2003:Evolutionary-Dy} for more on the difference between dynamics on mixed or
  behavioral strategies.)} The sender's behavioral strategies are functions
$\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$ to a
probability of each message $\messg \in \Messgs$ being sent in $\state$; the receiver's are
functions $\Rstrat \in \Delta(\Acts)^\Messgs$, thus mapping each message $\messg \in \Messgs$
to a probability of each act $\act \in \Acts$ being chosen in response to $\messg$. Although
behavioral strategies are probabilistic, we may conveniently imagine that every individual
agent has a non-probabilistic strategy nonetheless. If the population is large enough, the
number $\Sstrat(\messg \probbar \state)$, for instance, is then the probability that a randomly
sampled sender would have a strategy that responds with $\messg$ to $\state$.

The \emph{expected utility} of choices at each choice point is:
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\act \in \Acts}
  \Rstrat(\act \probbar \messg) \cdot U_\sen(\state, \messg, \act) \\
  \EU(\act, \messg, \Sstrat) & = \sum_{\state \in
    \States} \frac{\Pr(\state) \cdot \Sstrat(\messg \probbar \state)}{\sum_{\state'} \Pr(\state') \cdot \Sstrat(\messg \probbar \state')} \cdot
  U_\rec(\state, \messg, \act) \,.
\end{align*}
Expected utilities are sums, over all possible concrete outcomes, of the utilities of these
outcomes, weighted by how likely they are to occur. For instance, the expected utility for the
sender of choosing message $\messg \in \Messgs$, given tate $\state \in \States$ and a receiver
strategy $\Rstrat$ is the sum, for each act $\act \in \Acts$, of the probability that the
receiver will choose act $\act$ given message $\messg$ times the actual utility of using act
$\act$, given $\state$ and $\messg$.

The continuous-time replicator dynamic for behavioral strategies describes the mean rate of
change in the proportion of each choice, independently for each choice point:
\begin{align}
  \label{eq:RDContinuous}
  \dot{\Sstrat}(\messg \probbar \state) & = \underbrace{\Sstrat(\messg \probbar
    \state)}_{\text{frequency of $\messg$ at $\state$}}  \Big ( \underbrace{\EU(\messg,
  \state, \Rstrat)}_{\text{EU of $\messg$ at $\state$}} -   \underbrace{\sum_{\messg'}
      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat)}_{\text{average
      EU at choice point $\state$}} \Big ) \\
  \dot{\Rstrat}(\act \probbar \messg) & = \Rstrat(\act \probbar \messg) \thinspace \Big ( \EU(\act,
  \messg, \Sstrat) -  \sum_{\act'} \Rstrat(\act' \probbar \messg) \cdot \EU(\act',
  \messg,\Sstrat)  \Big ) \nonumber
\end{align}
The rate of change in proportion of $\messg$ choices at choice point $\state$, for instance, is
given by the difference of $\messg$'s expected utility at $\state$ and the average expected
utility at $\state$, times the probability of seeing $\messg$ at $\state$ in the first place.

For numerical simulations, it can be more convenient to work with a discrete-time formulation,
which maps current strategies $\Sstrat$ and $\Rstrat$ to strategies $\Sstrat'$ and $\Rstrat'$
that describe the population at the next discrete time step. Like the continuous-time version,
the discrete-time replicator dynamics tracks changes in frequency of choices as proportional to
their expected utilities \citep[e.g.][]{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The
  formulation given in Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one
  we will be looking at in Section~\ref{sec:exploring-rdd}, where utilities are always
  non-negative and expected utilities are always positive.}\textsuperscript{,}\footnote{The
  notation $\propto$, for ``proportional to'', means that the right-hand side must possibly yet
  be normalized so as to have probabilities sum to one. In general, writing $P(x) \propto f(x)$
  for any function $f \mycolon X \rightarrow \mathds{R}$ is shorthand for
  $P(x) = \frac{f(x)}{\sum_{x' \in X} f(x')}$. The implicit normalizing denominators for the
  replicator equations are average expected utilities, e.g., for the sender:
  $\sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg', \state,\Rstrat)$.}
\begin{align}
  \label{eq:RDDiscrete}
  \Sstrat'(\messg \probbar \state) & \propto \Sstrat(\messg \probbar \state) \cdot
    \EU(\messg, \state,\Rstrat)\,, & \Rstrat'(\act \probbar \messg) & \propto \Rstrat(\act \probbar \messg) \cdot
    \EU(\act, \messg,\Sstrat)\,.   
\end{align}
For a given choice point, say a state $\state$, the probability of seeing $\messg$ played by an
average agent in the population \emph{after} the update is proportional to the probability of
seeing it \emph{before} the update, which is $\Sstrat(\messg \probbar \state)$, times the
expected utility of $\messg$ at state $\state$. Intuitively put, frequencies of choices change
by a gradient of current frequency and a measure of how good they are.

\subsection{Conditional imitation of success}
\label{sec:cond-imit-succ}

The replicator dynamic is an abstract population-level dynamic that describes the mean expected
change of behavioral dispositions in a population of signalers. There are several ways of
deriving the replicator dynamic from agent-level processes of behavioral adaptation. We focus
here on one of the simplest: \emph{conditional imitation of success} \citep[see][for this and
many other derivations]{Sandholm2010:Population-Game}. The main idea is that agents imitate the
behavior of another agent at some choice point with a probability that is proportional to the
expected utility of the latter agent's act for that choice point. Since the sim-max games that
we will be looking at in Section~\ref{sec:exploring-rdd} have positive utilities, we may
identify the switching probability with the expected utility.

Let's consider sender strategies, as the receiver case is parallel. Revision by imitation of
success proceeds as follows. Call (misleadingly!) the agent who gets a chance to change
behavior ``learner'' and the possibly to-be-imitated agent ``teacher.'' A random learner is
drawn from the population and given a chance to change behavior at choice point $\state$. The
probability that our learner plays $\messg$ is $\Sstrat(\messg \probbar \state)$. The learner
observes what a randomly sampled teacher does at $\state$. That would be $\messg'$ with
probability $\Sstrat(\messg' \probbar \state)$. The learner then starts using $\messg'$ instead
of $\messg$ with probability $\EU(\messg', \state, \Rstrat)$. (Of course, $\messg'$ and
$\messg$ could be the same.)

If agents get repeated update chances for their choice points, the \emph{expected change} of
frequency for $\messg$ at $\state$ becomes:
\begin{align}
  \label{eq:MeanChange}
  \dot{\Sstrat}(\messg \probbar \state) = P(\messg' \rightarrow m, \state) - P(\messg \rightarrow m', \state)\,,
\end{align}
where $P(\messg' \rightarrow m, \state)$ is the ``inflow'' probability that agents switch from
$\messg'$ to $\messg$ and $P(\messg \rightarrow m', \state)$ is the ``outflow'' probability
that agents switch from $\messg$ to $\messg'$. Since we are dealing with expectations in a huge
population, the probability $P(\messg' \rightarrow m, \state)$ of switching from $\messg'$ to
$\messg$ at $\state$ can be written as:
\begin{align*}
  P(\messg' \rightarrow m, \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
    \state)}_{\text{learner plays $\messg'$}} \cdot
  \underbrace{\Sstrat(\messg \probbar \state)}_{\text{teacher plays $\messg$}} \cdot
  \underbrace{\EU(\messg, \state, \Rstrat)}_{\text{EU teacher choice}} 
  % P(\messg \rightarrow m', \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg \probbar
  %   \state)}_{\text{learner plays $\messg$}} \cdot
  % \underbrace{\Sstrat(\messg' \probbar \state)}_{\text{teacher plays $\messg'$}} \cdot
  % \underbrace{\EU(\messg', \state, \Rstrat)}_{\text{EU teacher choice}}
\end{align*}
From this, we can simplify the expression of expected change in Equation~(\ref{eq:MeanChange})
to:
\begin{align*}
  \dot{\Sstrat}(\messg \probbar \state) & = \Sstrat(\messg \probbar \state) \cdot \EU(\messg,
  \state, \Rstrat) - \Sstrat(\messg \probbar \state) \cdot  \sum_{\messg'}
      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat) \\
& = \Sstrat(\messg \probbar \state) \cdot \Big ( \EU(\messg, \state,\Rstrat) - \sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',
  \state,\Rstrat) \Big )\,,
\end{align*}
which is the continuous-time formulation of the replicator dynamic in (\ref{eq:RDContinuous}).

\subsection{Imitation of success with noise}
\label{sec:imit-succ-with}

The above derivation of the replicator dynamic assumes that agents can discriminate choices and
choice points perfectly. Let's dispense with that assumption. With an eye towards sim-max
games, we will assume that receivers choose interpretation actions $\Acts = \States$ and that
any two states, but not messages, may be confused for one another.\footnote{It is relatively
  straightforward to also incorporate confusability of messages, but this is irrelevant to our
  present purposes.} Confusability of states will affect how agents behave (on average), how
they perceive others that they might want to imitate, and how good any behavioral disposition
actually is.

To keep matters simple, let us assume that agents carry pure dispositions to act, but noise can
affect their realization. As a sender, every agent maps states to messages: these are
subjectively perceived states, and no longer necessarily also the actually occurring state. As
a receiver, every agent maps messages to state interpretations: these are intended
interpretations that need not always be faithfully realized. This means that behavioral
strategies $\Sstrat \in (\Delta(\Messgs))^\States$ and $\Rstrat \in (\Delta(\States))^\Messgs$
represent the average proportions of \emph{actual} behavioral dispositions in the population,
the realization and observation of which can be distorted by confusion of similar states.

If $\mystate{a}$ is the actual state, let $P_o(\mystate{o} \probbar \mystate{a})$ be the
probability that a given agent observes state $\mystate{o}$. Similarly, if a given receiver
intends to select interpretation $\mystate{i}$, we will assume that the probability with which
state $\mystate{r}$ is realized is $P_r(\mystate{r} \probbar
\mystate{i})$.
Section~\ref{sec:exploring-rdd} gives a concrete specification of $P_o$ and $P_r$ in terms of
similarity of states. For the time being, it is best to treat $P_o$ and $P_r$ as arbitrary
conditional confusion probabilities.

A single round of play of a sim-max game is then governed by five pieces of stochastic
information, where previously there were only three. At the outset, the actual state is chosen
with probability $P_a(\mystate{a})$. A randomly sampled sender observes $\mystate{o}$ with
probability $P_o(\mystate{o} \probbar \mystate{a})$ and so we expect $\messg$ with probability
$\Sstrat(\messg \probbar \mystate{o})$. (Senders respond to observations, not the actual
state!) As perception of messages is flawless, we expect a randomly sampled receiver to intend
to realize interpretation $\mystate{i}$ with probability
$\Rstrat(\mystate{i} \probbar \messg)$.  The actually realized interpretation will be
$\mystate{r}$ with probability $P_r(\mystate{r} \probbar \mystate{i})$.

Expected utilities of choices at choice points should also take into account that actual states
need not be observed states, and intended interpretations need not be realized
interpretations. First, note that
\begin{align*}
  P_{\overline{o}}(\mystate{a} \probbar \mystate{o}) \propto P_a(\mystate{a}) P_o(\mystate{o}
    \probbar \mystate{a})
\end{align*}
is the probability that $\mystate{a}$ is actual if $\mystate{o}$ is observed. The probability
that a random sender produces $\messg$ when the actual state is $\mystate{a}$ is:
\begin{align*}
  P_{\Sstrat}(\messg \probbar \mystate{a}) = \sum_{\mystate{o}} P_o(\mystate{o} \probbar
  \mystate{a}) \Sstrat(\messg \probbar \mystate{o})\,. 
\end{align*}
The probability that the actual state is $\mystate{a}$ if a random sender produced $\messg$ is:
\begin{align*}
  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \propto P_a(\mystate{a})
  P_{\Sstrat}(\messg \probbar \mystate{a})\,. 
\end{align*}
The probability that $\mystate{r}$ is realized by a random receiver in response to
message $\messg$ is:
\begin{align*}
  P_{\Rstrat}(\mystate{r} \probbar \messg) = \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
  \mystate{i}) \Rstrat(\mystate{i} \probbar \messg)\,.
\end{align*}
This lets us capture the expected utilities for observed states (sender) and intended
interpretations (receiver) by taking into consideration what the likely actual states and
realized interpretations will be:
\begin{align*}
  \EU(\messg , \mystate{o}, \Rstrat) & = \sum_{\mystate{a}}  P_{\overline{o}}(\mystate{a}
  \probbar \mystate{o}) \sum_{\mystate{r}}  P_\Rstrat(\mystate{r} \probbar
  \messg) \utils(\mystate{a}, \messg, \mystate{r})\,, \\
  \EU(\mystate{i} , \messg, \Sstrat) & = \sum_{\mystate{a}}
  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \sum_{\mystate{r}}
  P_{r}(\mystate{r} \probbar \mystate{i})  \utils(\mystate{a}, \messg, \mystate{r})\,.
\end{align*}
If the conditional probabilities $P_o$ and $P_r$ are trivial, i.e., assign probability 0 to the
observability and realizability of non-identical states, above definitions reduce to the
previous definitions of expected utilities. This also legitimates the overload of notation.

Presence of potential imprecision in the form of non-trivial $P_o$ and $P_r$ will also affect
the dynamic that ensues from imitation of successes. Since imprecision works differently on
senders and receivers (the former confuse choice points, the latter confuse choices), we need
to look separately at each case.

As before, suppose that senders receive a chance to change their behavior independently for a
given choice point. In the present case, this would be a chance to change how to respond to a
subjectively perceived state $\mystate{o}$, which need not be the actual one. We must then
consult the probability $P_o(\messg \probbar \mystate{o})$ that, given that the learner
observed $\mystate{o}$, he will simultaneously observe a randomly sampled teacher play
$\messg$. This is (with $P_{\overline{o}}$ and $P_{\Sstrat}$ as defined above):
\begin{align*}
  P_o(\messg \probbar \mystate{o}) = \sum_{\mystate{a}} P_{\overline{o}}(\mystate{a}
  \probbar \mystate{o}) \myts P_{\Sstrat}(\messg \probbar \mystate{a})\,.
\end{align*}
The probability $P(\messg' \rightarrow \messg, \state)$ that a randomly sampled learner switches
from $\messg'$ to $\messg$ in subjectively perceived state $\mystate{o}$ is therefore:
\begin{align*}
  P(\messg' \rightarrow \messg, \mystate{o}) = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
    \mystate{o})}_{\text{learner plays $\messg'$ at $\mystate{o}$}} \cdot
  \underbrace{P_{o}(\messg \probbar \mystate{o})}_{\text{observe teacher play $\messg$}} \cdot
  \underbrace{\EU(\messg, \mystate{o}, \Rstrat)}_{\text{EU teacher choice in learner's view}}
\end{align*}
The mean change to the proportion of $\messg$ choices at state $\state$ are then:
\begin{align}
  \label{eq:nRDSender}
  \dot{\Sstrat}(\messg \probbar \mystate{o}) & = P(\messg' \rightarrow m, \mystate{o}) - P(\messg
  \rightarrow m', \mystate{o}) \nonumber \\
  & = \sum_{\messg'} \Sstrat(\messg' \probbar \mystate{o}) \myts P_o(\messg \probbar
  \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \sum_{\messg'} \Sstrat(\messg \probbar \mystate{o}) \myts
  P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat) \nonumber \\
  & = P_o(\messg \probbar \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \Sstrat(\messg \probbar \mystate{o}) \myts \sum_{\messg'} P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat)\,.
\end{align}

The case of the receiver is mostly analogous. Presented with an update opportunity for choice
point $\messg$, a learner will observe a random teacher choose interpretation $\mystate{o}$
with probability:\footnote{We assume that the noise $P_o(\mystate{o} \probbar \mystate{a})$
  that applies to the sender's observations also applies to the receiver's observations when
  trying to imitate other agents' strategies. We could easily make the model more complex and
  introduce another measure of noise for state confusability during receiver attempts to
  imitate. We refrain from it here, because we see no immediate theoretical gain.}
\begin{align*}
  P_o(\mystate{o} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
  \myts P_{\Rstrat}(\mystate{r} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
  \myts \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
  \mystate{i}) \myts \Rstrat(\mystate{i} \probbar \messg) \,.
\end{align*}
The probability $P(\state' \rightarrow \state, \messg)$ that a randomly sampled learner switches
from using $\state'$ to $\state$ in response to $\messg$ is therefore:
\begin{align*}
  P(\state' \rightarrow \state, \messg) = \sum_{\state'} \underbrace{\Rstrat(\state' \probbar
    \messg)}_{\text{learner plays $\state'$}} \cdot
  \underbrace{P_{o}(\state \probbar \messg)}_{\text{observe teacher play $\state$}} \cdot
  \underbrace{\EU(\state, \messg, \Rstrat)}_{\text{EU teacher choice}}
\end{align*}
Parallel to those for the sender case, this gives rise to:
\begin{align}
  \label{eq:nRDReceiver}
  \dot{\Rstrat}(\state \probbar \messg) & = \sum_{\state'} \Rstrat(\state' \probbar \messg)
  \myts P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \sum_{\state'}
  \Rstrat(\state \probbar \messg) \myts
  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) \nonumber \\
  & = P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \Rstrat(\state \probbar
  \messg) \myts \sum_{\state'}
  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) 
\end{align}

The continuous-time formulations in Equations~(\ref{eq:nRDSender}) and (\ref{eq:nRDReceiver})
have elegant and practical discrete-time solutions in:\footnote{To see how the discrete-time
  formulation gives rise to the continuous-time formulation, let's assume that update steps are
  infinitesimally small, so that
  \begin{align*}
    \dot{\sigma}(\messg \probbar \state) & = \sigma'(\messg \probbar \state) - \sigma(\messg
    \probbar \state)\\ 
    &= \frac{P_o(\messg \probbar \state) \myts \EU(\messg, \state, \Rstrat)
      - \Sstrat(\messg \probbar \state) \sum_{\messg'} P_o(\messg' \probbar \state) \myts
      \EU(\messg', \state, \Rstrat) }{ \sum_{\messg'} P_o(\messg' \probbar \state) \myts
      \EU(\messg', \state, \Rstrat)}\,.
  \end{align*}
By dropping the denominator, which is constant for all $\messg$ for fixed
$\state$, we obtain the above continuous-time formulation.}
\begin{align}
  \label{eq:nRD}
  \Sstrat'(\messg \probbar \state) & \propto P_o(\messg \probbar \state) \myts \EU(\messg,
  \state, \Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto P_o(\state \probbar \messg)
  \myts \EU(\state, \messg, \Sstrat)\,.
\end{align}
This looks very much like the discrete-time formulation of the standard replicator dynamic in
(\ref{eq:RDDiscrete}). A minor difference is that expected utility here takes stochastic
confusability of states into account. The main difference is that here we are not looking at,
say, the probability $\Sstrat(\messg \probbar \state)$ with which $\messg$ is be played in
response to $\state$ on average in the population, but at the probability $P_o(\messg \probbar
\state)$ of seeing one agent pick $\messg$ given that another agent has observed
$\state$. Similarly for the receiver, we are not looking at occurrence probability
$\Sstrat(\state \probbar \messg)$ but at the probability $P_o(\state \probbar \messg)$ with
which an agent who observes $\messg$ observes another agent play $\state$. It is easy to check
formally and intuitive to appreciate that this major difference vanishes if confusion
probabilities $P_o$ and $P_r$ are trivial, i.e., assigning zero probability to all pairs of
non-identical states.

\section{Exploring noisy imitation}
\label{sec:exploring-rdd}

How does a tendency to confuse similar states interact with a drive towards more efficient
signaling strategies under the imprecise conditional imitation dynamic? To be able to properly
address this question, it pays to be more precise about similarity of states, the conditional
confusion probabilities $P_o$ and $P_r$ and the utilities of our sim-max games.

\subsection{Confusion \& utilities}
\label{sec:iterated-diffusion}

Confusion of states should intuitively be a function of their perceptual similarity: the more
similar two states are, the more likely they could be mistaken for each other.  To make this
concrete, let us assume that the state space of the sim-max game consists of $n \ge 2$ states
that are equally spaced across the unit interval, including $0$ and $1$. The distance
$\card{\mystate{i} - \mystate{j}}$ is the objective, physical similarity between two states
$\mystate{i}$ and $\mystate{j}$. Distance in physical space feeds into a perceptual similarity
function, as described by \citet{Nosofsky1986:Attention-Simil}:
\begin{align*}
  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \imprecision) =
      \begin{cases}
    1 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \imprecision^2} \right ) & \textrm{otherwise,} \\
    \end{cases}
\end{align*}
where $\imprecision \ge 0$ is an imprecision parameter. When $\imprecision=0$ agents perfectly
discriminate between states; when $\imprecision \rightarrow \infty$ agents cannot discriminate
states at all. Figure~\ref{fig:NosofskySim} gives an impression of Nosofsky-similarity for
different parameter values. Other formalizations of perceptual similarity are possible,
including ones that allow for different discriminability in different areas of the state space,
but we stick with Nosofsky's similarity function for the time being, because it is
mathematically simple, yet an established notion in mathematical psychology.

\begin{figure}
  \centering

  \includegraphics[width=0.7\textwidth]{plots/NosofskySim.pdf}

  \caption{Examples of Nosofsky-similarity for different values of
    imprecision.}
  \label{fig:NosofskySim}
\end{figure}

We further assume that the probability of confusing any two states $\mystate{i}$ and
$\mystate{j}$ is proportional to their perceived similarity and, to keep matters simple, that
observational confusability $P_o$ is simply the same as realizational confusability $P_r$ and
that both are governed by the same imprecision parameter $\alpha$:
\begin{align*}
  P_o(\mystate{o} \probbar \mystate{a}) & \propto \similarity(\mystate{o}, \mystate{a} \myts ; \myts
  \imprecision) &   P_r(\mystate{r} \probbar \mystate{i}) & \propto \similarity(\mystate{r}, \mystate{i} \myts ; \myts
  \imprecision) \,.
\end{align*}
For $\imprecision = 0$, we obtain trivial confusion probabilities, i.e., confusion probability
zero between non-identical states and our dynamic reduces to the replicator dynamics. For
$\imprecision > 0$, any state can be confused for any other state with some positive
probability. For the uninteresting case of $\imprecision \rightarrow \infty$, confusion is
maximal and every state can be perceived or realized as any other state with probability
$\nicefrac{1}{\card{\States}}$.

As for utility functions, we take another Nosofsky-style function:
\begin{align*}
  \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
      \begin{cases}
    1 & \textrm{if } \toler = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \toler = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \toler^2} \right ) & \textrm{otherwise.} \\
    \end{cases}
\end{align*}
The tolerance parameter $\toler \ge 0$ models the amount of tolerable pragmatic slack or
communicative imprecision. This choice of utility function is governed partly by convenience in
parallel to the similarity function, but also because it has, as we believe, the right general
properties for a communicative payoff function. Unlike utilities that are, say, linearly or
quadratically decreasing in physical distance
\citep[c.f.][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa}, utilities
that are exponentially decreasing in negative quadratic distance can model situations where a
small amount of imprecision in communication is tolerable, whereas similarly small differences
in intolerably far away interpretations matter very little, with a smooth transition between
these regimes \citep[c.f.][]{OConnor2013:The-Evolution-o}.

In conclusion, to structure thinking about the behavior of the \icid, the system is governed by
two parameters: imprecision $\imprecision$ and tolerance $\toler$, which can vary
independently.

\subsection{Numerical simulations}
\label{sec:simulations}

The main question we should address is whether state confusability and utility-driven imitation
reasonably interact, and if so, whether the inclusion of confusability has any noteworthy
effects on the evolving meaning of signals. It transpires that moderate levels of imprecision
$\imprecision$ can lead to reasonably efficient, yet vague signaling behavior. Moreover,
imprecision can have further accelerating and, surprisingly, unifying effects on meaning
evolution. It appears that state confusability has an effect on aggregate signaling behavior
that can be described as generalization or systematization, as if ``conclusions'' about what
works for one state are ``carried over'' to similar states. This, however, is merely an
emergent phenomenon in the sense no single agent genuinely generalizes over stimuli or reasons
about what a more systematic signaling strategy would be.

For all simulations, priors were flat: $P_a(\state) = P_a(\state')$ for all
$\state, \state'$. We ran 50 trials of the discrete-time dynamic in (\ref{eq:nRD}), starting
with randomly sampled sender and receiver strategies, for each triplet of independent parameter
values: $\ns = \card{T} \in \set{6, 10, 50, 90}$,
$\imprecision \in \set{0, 0.05, 0.1, 0.2, 0.3}$, $\toler \in \set{0.05, 0.1, 0.2, 0.3}$. Each
trial ran for a maximum of 200 update steps of the \rdd. A trial was considered converged, and
thus stopped before the maximum of 200 rounds, if the total amount of change between strategies
before and after an \rdd step was smaller than a suitably chosen threshold. It is not
guaranteed that strategies at halting time had converged to the eventual attracting state,
whether they ran for 200 rounds or not. Our notion of convergence is therefore only a
categorical measure for reaching a certain (well-considered, but eventually arbitrary) degree
of stability. In principle, rather than setting a hard limit on the maximum number of rounds
and analyzing convergence categorically, one could also run each trial until reaching our
criterion for convergence and analyze the number of rounds it took to get there.  Our choice
was initially motivated by pragmatic concerns regarding length of simulation time, but it is
also justifiable if we reflect on the nature of language evolution as a continuous process.
Language adapts to an ever-changing reality; it is rare, if not impossible, that a lexicon has
an infinite amount of time to neatly converge to a stable state. Therefore, investigating speed
of convergence within a limited time frame can be seen as better reflecting the natural
constraints of language evolution.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_ind3098.pdf}
    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0$}
    \label{fig:example_stratsA}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_ind1.pdf}
    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0.05$}
    \label{fig:example_stratsB}
  \end{subfigure}

  % \begin{subfigure}[]{0.45\textwidth}
  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp0_ind1001.pdf}
  %   \caption{$\ns = 10$, $\toler = 0.1$, $\imprecision = 0$}
  %   \label{fig:example_stratsC}
  % \end{subfigure}
  % \hfill
  % \begin{subfigure}[]{0.45\textwidth}
  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp005_ind1201.pdf}
  %   \caption{$\ns = 10$, $\toler = 0.1$, $\imprecision = 0.05$}
  %   \label{fig:example_stratsD}
  % \end{subfigure}

  \caption{Example strategies under \rdd at stopping time of our simulations.}
  \label{fig:example_strats}
\end{figure}

 Representative examples for resulting strategy pairs are given in
Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA} shows a strategy pair at
stopping time with 90 states, tolerance $\toler = 0.1$ and imprecision $\imprecision = 0$. Zero
imprecision means that the trial was effectively an application of the plain \rd. Noteworthily,
the given sender strategy approximates a pure sender strategy that crisply partitions the state
space into non-convex sets. The irregular shape of the receiver strategy suggests that the
pictured strategy pair has not yet reached a dynamically stable state under the \rd. Indeed,
the trial was stopped for reaching the maximum of 200 rounds. In contrast, the outcome of a
trial with identical parameters, except with imprecision $\imprecision = 0.0.5$, which is shown
in Figure~\ref{fig:example_stratsB}, had converged (in our technical sense) after 99 rounds of
iterated \rdd. The sender strategy shows a smooth blending from one ``category'' to the other,
and the receiver's interpretations are rather extended curves, peaking at a central point in
the relevant ``categories.''

These examples already show two interesting things. Firstly, inclusion of imprecision can lead
to seemingly well-behaved, yet vague strategies in the sense that we are after. The sender
strategy in Figure~\ref{fig:example_stratsB} identifies clear positive and clear negative cases
for each signal, with a smooth transition in-between. The receiver's interpretations of signals
can be seen as smoothed-out prototype regions. Secondly, (sender) strategies can approach
non-convex pure strategies under the replicator dynamic and linger there for vast amounts of
time, possibly indefinitely. We see this in our limited-time simulations (e.g.,
Figure~\ref{fig:example_stratsA}), but this also holds, for some types of utility functions,
for the limiting case. This was first observed by Elliott Wagner, as mentioned by
\citet{OConnor2014-OCOEPC}.  A full analysis of the dynamics of sim-max games is beyond the
scope of this paper, but we will shortly see that diffusion from confusability of states
clearly prevents evolutionary paths that meander for a long time in the vicinity of non-convex
strategies.

\subsubsection{Dependent measures}

To further explore our simulation results, we calculated metrics that
aim to numerically capture how vague, generally well-structured, and
communicatively efficient the recorded strategy pairs
were. \emph{Entropy} captures the amount of systematicity or
regularity in signal use. \emph{Convexity} captures whether a
behavioral strategy would project onto a convex pure
strategy. \emph{Expected utility} measures the communicative
efficiency of evolved strategy pairs.

\paragraph{Entropy.} This classic information-theoretic notion
captures the amount of uncertainty in a probability
distribution. Roughly put, entropy of a signaling strategy captures
inverse distance from a pure strategy. The usual definition of entropy
applies directly to mixed strategies, but provably equivalent metrics
for behavioral strategies are ready to hand:
\begin{align*}
  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs}
  \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar
  \state)) \\
  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\state \in \States}
  \Rstrat(\state \probbar \messg) \cdot \log(\Rstrat(\state \probbar
  \messg)) \,. 
\end{align*}
Values obtained by these definitions are lower bounded by $0$ and
upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States|
\cdot \log(|\Messgs|)$ and $\log(|\States^\Messgs|) = |\Messgs| \cdot
\log(|\States|)$. We work with normalized values. The sender
strategies in Figures~\ref{fig:example_stratsA} and
\ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.08$,
respectively. The receiver strategies have respective entropies $0.43$
and $0.81$. In general, we expected that vague languages would have
higher entropy than crisp ones and that increasing imprecision would
lead to increased entropy, all else being equal.

\paragraph{Convexity.} At least for sender strategies, which develop
faster than receiver strategies, it also makes sense to define a
categorical measure of convexity that compensates for potential
vagueness. To determine whether a sender strategy $\Sstrat$ is convex
despite possibly being vague, we look at the derived pure strategy
$\Spure$ for which $\Spure(\state) = \argmax_{\messg' \in \Messgs}
\Sstrat(\state,\messg')$. If that $\Spure$ is convex, we also count
$\Sstrat$ as convex. The sender strategy in
Figure~\ref{fig:example_stratsA} is not convex, while the one in
Figure~\ref{fig:example_stratsB} is.  One would a priori not expect a
systematic relation between imprecision and convexity.

\paragraph{Expected utility.} We also recorded the expected utility of
a strategy pair:
\begin{align*}
  \EU(\Sstrat,\Rstrat \myts ; \myts \toler) = \sum_{\state \in
    \States} \sum_{\messg \in \Messgs} \sum_{\state' \in \States}
  \Pr(\state) \cdot \Sstrat(\state, \messg) \cdot \Rstrat(\messg,
  \state') \cdot \Util(\state,\state' \myts ; \myts \toler)\,.
\end{align*}
To make direct comparisons across different parameter settings, we
normalize expected utility by the maximal amount of expected utility
obtainable in the relevant game. The strategy pair in
Figure~\ref{fig:example_stratsA} has a normalized expected utility of
$0.99$, the pair in Figure~\ref{fig:example_stratsB} has
$0.95$. Generally, vagueness and imprecision can be expected to
decrease expected utility
\citep[c.f.][]{Lipman2009:Why-is-Language}. The crucial question is
whether communicative success drops unacceptably fast with moderate
levels of vagueness and imprecision.

\subsubsection{Results}

\begin{figure}[t]
  \centering

  \includegraphics[width=0.75\textwidth]{plots/MeanMetrics3.pdf}

  \caption{Means of gradient and proportions of categorical measures
    for $\toler = 0.1$, $\ns \in \set{6, 10, 50}$, and $\imprecision
    \in \set{0, 0.05, 0.1, 0.2, 0.3}$. The plot shows the average of
    the entropies for the sender and receiver strategy.}
  \label{fig:MeanMetrics}
\end{figure}

Figure~\ref{fig:MeanMetrics} shows plots summarizing some results.  For perspicuity, we only
plot results for one level of tolerance $\beta = 0.1$, and leave out the case of $n_s =
90$. Still, every qualitative trend mentioned in the following applies to the whole set of
results.

As expected, increasing imprecision leads to higher entropy and lower
expected utility. Importantly, however, imprecision does not
necessarily lead to disastrous decline of communicative success. What
was not expected, on the other hand, was a relation between
imprecision and convexity. However, the corresponding plot clearly
shows that higher imprecision led to a higher number of outcomes with
convex sender strategies. Moreover, it also led to higher likelihood
of convergence. In fact, sufficient imprecision always ensured
convergence and convexity. It appears that perceptual imprecision
leads to more vagueness, slightly less communicative efficiency, but
more regular, well-behaved languages in shorter time.

Beyond promoting convexity and convergence, diffusion also has another interesting
regularizing effect on the evolution of signaling. There was very
little variation in the recorded metrics for evolved strategies, at
least for higher values of imprecision. On closer inspection, it turns
out that variability in low-imprecision conditions is not only due to
non-convergence or non-convexity. Figure~\ref{fig:MoreExample_strats}
gives two more examples of strategy pairs at stopping time. Both are
obtained for the same triple of parameters, both converged before the
maximum number of rounds, and both have
convex sender strategies. However, they are not equally efficient. In
fact, the pair in Figure~\ref{fig:example_stratsC} has a normalized
expected utility of $0.99$ while the one in
Figure~\ref{fig:example_stratsD} only has $0.89$.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_ind21.pdf}
    \caption{$\ns = 6$, $\toler = 0.2$, $\imprecision = 0$}
    \label{fig:example_stratsC}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_ind23.pdf}
    \caption{$\ns = 6$, $\toler = 0.2$, $\imprecision = 0$}
    \label{fig:example_stratsD}
  \end{subfigure}

  \caption{More example strategies under \rdd at stopping time of our simulations.}
  \label{fig:MoreExample_strats}
\end{figure}

Interestingly, this type of variability in evolutionary outcomes can
be weeded out by imprecision. To investigate this, we calculated the average
distance between evolved sender strategies within each group of trials
that had identical parameter values. We determined the distance
between strategies $\Sstrat$ and $\Sstrat'$ as the average Hellinger
distance between distributions $\Sstrat(\state)$ and
$\Sstrat'(\state)$ at each choice point $\state$:
\begin{align*}
  \text{HD}(\Sstrat,\Sstrat') & = \frac{1}{{\card{\States} \cdot
     \sqrt{2}}} \cdot  \sum_{\state \in \States} 
 \sqrt{\sum_{\messg \in  \Messgs}
         \left ( \sqrt{\Sstrat(\state,\messg)} -
         \sqrt{\Sstrat'(\state,\messg)} \right )^2 }\,.
\end{align*}
To compensate for the arbitrariness of message use, we set the distance between strategies
$\Sstrat$ and $\Sstrat'$ to be the maximum of $\text{HD}(\Sstrat,\Sstrat')$ and
$\text{HD}(\Sstrat^*,\Sstrat')$ where $\Sstrat^*$ is $\Sstrat$ with reversed message
indices. An example of the \emph{within group distance}, i.e., the average distances between
all sender strategies obtained for the same parameter values, is plotted in
Figure~\ref{fig:GroupMeasuresA} for $\beta = 0.1$ and $n_s = 10$. Despite some quantitative
differences, the general trend is the same for all other parameter settings that we tested:
with increasing imprecision, the sender strategies evolving under the \rdd were much more alike
(modulo swapping of messages). This means that perceptual imprecision unifies evolutionary
outcomes and guarantees sender strategies that are not only convex, but also regular in that
they induce a vague category split exactly in the middle of the unit interval.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/WithinGroupDistanceConcise.pdf}
    \caption{Within group distance}
    \label{fig:GroupMeasuresA}
  \end{subfigure}
    \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/WithinGroupEUConcise.pdf}
    \caption{Within group expected utility}
    \label{fig:GroupMeasuresB}
  \end{subfigure}

    \caption{Within group measures for all runs with $\toler = 0.1$ and $n_s = 10$.}
  \label{fig:GroupMeasures}
\end{figure}

This unifying property of perceptual imprecision could be considered
an evolutionary beneficial side-effect. Higher imprecision can lead to
higher \emph{within group expected utility}, defined as the average
expected utility that each evolved language scored when playing against an
arbitrary other language obtained for the same parameter values. Figure~\ref{fig:GroupMeasuresB} gives a representative
example. The observation repeats for other parameter values: while
imprecision might decrease the communicative efficiency of individual
languages, it increases the conceptual coherence and communicative
success between independently evolving strategies. It is almost as if
mere confusability of states imposes a regularity constraint on
evolving categories.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}
\label{sec:discussion}

\subsection{Interpretation of model and results}
\label{sec:model-interpretation}

The \rdd adds expected confusion of similar states to the \rd in behavioral strategies. The
result is an abstract evolutionary dynamic that describes the most likely path of development
of aggregate population behavior that is consistent with the idea that agents imitate other
agents' behavior (Appendix~\ref{sec:deriv-rdd-imit}). This would be a process of cultural
evolution operating on behavioral strategies. But we speculate that the \rdd is also extensible
to biological evolution on mixed strategies, in which case it is a special case of the \rmd
with mutation matrices derived from confusion matrices, as shown in
Appendix~\ref{sec:diffusion-as-special}.

The \rdd is not only compatible with several interpretations of how,
at an agent level, the population-level changes in strategy
proportions come about. It is also compatible with several
interpretations of diffusion. Diffusion can be pictured as a flaw or a
virtue. If we think of perceptual imprecision as a hard limitation on
information processing, vagueness is a natural byproduct of language
evolution. Results from numerical simulations show that such
imprecision leads to vagueness, but does not need to undermine the
prospect of evolving regular and by-and-large efficient signaling
behavior. On the other hand, there appears to be a possibly beneficial
side-effect of state confusability: it speeds up convergence to
regularly shaped and convex categories. This suggests that diffusion
of behavior, as we implemented here, could also be seen as a form
of stimulus-generalization that is evolutionarily beneficial, because
it accelerates the emergence of uniform category formation. This
latter view is also suggested by a like-minded account to which we
turn next.

\subsection{Relation with previous accounts}
\label{sec:relat-with-prev}

\paragraph{Generalized reinforcement.}

% Under \emph{Herrnstein reinforcement learning}, sender and receiver
% adjust their dispositions to act after each round of play, in such a
% way that the actual (non-negative) payoff gained in the previous
% interaction is added to the non-normalized propensities for repeating
% the same behavior. For sim-max games, this means that, when the sender
% chose $\messg$ in state $\state$, and this resulted in some
% non-negative payoff, the probability that the sender chooses $\messg$
% again in $\state$ is increased, but nothing else changes. In
% particular, the sender's behavior in other choice points does not
% change. \emph{Generalized reinforcement learning} is different. When
% the use of $\messg$ in $\state$ gave positive payoff, then not only
% will the probability of its future use be promoted at $\state$ but also at
% other states, proportional to how similar these are to
% $\state$. Similar amendments take care of the way that the receiver
% updates his choice dispositions.

\citet{OConnor2013:The-Evolution-o} introduces a variant of classical Herrnstein reinforcement
learning for sim-max games, in which successful play leads to reinforcement of acts also for
states similar to the one that actually occurred. This is a form of generalization to unseen
contingencies. It not only leads to vague signaling of the appropriate kind, but also speeds up
learning in such a way that, especially for sim-max games with higher numbers of states, higher
levels of communicative success are reached in shorter learning
periods. % Technically, this result is
% partly due to the fact that signalers make bigger changes to their behavioral strategies
% after each round of play under generalized reinforcement than under the Herrnstein
% variety. That only explains the speed of adaptation, not necessarily that generalization also
% leads to regularity and communicative efficiency, but it does.

Diffusion of strategies in the \rdd can also be conceived of as a form of generalization, and
works in large part quite analogously to stimulus generalization in
\citeauthor{OConnor2013:The-Evolution-o}'s approach. But there are still
differences. \citeauthor{OConnor2013:The-Evolution-o} showed that her generalized reinforcement
learning can speed up the development of efficient signaling, especially for higher numbers of
states. Complementing this, we showed that diffusion regularizes evolutionary outcomes and
prevents meandering around sub-optimal and non-convex strategies, also for low numbers of
states.

Conceptually, the \rdd is a more abstract framework than generalized reinforcement
learning. The latter is most naturally seen as an agent-based learning dynamic that has two
players adapt their individual strategies after each concrete round of play. In contrast, the
\rdd describes a more abstract, average dynamical change in behavioral dispositions at the
aggregate population level. Although the behavior of (some forms of) reinforcement learning
mirror those of the \rd at some stage in time
(\cite{BorgersSarin997:Learning-Throug,HopkinsPosch2005:Attainability-o,Beggs2005:On-the-Converge}),
this does not mean that \emph{generalized} reinforcement learning is also necessarily a
plausible population-level description of, say, generalized learning in a population of
agents. Seen in this light, generalized reinforcement learning and the \rdd nicely complement
each other, as similarly-minded accounts operating at different levels of abstraction.
Although it would be interesting to explore the possibility of a mathematical connection
between the two, this is beyond the scope of this paper.

\paragraph{Quantal response equilibria.}
\citet{FrankeJager2010:Vagueness-Signa} suggested a number of ways in which
information-processing limitations could lead to vague strategies. The model that is most
clearly related to the present approach uses the notion of a \emph{quantal response}, also
known as a \emph{soft-max function}
\citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,GoereeHolt2008:Quantal-Respons}. A
quantal response is a stochastic generalization of the best response function (i.e., the
deterministic choice of an action that maximizes expected utility). Quantal responses factor in
stochastic trembles at the level of the computation or maximization of expected utilities:
agents may make mistakes in calculating how good given choice options really are for
them. \citet{FrankeJager2010:Vagueness-Signa} demonstrate that quantal response equilibria of
sim-max games can show the desired marks of vague
signaling. Figure~\ref{fig:exampleQRE_stratsA} gives an example of a quantal response
equilibrium for a sim-max game, as used in our set-up but with higher tolerance $\toler =
0.5$.
Sender and receiver strategies look very much like what evolves under \rdd with modest values
of perceptual imprecision.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/exampleStratQRE_tolerance05.pdf}
    \caption{$\ns = 90$, $\lambda = 15$, $\toler = 0.5$}
    \label{fig:exampleQRE_stratsA}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/exampleStratQRE_tolerance005.pdf}
    \caption{$\ns = 90$, $\lambda = 15$, $\toler = 0.05$}
    \label{fig:exampleQRE_stratsB}
  \end{subfigure}

  \caption{Examples of vague quantal response equilibria. Here, $\lambda$ is an imprecision
    parameter measuring the extent of noise in the calculation of expected utilities
    \citep[see][for details]{FrankeJager2010:Vagueness-Signa}.}
  \label{fig:exampleQREs}
\end{figure}

But not all quantal response equilibria are equally plausible.  To see what the problem is,
consider Figure~\ref{fig:exampleQRE_stratsB}, which is the quantal response equilibrium for a
game with lower tolerance $\toler = 0.05$. Unlike what evolves under \rdd in this case, sender
strategies have vague boundaries also towards the end of the unit intervals. Technically, this
is because quantal responses equalize message use far away from the ``prototypical''
interpretation, not just in-between categories, so to speak. Conceptually, this is because
quantal responses introduce noise at the level of computing or maximizing expected utility of
choices.

In contrast, the \rdd implements noise at the level of the perception of different
states. Unintuitive outcomes like in Figure~\ref{fig:exampleQRE_stratsB} are thereby
avoided. Still, we believe that it would be most plausible to assume that both perceptual
confusion \emph{and} computational errors in maximizing expected utility play a role in making
vague categorization a pervasive feature of natural language meaning. Perhaps even further
factors, such as limited observations, imperfect memory or uncertainty about the exact prior
distribution of states, should be considered in order to form a more complete understanding of
the phenomenon of vagueness.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{sec:conclusions}

Vagueness is a pervasive but seemingly non-disruptive feature of
natural communication and classification systems. From an evolutionary
point of view, the challenge arises to explain how vagueness can
persist under selective pressure for precision. This is foremost a
technical challenge, probing for the possibility of integrating into a
consistent model forces that lead to vagueness and forces that lead
to efficient information transfer. This paper proposed one such model
in the context of sim-max games and explored some of its consequences.

We introduced the replicator diffusion dynamic and identified it as a special case of the
replicator mutator dynamic (Appendix~\ref{sec:diffusion-as-special}). The \rdd incorporates a
special kind of stochastic mutation due to differential confusability of similar states. It can
be thought of as the aggregate population-level behavior process that results from imitation
among agents that are unaware of perceptual confusion
(Appendix~\ref{sec:deriv-rdd-imit}). Based on data from numerical simulations, we demonstrated
that the inclusion of mild levels of perceptual imprecision does not undermine the possibility
of evolving relatively successful signaling strategies. On the contrary, strategies that
evolved under mild imprecision induce a highly regular and systematic category structure that
shows the signs of vagueness as desired. There might even be a higher-order benefit to the
presence of imprecision, in that it can accelerate convergence to optimal categorization, by
preventing evolutionary paths to stay near inefficient non-convex strategies for a long time. A
little imprecision thus also unifies and regularizes evolutionary outcomes in such a way that
sub-optimal categorization is avoided in short-term trajectories.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\printbibliography[heading=bibintoc]

\end{document}