### 2015-11-17 presentations/01_toulouse.Rmd
slides Toulouse

---
title: "Imprecise imitation in signaling games"
author: "Michael Franke & Jose Pedro Correia"
runtime: shiny
output:
  ioslides_presentation:
    css: mistyle.css
    smaller: yes
    transition: faster
---
```{r setup, include=FALSE, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, dev.args = list(bg = 'transparent'))
require('ggplot2')
require('reshape2')
require('coda')
require('ggmcmc')
require('grid')
require('rjags')
require('runjags')
```