### 2016-02-15 paper/version_05/paper.tex
Added some motivation for talking about orderly categories.

\documentclass[fleqn,reqno,10pt]{article}

%========================================
% Packages
%========================================

\usepackage[]{../../helpers/mypackages}
%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
%\bibliography{MyRefGlobal}
\bibliography{../../helpers/MyRefGlobal}
\bibliography{paper} 
\usepackage{../../helpers/myenvironments}
\usepackage{../../helpers/mycommands}
\usepackage{todonotes}
\usepackage{subcaption}

\usepackage{a4wide}

%========================================
% Standard Layout
%========================================

% Itemize
\renewcommand{\labelitemi}{\large{$\mathbf{\cdot}$}}    % itemize symbols
\renewcommand{\labelitemii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiv}{\large{$\mathbf{\cdot}$}}
% Description
\renewcommand{\descriptionlabel}[1]{\hspace\labelsep\textsc{#1}}

% Figure Captions
\usepackage{caption} % use corresponding myfiguresize!
\setlength{\captionmargin}{20pt}
\renewcommand{\captionfont}{\small}
\setlength{\belowcaptionskip}{7pt} % standard is 0pt

%========================================
% Additional layout & commands
%========================================

\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}

% Annotations
\newcommand{\mytodo}[2]{\todo[inline,color=yellow,author=#1]{#2}}
\newcommand{\question}[2]{\todo[inline,color=blue,author=#1]{#2}}
\newcommand{\answer}[2]{\todo[inline,color=green,author=#1]{#2}}

\newcommand{\rd}{\acro{rd}} % replicator dynamic
\newcommand{\rmd}{\acro{rmd}} % replicator mutator dynamic
\newcommand{\rdd}{\acro{rdd}} % replicator diffusion dynamic
\newcommand{\RD}{\ensuremath{\mathrm{RD}}} % replicator dynamic
\newcommand{\RDD}{\ensuremath{\mathrm{RDD}}} % replicator diffusion dynamic
\newcommand{\RMD}{\ensuremath{\mathrm{RMD}}} % replicator mutator
\newcommand{\icid}{\acro{icid}} % imprecise conditional imitation dynamic
\newcommand{\ICID}{\ensuremath{\mathrm{ICID}}} % imprecise conditional imitation dynamic

\newcommand{\Diff}{\ensuremath{\mathrm{D}}} % Difusion 
\newcommand{\Mutate}{\ensuremath{\mathrm{M}}} % Mutation

\newcommand{\imprecision}{\ensuremath{\alpha}} % imprecision
\newcommand{\toler}{\ensuremath{\beta}} % tolerance
\newcommand{\ns}{\ensuremath{n_s}} % number of states

\newcommand{\similarity}{\ensuremath{\mathrm{Sim}}} % similarity function

\newcommand{\myred}[1]{\textcolor{red}{#1}} 
\renewcommand{\Pr}{\ensuremath{P}}

% \doublespacing

\title{Vagueness and imprecise imitation in signaling games}
% \author{Michael Franke \and Jos\'e Pedro Correia}
\date{}

\begin{document}

\maketitle

\begin{abstract}
  Signaling games are popular models for studying the evolution of meaning, but typical
  approaches do not incorporate vagueness as a feature of successful signaling.  Complementing
  recent like-minded models, we describe an aggregate population-level dynamic that describes a
  process of imitation of successful behavior under imprecise perception and realization of
  similar stimuli. Applying this new dynamic to a generalization of Lewis' signaling games, we
  show that stochastic imprecision leads to vague, yet by-and-large efficient signal use, and,
  moreover, that it unifies evolutionary outcomes and helps avoid suboptimal
  categorization. The upshot of this is that we see \emph{as-if}-generalization at an aggregate
  level, without agents actually generalizing.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

Many concepts and expressions are vague. A vague category knows clear
cases that fall under it, clear cases that do not, and also so-called
borderline cases. Borderline cases do not clearly apply, nor do they
clearly not apply, and there may be differences between borderline
cases in terms of how well they represent the category in
question. Vagueness does not seem to dramatically affect the success
of everyday communication, but it is troublesome for some of the most
prominent theories of language and meaning. This is especially so for the logico-positivist tradition of Frege, Russell, and early
Wittgenstein, which is challenged by the paradoxes vagueness gives rise
to.

But there are other intriguing aspects about vagueness. The puzzle
that we are concerned with here is how vagueness could arise and be
maintained in the first place. This is a serious issue for
functionalist accounts that maintain that concepts and linguistic
meanings evolved driven towards efficiency. Since the existence of
unclear borderline cases seems to entail inefficiency of
categorization or communication, the challenge, succinctly put forward
by \citet{Lipman2009:Why-is-Language}, is to explain how vagueness can
persist under evolutionary pressure to be optimally discriminative. We
address \emph{Lipman's problem} as a technical challenge, namely to
find a conceptually sound and mathematically coherent formal model of
the evolution of categorization and language use that yields
by-and-large efficient categorization \emph{and} vagueness.

A number of authors have recently tried to explain why vagueness evolved as something that is
itself useful
\citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,BlumeBoard2013:Intentional-Vag}.
Others have argued that vagueness is a natural byproduct of limitations in information
processing \citep[e.g.][]{FrankeJager2010:Vagueness-Signa} or the result of particular
low-level learning strategies \citep[e.g.][]{OConnor2013:The-Evolution-o}. This paper
contributes to the latter line of thought. Concretely, we introduce a dynamic that describes
changes in aggregate population behavior as the result of imitation with probabilistic
confusability of similar stimuli. We show that stochastic noise can not only lead to vague, yet
relatively efficient signal use, but can also unify evolutionary outcomes and help avoid
suboptimal categorization: regularities in confusability of stimuli also regularize signaling
behavior and evolving categories. The system gives rise to generalization behavior without
agents actually generalizing.

The next section introduces the background against which the work presented here can be
appreciated. Section~\ref{sec:repl-diff-dynam} introduces a generalization of the replicator
dynamic which is derived from the idea that agents imitate other agents' behavior while
possibly confusing similar states. This imprecise imitation dynamic is explored in
Section~\ref{sec:exploring-rdd}. Section~\ref{sec:discussion} reflects and compares our
approach to others. Appendix~\ref{sec:formal-stuff} provides formal detail.

\section{Background}
\label{sec:background}

% The view that vagueness is a natural concomitant of cognitive
% limitations of language users has been formalized in a number of ways,
% using evolutionary game theory and certain generalizations of Lewis'
% signaling games \citep{Lewis_1969:Convention}, so called
% \emph{similarity-maximizing games}, or \emph{sim-max games}, for short
% \citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct}.
% %\citep{FrankeJager2010:Vagueness-Signa,OConnor2013:Evolving-Percep,OConnor2013:The-Evolution-o}.
% Our contribution is best seen in relation to these accounts, as it also
% relies on sim-max games. Let's introduce these first, and then zoom in
% on the problem of vagueness.

\subsection{Sim-max games \& conceptual spaces}

Signaling games, as introduced by \citet{Lewis_1969:Convention}, have
a sender and a receiver. The sender knows the true state of the world,
but the receiver does not. The sender can select a signal, or message,
to reveal to the receiver, who then chooses an act. In Lewis' games,
if the receiver chooses the act that corresponds to the actual state,
the play is a success, otherwise a failure. Certain regular
combinations of sender signaling and receiver reaction make messages
meaningful, in the sense that their use is correlated systematically
to certain states or acts. To investigate the conditions under which
such meaning-generating behavior can evolve is a highly interesting
topic that we are only beginning to fully understand
\citep[e.g.][]{Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,Huttegger2007:Evolution-and-t,Pawlowitsch2008:Why-Evolution-d,Barrett2009:The-Evolution-o,HutteggerSkyrms2010:Evolutionary-Dy,Skyrms2010:Signals}.

Similarity-maximizing (short: sim-max) games are a variation of
Lewis' games where the receiver's actions are equated with the state
space (one can think of the actions as choosing states) and different states are allowed to be more or less
similar to one another. While Lewis' games treated communicative
success as a matter of black and white, sim-max games allow for a
gradient notion: the more similar the receiver's interpretation is to
the actual state, the better. Signaling games with utility-relevant
similarities in the state space are fairly standard in economics
\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor},
but have received particular attention in a more philosophical context
for reasons that will become clear presently.

Formally, a sim-max game consists of a set of states $\States$, a set of messages $\Messgs$ typically with much
fewer messages than states, a probability distribution $\Pr \in \Delta(\States)$ such that
$\Pr(\state)$ gives the probability that state $\state$ occurs, a similarity metric\footnote{A
  metric is a function of distance between any two points in a given space. It should satisfy
  certain axioms that ensure behavior that one would intuitively expect from the words
  ``similarity'' and ``distance'' alone, but these details do not matter for the purposes of
  this paper.}
  % : similarity should always be non-negative, zero only for the distance
  % between an element and itself, symmetric, and satisfy the triangle inequality that
  % $\similarity(x,z) \le \similarity(x,y) + \similarity(y,z)$
on states $\similarity \mycolon \States \times \States \rightarrow \mathds{R}$ such that
$\similarity(\mystate{1}, \mystate{2})$ is the (perceptual) similarity between $\mystate{1}$
and $\mystate{2}$, and a utility function
$\Utils \mycolon \States \times \States \rightarrow \mathds{R}$ such that
$\Utils(\mystate{1}, \mystate{2})$ is the payoff for sender and receiver for a play with actual
state $\mystate{1}$ and receiver interpretation $\mystate{2}$. We identify the receiver's acts
with the states of the world, so that the game is one of guessing the actual state, so to
speak. The utility function should be a monotonically increasing function of similarity.
Non-probabilistic sender behavior can be represented by a pure strategy
$\Spure \in \Messgs^\States$, which deterministically defines which message would be used for
each state. Similarly, a pure receiver strategy is a function $\Rpure \in \States^\Messgs$.

\citet{JagerMetzger2011:Voronoi-Languag} showed that the evolutionarily stable states of sim-max games with infinitely many states in $n$-dimensional Euclidean space $\States \subseteq \mathds{R}^n$ and a quadratic loss function for utilities $\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} - \mystate{2})^2$ are remarkably systematic: the evolutionarily stable states are demonstrably so-called Voronoi languages.
Roughly put, a Voronoi
language is a pair of sender and receiver strategies, such that the sender strategy partitions
the state space into convex categories, while the receiver's interpretations are the central
spots in each category. A subset $X$ of $\mathds{R}^n$ is convex if, informally put, all points
in $X$ are connected via a straight line that lies entirely in $X$; $X$ has no gaps or
dents. For example, if \States is the unit interval and all states are equiprobable, a Voronoi
language with two messages could have the sender use one message exclusively for all points in
the lower half of the unit interval and another for all points in the upper half; the
receiver's interpretations of messages are the central points, .25 and .75, in the respective
intervals.  See Figure~\ref{fig:Voronoi-language-example} for an illustration.

\begin{figure}
  \centering

  \begin{tikzpicture}[scale=3]

        \begin{scope}

          \draw[-,thick] (0,0) -- (1.5,0);

          \draw (0,0) node {\color{red}{\big{[}}};

          \draw (.75,0) node {\color{red}{\big{)}}};

          \draw (.76,0) node {\color{red}{\big{(}}};

          \draw (1.5,0) node {\color{red}{\big{]}}};

            \draw (.325,.15) node
            {\color{red}{$\Spure(\mymessg{1}  \probbar \cdot)$}};

            \draw (1.075,.15) node
            {\color{red}{$\Spure(\mymessg{2} \probbar \cdot)$}};

            \node at (.375,0) (node1) {\color{blue}{$\bullet$}};

            \node at (.375,-.3) (label1)
            {\color{blue}{${\Rpure}(\cdot \probbar \mymessg{1})$}};

            \draw[-,color=blue] (node1) -- (label1);

            \node at (1.125,0) (node2) {\color{blue}{$\bullet$}};

            \node at (1.125,-.3) (label2)
            {\color{blue}{${\Rpure}(\cdot \probbar \mymessg{2})$}};

            \draw[-] (node2) -- (label2);

            \draw (0,-.2) node {$0$};

            \draw (.75,-.2) node {$0.5$};

            \draw (1.5,-.2) node {$1$};

          \end{scope}

        \end{tikzpicture}

        \caption{Example of a Voronoi language on $\States = [0,1]$. The pure sender strategy
          $\Spure$ uses one signal for the lower half, and another for the upper half of the
          unit interval. The pure receiver strategy $\Rpure$ selects the central elements in
          the respective intervals. See \citet{JagerMetzger2011:Voronoi-Languag} for further
          details.}
  \label{fig:Voronoi-language-example}
\end{figure}

It is intuitive to think that linguistic and conceptual categories are orderly in such a manner.
For example, if you consider two people, one with a height of 2m and another with a height of 2.2m, both `tall', it would be difficult to defend considering a person with a height of 2.1m `short'.
Another example is that of color categorization.
The World Color Survey project~\citep{WCS} collected color naming data for 110 unwritten languages of 45 language families.
For the great majority of these languages a pattern can be observed: basic color terms are by and large convex~\citep{Regier07,Jager2010}.
It is premature to argue that these observations can be extended to \emph{all} cases of categorization.
However, for the cases where it does apply, the result of \citet{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signaling can impose this kind of orderly categories on a metric space without that being the ulterior purpose of it all.

There are at least two potential ways of interpreting the signaling setup.
Sender and receiver can be distinct
entities, whose purpose is to communicate effectively about the actual
state. In that case, evolving Voronoi languages would explain why
\emph{linguistic categories} are well-behaved and orderly in the way
they appear to be. More abstractly, sender and receiver can also be
thought of as distinct modules in a single system, where the first
module must discretize the information it is fed by selecting a small
sample of, suggestively, category labels. These are passed to a second
module that tries to decode the original information. In this case,
evolving Voronoi languages would explain why \emph{conceptual
  categories} are well-behaved and orderly in the way that they appear
to be.
Seen in this light, sim-max games may provide a foundation to approaches in cognitive semantics
that rely on the notion of conceptual spaces.
\citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued that
natural categories are convex regions in conceptual space. If the conceptual space has a
suitable metric, convex categories can be derived from a set of prototypes. The category
corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient categorization
system can be obtained: storing the prototypes lets us recover the categories without having to
store each category's extension. However, what is left unexplained is where the prototypes
come from, and why we would not see just any distribution of prototypes as an equally efficient
classification system. This is where sim-max games can contribute a principled approach to
deriving, in an independent way, not only convex categories but also prototypical exemplars
belonging to them.  These ideas, and more, are developed further by
\citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC},
among others.

\subsection{Vague signaling in sim-max games}

This outline of an approach to categorization using sim-max games
leaves many problems unaddressed. One of them is that natural
categories for continuously variable stimuli usually do not have
unique, point-valued prototypes and clear category boundaries. We
would like to account for the possibility of such vagueness, in
particular: (i) clear positive examples of a vague category should
show a gradient transition to clear negative examples;
(ii) prototypes should likewise be gradient regions, peaking at the
center of the vague category they represent.

\citet{DouvenDecock2011:Vagueness:-A-Co} show that
\citeauthor{Gardenfors2000:Conceptual-Spac}'s conceptual spaces
approach can be extended to account for the existence of borderline
cases. From the assumption that prototypes are extended, yet convex
regions in conceptual space, a construction algorithm is available
that yields ``collated Voronoi diagrams'' with thick boundaries
representing borderline
regions. \citet{DecockDouven2012:What-is-Graded-} show further how it
is possible to arrive at a gradient transition between categories, by
weighing in the distance of different borderline cases to various
prototypical regions. This accounts for the first of the two
desiderata mentioned above, but still assumes that crisp prototype
regions must be given.

Alternative approaches are taken by, for example, \citet{FrankeJager2010:Vagueness-Signa} and
\citet{OConnor2013:The-Evolution-o}, who show different ways how the above desiderata can be met by evolving
strategies in sim-max games. The approach we take here is similar in spirit, but different in
relevant detail. A more in-depth comparison is deferred until Section~\ref{sec:discussion}. Let
us first look into our own proposal in more detail.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Imprecise imitation}
\label{sec:repl-diff-dynam}

Signaling agents can adapt their dispositions to act, given some feedback about their past
success, in multiple ways. Usually, we would assume that changes in behavior should, at least
on average, lean towards increasing chances of communicative success. Such behavioral
adaptations can be described at different levels of abstraction. At the level of individual
agents, we can picture a more or less idealized process of how each agent adapts dispositions
for future actions based on various pieces of information available to the agent. More
abstractly, at the level of a population of agents, we can describe how average behavioral
dispositions will evolve. The population-level perspective abstracts over small stochastic
fluctuations and zooms in on the general tendency or direction of evolution that ensues from
behavior at the agent-level.

To better understand the interaction of confusability of stimuli and selective pressure towards
successful communication, we look at a population-level dynamic that describes the most likely
evolutionary path of a population of signaling agents who imitate other agents' behavior but
are liable to confuse states for one another. In the special limiting case where state
confusability vanishes, the process is just the well-known replicator dynamic
\citep{TaylorJonker1978:Evolutionary-St}, which should therefore be briefly reviewed first.

\subsection{Replicator dynamic in behavioral strategies}
\label{sec:repl-dynam-behav}

Fix a sim-max game with finite states $\States$ and messages $\Messgs$. As usual, we assume
that the receiver chooses states in $\States$ in response to messages. Let
$\Pr(\cdot) \in \Delta(\States)$ be the prior distribution over states and
$\Util \mycolon \States \times \States \rightarrow \mathds{R}$ the utility function shared by
senders and receivers in the population. A \emph{behavioral strategy} is a function that maps
an agent's choice points to a probability distribution over available
choices.\footnote{\label{fn:mixedStrats}Our focus is on behavioral strategies, not \emph{mixed
    strategies}, i.e., probability distributions over functions from each choice point to an
  act, such as $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioral strategies assume
  that agents can adapt their behavior locally, i.e., independently at each choice point. Our
  focus on behavioral strategies greatly reduces the complexity of the dynamic and simplifies
  numerical simulations. But it also seems the more plausible choice for imitation-based update
  protocols of the kind we consider here: agents only observe how, on some occasion, some other
  agent behaved in \emph{one} particular situation, not how that agent would behave in
  \emph{all} relevant choice situations; they imitate the use of a single word, so to speak,
  not a whole lexicon. (See \citet{Cressman2003:Evolutionary-Dy} for more on the difference
  between dynamics on mixed or behavioral strategies.)} The sender's behavioral strategies are
functions $\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$
to a probability of each message $\messg \in \Messgs$ being sent in $\state$; the receiver's
are functions $\Rstrat \in \Delta(\States)^\Messgs$, thus mapping each message
$\messg \in \Messgs$ to a probability of each interpretation $\state \in \States$ being chosen
in response to $\messg$. Although behavioral strategies are probabilistic, evolutionary
modeling usually imagines that every individual agent has a non-probabilistic
strategy. Behavioral strategies then capture average population behavior. Assuming a virtually
infinite population, the number $\Sstrat(\messg \probbar \state)$, for instance, is then the
probability that a randomly sampled sender would send message $\messg$ if the actual state was
$\state$. Similarly, $\Rstrat(\state \probbar \messg)$ is then the probability with which a
randomly sampled receiver interprets $\messg$ as $\state$. A play of single evolutionary game
with behavioral strategies is illustrated in Figure~\ref{fig:sim-max-illustration}.

\begin{figure}
  \centering

  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]

  \node (1) {};
  \node (2) [right=of 1,label=below:$\Pr(\mystate{1})$] {$\mystate{1} \in \States$};
  \node (3) [right=of 2,label=below:$\Sstrat(\messg \probbar \mystate{1})$] {$\messg \in \Messgs$};
  \node (4) [right=of 3,label=below:$\Rstrat(\mystate{2} \probbar \messg)$] {$\mystate{2} \in \States$};

  \path
    (1) edge node {N} (2)
    (2) edge node {S} (3)
    (3) edge node {R} (4);

  \end{tikzpicture}

  \caption{A round of play in a sim-max game with behavioral strategies: Nature (N) chooses a
    state $\mystate{1}$ with probability $\Pr(\mystate{1})$; a random sender (S) selects a
    message $\messg$ with probability $\Sstrat(\messg \probbar \mystate{1})$; a random receiver
    (R) selects a state $\mystate{2}$ with probability $\Rstrat(\mystate{2} \probbar \messg)$.
    Payoff for both sender and receiver is given by $\Utils(\mystate{1}, \mystate{2})$.}
   %which is typically proportional to state similarity $\similarity(\mystate{1}, \mystate{2})$.}
  \label{fig:sim-max-illustration}
\end{figure}

The \emph{expected utility} of choices at each choice point is:\footnote{The notation
  $\propto$, for ``proportional to'', that is used here and hereafter means that the right-hand
  side might still need to be normalized so as to have probabilities sum to one. In general,
  writing $P(x) \propto f(x)$ for any function $f \mycolon X \rightarrow \mathds{R}$ is
  shorthand for $P(x) = \frac{f(x)}{\sum_{x' \in X} f(x')}$. }
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\state \in \States}
  \Rstrat(\state' \probbar \messg) \myts U(\state, \state') \\
  \EU(\state', \messg, \Sstrat) & = \sum_{\state \in
    \States} P(\state \probbar \messg)  \myts
  U(\state, \state')\,, \text{\ \  where } &
P(\state \probbar \messg) & \propto \Pr(\state) \myts \Sstrat(\messg \probbar \state) \,.
\end{align*}
Expected utilities are sums, over all possible concrete outcomes, of the utilities of these
outcomes, weighted by how likely they occur. For instance, the expected utility for the sender
of choosing message $\messg$, given state $\state$ and a receiver strategy $\Rstrat$ is the
sum, for each interpretation $\state'$, of the probability that some random receiver will
choose $\state'$ given $\messg$ times the actual utility of triple $\state$, $\messg$ and
$\state'$.

The discrete-time replicator dynamic tracks changes in frequency of choices in the population
as proportional to their expected utilities
\citep[e.g.][]{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The formulation given in
  Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one we will be looking at
  in Section~\ref{sec:exploring-rdd}, where utilities are always non-negative and expected
  utilities are always positive.}
\begin{align}
  \label{eq:RDDiscrete}
  \Sstrat'(\messg \probbar \state) & \propto \Sstrat(\messg \probbar \state) \myts
    \EU(\messg, \state,\Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto \Rstrat(\state \probbar \messg) \myts
    \EU(\state, \messg,\Sstrat)\,.   
\end{align}
For a given choice point, say a state $\state$, the probability of seeing $\messg$ played by an
average agent in the population \emph{after} the update is proportional to the probability of
seeing it \emph{before} the update, which is $\Sstrat(\messg \probbar \state)$, times the
expected utility of $\messg$ at state $\state$. Intuitively put, frequencies of choices change
by a gradient of current frequency and a measure of how good they are.

The replicator dynamic is an abstract population-level dynamic that describes the mean expected
change of behavioral dispositions in a population of signalers. There are several ways of
deriving the replicator dynamic from agent-level processes of behavioral adaptation. We focus
here on one of the simplest: \emph{imitation of success}
\citep[see][]{Sandholm2010:Population-Game}. The intuitive idea is the following (see
Appendix~\ref{sec:cond-imit-succ} for a derivation of the standard replicator dynamic from this
update scheme).  Every now and then a random agent gets a chance to alter his behavior for one
of his choice points. Say, this is a sender who gets to ``reconsider'' his choice of message
for state $\state$. The revising agent then observes what some random agent does at $\state$,
say $\messg$, and will henceforth play $\messg$ in $\state$ with a probability given by the
expected utility of $\messg$ for state $\state$.

\subsection{Noisy replicator dynamic}
\label{sec:noisy-repl-dynam}

Imitation of success, as described above, presupposes that agents make no mistakes when
observing states, or choosing interpretations. This may not always be an appropriate
assumption, especially when some states can be perceptually similar and therefore likely to be
confused for one another. Therefore, imitation dynamics could be affected by at least two
sources of probabilistic noise:

\begin{enumerate}
\item \emph{observation noise:} whenever a state $\mystate{a}$ actually occurs, the probability
  that an agent observes it as $\mystate{o}$ is $P_o(\mystate{o} \probbar \mystate{a})$;
\item \emph{realization noise:} whenever an agent intends to realize interpretation
  $\mystate{i}$, the probability that $\mystate{r}$ is realized is
  $P_r(\mystate{r} \probbar \mystate{o})$.
\end{enumerate}

\noindent A round of play of a sim-max game with these two sources of confusability of states
is pictured in Figure~\ref{fig:imprecise-sim-max-illustration}. Note that, and this is crucial,
senders respond to observations not actual states, receiver strategies determine intentions not
realized states, but payoff is calculated based on actual and realized states. Behavioral
strategies thus encode what agents actually do from their subjective point of view; or, put
differently, what they would do in a noise-free world. The noise-free situation depicted in
Figure~\ref{fig:sim-max-illustration} is then the special case where
$P_o(\mystate{o} \probbar \mystate{a}) = 1$ iff $\mystate{a} = \mystate{o}$ and
$P_r(\mystate{r} \probbar \mystate{i}) = 1$ iff $\mystate{i} = \mystate{r}$.

\begin{figure}
  \centering

  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]

  \node (1) {};
  \node (2) [right=of 1,label=below:$P_a(\mystate{a})$] {$\mystate{a} \in \States$};
  \node (2') [right=of 2,label=below:$P_o(\mystate{o} \probbar \mystate{a})$] {$\mystate{o} \in \States$};
  \node (3) [right=of 2',label=below:$\Sstrat(\messg \probbar \mystate{o})$] {$\messg \in \Messgs$};
  \node (4) [right=of 3,label=below:$\Rstrat(\mystate{i} \probbar \messg)$] {$\mystate{i} \in \States$};
  \node (4') [right=of 4,label=below:$P_r(\mystate{r} \probbar \mystate{i})$] {$\mystate{r} \in \States$};

  \path
    (1) edge node {N} (2)
    (2) edge node {} (2')
    (2') edge node {S} (3)
    (3) edge node {R} (4)
    (4) edge node {} (4');

  \end{tikzpicture}

  \caption{A round of play in a sim-max game with probabilistic confusability of states: Nature
    (N) chooses an actual state with probability $P_a(\mystate{a})$; a randomly sampled sender
    (S) observes $\mystate{o}$ with probability $P_o(\mystate{o} \probbar \mystate{a})$ and
    subsequently selects $\messg$ with probability $\Sstrat(\messg \probbar \mystate{o})$; a
    randomly sampled receiver (R) intends to realize interpretation $\mystate{i}$ with
    probability $\Rstrat(\mystate{i} \probbar \messg)$ but actually realizes interpretation
    $\mystate{r}$ with probability $P_r(\mystate{r} \probbar \mystate{i})$. Payoff for both
    sender and receiver is given by $\Utils(\mystate{a}, \mystate{r})$.}
  %which is typically proportional to state similarity $\similarity(\mystate{a}, \mystate{r})$.}
  \label{fig:imprecise-sim-max-illustration}
\end{figure}

The presence of observation and realization noise also affects imitation of successes. Here, we
focus on the main ideas. Formal detail is provided by Appendix~\ref{sec:imit-succ-with}. Take a
sender who gets to revise behavior for state $\state$. What agents can plausibly revise by
imitation is their pure strategy, which maps perceived states onto messages. When an agent gets
to revise his strategy for the perceived state $\state$, this need not necessarily be the
actual state. Moreover, when that agent observes $\state$, another agent may perceive yet a
different state. Given what that latter agent perceives, his actual (pure) strategy will
determine what he plays. In sum, to describe how likely the potential imitator observes a
message choice $\messg$, we are interested in the conditional probability
$P_o(\messg \probbar \state)$ that some other random agent selects $\messg$ when the first
agent perceives $\state$. This $P_o(\messg \probbar \state)$ is derived from the prior
probability of states, the current sender population behavior $\Sstrat$ and the given
observation noise (see Appendix~\ref{sec:imit-succ-with}). Eventually, the imitating agent adopts
$\messg$ as his choice for $\state$ with probability given by the expected utility of sending
$\messg$ when perceiving state $\state$. Expected utility should, of course, take the
probabilistic confusability of states into account as well.

Similar considerations apply to the receiver side. If an agent gets a chance to change his
intended interpretation of $\messg$, we need to look at the conditional probability
$P_o(\state \probbar \messg)$ of observing another agent realize interpretation $\state$ given
that the first agent (and therefore the second as well) perceived $\messg$. This depends on
observation and realization noise, as well as on the current receiver population behavior
$\Rstrat$.

Appendix~\ref{sec:imit-succ-with} shows how imprecise imitation of this sort leads to mean changes in population
frequencies of choices that can be covered by the following discrete-time formulation:
\begin{align}
  \label{eq:nRD}
  \Sstrat'(\messg \probbar \state) & \propto P_o(\messg \probbar \state) \myts \EU(\messg,
  \state, \Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto P_o(\state \probbar \messg)
  \myts \EU(\state, \messg, \Sstrat)\,.
\end{align}
\noindent This looks very much like the discrete-time formulation of the standard replicator
dynamic in (\ref{eq:RDDiscrete}), but there are, of course, the aforementioned
differences. Firstly, expected utility here takes stochastic confusability of states into
account. Secondly, where the standard replicator dynamic had probabilities
$\Sstrat(\messg \probbar \state)$ and $\Rstrat(\state \probbar \messg)$, we now have
$P_o(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ respectively. If there is no
observation or realization noise, $P_o(\messg \probbar \state)$ reduces to
$\sigma(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ reduces to
$\Rstrat(\state \probbar \messg)$. The noisy replicator dynamic in (\ref{eq:nRD})
conservatively extends the classic case in (\ref{eq:RDDiscrete}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Exploring imprecise imitation}
\label{sec:exploring-rdd}

How does a tendency to confuse similar states interact with selective pressure towards more
efficient signaling strategies under the imprecise conditional imitation dynamic? If confusion
probabilities are moderate and regular, in that they track similarity of states, we might
expect a regularizing effect also on evolving signaling strategies. Indeed, we hypothesize that
confusion of states can give rise to population-level aggregate behavior that looks \emph{as
  if} information about what is good for one state percolates also to similar states. In other
words, imprecise imitation may make signaling behavior on average look like agents generalize
across similar states, even if no agent actually generalizes. To explore whether confusion of
states can have this effect, we turn to numerical simulation.

\subsection{Setting the stage}
\label{sec:setting-stage}

To obtain concrete results, we must fix how to represent states, similarity of states, the
conditional confusion probabilities $P_o$ and $P_r$ and the utilities of our sim-max
games. Confusion probabilities between states should be a function of perceptual similarity:
the more similar two states are, the more likely they could be mistaken for each other.

Let the state space consist of $\ns \ge 2$ states that are equally spaced across the unit
interval, including $0$ and $1$. All states occur, for simplicity, with the same probability
(i.e., $P_a$ is uniform).  The distance $\card{\mystate{i} - \mystate{j}}$ is the objective,
physical similarity between two states $\mystate{i}$ and $\mystate{j}$. Distance in physical
space feeds into a perceptual similarity function, as described by
\citet{Nosofsky1986:Attention-Simil}:
\begin{align*}
  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \imprecision) =
      \begin{cases}
    1 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^2}{ \imprecision^2} \right ) & \textrm{otherwise,} \\
    \end{cases}
\end{align*}
where $\imprecision \ge 0$ is an imprecision parameter. When $\imprecision=0$ agents perfectly
discriminate between states; when $\imprecision \rightarrow \infty$ agents cannot discriminate
states at all. Figure~\ref{fig:NosofskySim} gives an impression of Nosofsky-similarity for
different parameter values. Other formalizations of perceptual similarity are possible,
including ones that allow for different discriminability in different areas of the state space,
but we stick with Nosofsky's similarity function for the time being, because it is
mathematically simple, yet an established notion in mathematical psychology.

\begin{figure}
  \centering

  \includegraphics[width=0.7\textwidth]{plots/NosofskySim.pdf}

  \caption{Examples of Nosofsky-similarity for different values of
    imprecision.}
  \label{fig:NosofskySim}
\end{figure}

We further assume that the probability of confusing any two states $\mystate{i}$ and
$\mystate{j}$ is proportional to their perceived similarity and, to keep matters simple, that
observation noise $P_o$ is simply the same as realizational confusability $P_r$ and
that both are governed by the same imprecision parameter $\alpha$:
\begin{align*}
  P_o(\mystate{o} \probbar \mystate{a}) & \propto \similarity(\mystate{o}, \mystate{a} \myts ; \myts
  \imprecision)\,, &   P_r(\mystate{r} \probbar \mystate{i}) & \propto \similarity(\mystate{r}, \mystate{i} \myts ; \myts
  \imprecision) \,.
\end{align*}
For $\imprecision = 0$, we obtain trivial confusion probabilities: everything is reduced to
perfect imitation and the replicator dynamic. For $\imprecision > 0$, any state can be confused
for any other state with some positive probability. For the uninteresting case of
$\imprecision \rightarrow \infty$, confusion is maximal and every state can be perceived or
realized as any other state with probability $\nicefrac{1}{\card{\States}}$.

As for utility, we define it in terms of similarity and introduce another free parameter, $\toler \ge 0$.
 The intention is for this parameter to model the amount of tolerable pragmatic slack, which we felt should be allowed to vary separately from perceptual imprecision:
\begin{align*}
  \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
      \similarity(\mystate{i}, \mystate{j} \myts ; \myts \toler)\,.
\end{align*}
This choice of utility function is governed partly by convenience, but also because we believe it has the right general properties for a communicative payoff function.
Unlike utilities that are, say, linearly or
quadratically decreasing in physical distance
\citep[c.f.][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa}, utilities
that are exponentially decreasing in negative quadratic distance can model situations where a
small amount of imprecision in communication is tolerable, whereas similarly small differences
in intolerably far away interpretations matter very little, with a smooth transition between
these regimes \citep[c.f.][]{OConnor2013:The-Evolution-o}.

In order to simplify the analysis, we focus on games with two messages, \emph{i.e.}~we fix $\card{M} = 2$.
In sum, to structure thinking about the behavior of our imprecise imitation dynamic, the system
is governed by three parameters: the number $\ns = \card{T}$ of states in the state space,
imprecision $\imprecision$ and tolerance $\toler$.

\subsection{Simulation set-up}
\label{sec:simulations}

We ran 50 trials of the discrete-time dynamic in (\ref{eq:nRD}), starting with randomly sampled
sender and receiver strategies, for each triplet of independent parameter values:
$\ns \in \set{6, 10, 50, 90}$, $\imprecision \in \set{0, 0.05, 0.1, 0.2, 0.3}$,
$\toler \in \set{0.05, 0.1, 0.2, 0.3}$. Each trial ran for a maximum of 200 update steps. A
trial was considered converged, and thus stopped before the maximum of 200 rounds, if the total
amount of change between strategies before and after an update step was smaller than a suitably
chosen threshold. It is not guaranteed that strategies at halting time had converged to the
eventual attracting state, whether they ran for 200 rounds or not. Our notion of convergence is
therefore only a categorical measure for reaching a certain (well-considered, but eventually
arbitrary) degree of stability. In other words, our notion of ``convergence'' is a measure of
relative speed: is it true that the system reached a state in which evolutionary adaptations
had slowed down almost to a halt before 200 update steps? This is motivated by pragmatic
concerns regarding length of simulation time, but also theoretically justified, because we
hypothesize that confusability of states leads to regularization of evolving strategies, which
would show exactly in an increased speed of evolutionary trajectories towards well-behaved and
regular signaling behavior.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_ind3098.pdf}
    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0$}
    \label{fig:example_stratsA}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_ind1.pdf}
    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0.05$}
    \label{fig:example_stratsB}
  \end{subfigure}

  % \begin{subfigure}[]{0.45\textwidth}
  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp0_ind1001.pdf}
  %   \caption{$\ns = 10$, $\toler = 0.1$, $\imprecision = 0$}
  %   \label{fig:example_stratsC}
  % \end{subfigure}
  % \hfill
  % \begin{subfigure}[]{0.45\textwidth}
  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp005_ind1201.pdf}
  %   \caption{$\ns = 10$, $\toler = 0.1$, $\imprecision = 0.05$}
  %   \label{fig:example_stratsD}
  % \end{subfigure}

  \caption{Example strategies under \rdd at stopping time. Each line corresponds to a message and plots, for each state, the probability that the message is used.}
  \label{fig:example_strats}
\end{figure}

Representative examples for resulting strategy pairs are given in
Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA} shows a strategy pair at
stopping time with 90 states, tolerance $\toler = 0.1$ and imprecision $\imprecision = 0$. Zero
imprecision means that the trial was effectively an application of the plain \rd. Noteworthily,
the given sender strategy approximates a pure sender strategy that crisply partitions the state
space into non-convex sets. The irregular shape of the receiver strategy suggests that the
pictured strategy pair has not yet reached a stable state. Indeed,
the trial was stopped when reaching the maximum of 200 rounds. In contrast, the outcome of a
trial with identical parameters, except with imprecision $\imprecision = 0.05$, which is shown
in Figure~\ref{fig:example_stratsB}, had converged (in our technical sense) after 99
rounds. The sender strategy shows a smooth blending from one ``category'' to the other, and the
receiver's interpretations are rather extended curves, peaking at a central point in the
relevant ``categories.''

These examples already show two interesting things. Firstly, inclusion of imprecision can lead
to seemingly well-behaved, yet vague strategies in the sense that we are after. The sender
strategy in Figure~\ref{fig:example_stratsB} identifies clear positive and clear negative cases
for each signal, with a smooth transition in-between. The receiver's interpretations of signals
can be seen as smoothed-out prototype regions. Secondly, (sender) strategies can approach
non-convex pure strategies under the replicator dynamic and linger there for vast amounts of
time, possibly indefinitely. We see this in our limited-time simulations (e.g.,
Figure~\ref{fig:example_stratsA}), but this also holds, for some types of utility functions,
for the limiting case. This was first observed by Elliott Wagner, as mentioned by
\citet{OConnor2014-OCOEPC}.  A full analysis of the dynamics of sim-max games is beyond the
scope of this paper, but we will shortly see that diffusion from confusability of states
clearly prevents evolutionary paths that meander for a long time in the vicinity of non-convex
strategies.

\subsection{Dependent measures}

To further explore our simulation results, we calculated metrics that aim to numerically
capture how vague, generally well-structured, and communicatively efficient the recorded
strategy pairs were. \emph{Entropy} captures the amount of systematicity or regularity in
signal use. \emph{Convexity} captures whether a behavioral strategy would project onto a convex
pure strategy. \emph{Expected utility} measures the communicative efficiency of evolved
strategy pairs.

\paragraph{Entropy.} This classic information-theoretic notion captures the amount of
uncertainty in a probability distribution. Roughly put, entropy of a signaling strategy
captures inverse distance from a pure strategy. The usual definition of entropy applies
directly to mixed strategies (see Footnote~\ref{fn:mixedStrats}), but provably equivalent
metrics for behavioral strategies are ready to hand:
\begin{align*}
  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs}
  \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar
  \state))\,, \\
  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\state \in \States}
  \Rstrat(\state \probbar \messg) \cdot \log(\Rstrat(\state \probbar
  \messg)) \,. 
\end{align*}
Values obtained by these definitions are lower bounded by $0$ and
upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States|
\cdot \log(|\Messgs|)$ and $\log(|\States^\Messgs|) = |\Messgs| \cdot
\log(|\States|)$. We work with normalized values. The sender
strategies in Figures~\ref{fig:example_stratsA} and
\ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.08$,
respectively. The receiver strategies have respective entropies $0.43$
and $0.81$. In general, we expected that vague languages would have
higher entropy than crisp ones and that increasing imprecision would
lead to increased entropy, all else being equal.

\paragraph{Convexity.} At least for sender strategies, which develop faster than receiver
strategies, it also makes sense to define a categorical measure of convexity that compensates
for potential vagueness. To determine whether a sender strategy $\Sstrat$ is convex despite
possibly being vague, we look at the derived pure strategy $\Spure$ for which
$\Spure(\state) = \argmax_{\messg' \in \Messgs} \Sstrat(\state,\messg')$. If that $\Spure$ is
convex, we also count $\Sstrat$ as convex. The sender strategy in
Figure~\ref{fig:example_stratsA} is not convex, while the one in
Figure~\ref{fig:example_stratsB} is.  If confusion of states can regularize signaling
strategies, like we hypothesized, we should see more convexity with increasing impairment all
else equal.

\paragraph{Expected utility.} We also recorded the expected utility of
a strategy pair:
\begin{align*}
  \EU(\Sstrat,\Rstrat \myts ; \myts \toler) = \sum_{\state \in
    \States} \sum_{\messg \in \Messgs} \sum_{\state' \in \States}
  \Pr(\state) \cdot \Sstrat(\state, \messg) \cdot \Rstrat(\messg,
  \state') \cdot \Util(\state,\state' \myts ; \myts \toler)\,.
\end{align*}
To make direct comparisons across different parameter settings, we
normalize expected utility by the maximal amount of expected utility
obtainable in the relevant game. The strategy pair in
Figure~\ref{fig:example_stratsA} has a normalized expected utility of
$0.99$, the pair in Figure~\ref{fig:example_stratsB} has
$0.95$. Generally, vagueness and imprecision can be expected to
decrease expected utility
\citep[c.f.][]{Lipman2009:Why-is-Language}. The crucial question is
whether communicative success drops unacceptably fast with moderate
levels of vagueness and imprecision.

\subsection{Results}

\begin{figure}[t]
  \centering

  \includegraphics[width=0.75\textwidth]{plots/MeanMetrics3.pdf}

  \caption{Means of gradient and proportions of categorical measures
    for $\toler = 0.1$, $\ns \in \set{6, 10, 50}$, and $\imprecision
    \in \set{0, 0.05, 0.1, 0.2, 0.3}$. The plot shows the average of
    the entropies for the sender and receiver strategy.}
  \label{fig:MeanMetrics}
\end{figure}

Figure~\ref{fig:MeanMetrics} shows plots summarizing a selected part of our findings.  For
perspicuity, we only plot results for one level of tolerance $\beta = 0.1$, and leave out the
case of $n_s = 90$. Still, every qualitative trend mentioned in the following applies to the
whole set of results.

As expected, increasing imprecision leads to higher entropy and lower expected
utility. Importantly, however, imprecision does not necessarily lead to disastrous decline of
communicative success. What is more, in line with our hypothesis that mere imprecision in
imitation behavior can lead to behavior that looks as if agents generalize across similar
stimuli, higher imprecision led to a higher number of outcomes with convex sender
strategies. It also led to higher rates of convergence. In fact, sufficient imprecision always
ensured convergence and convexity. It appears that perceptual imprecision leads to more
vagueness, slightly less communicative efficiency, but more regular, well-behaved languages in
shorter time.

Beyond promoting convexity and convergence, diffusion also has another interesting regularizing
effect on the evolution of signaling. There is very little variation in the recorded metrics
for evolved strategies, at least for higher values of imprecision. On closer inspection, it
turns out that variability in low-imprecision conditions is not only due to non-convergence or
non-convexity. Figure~\ref{fig:MoreExample_strats} gives two more examples of strategy pairs at
stopping time. Both are obtained for the same triple of parameters, both converged before the
maximum number of rounds, and both have convex sender strategies. However, they are not equally
efficient. In fact, the pair in Figure~\ref{fig:example_stratsC} has a normalized expected
utility of $0.99$ while the one in Figure~\ref{fig:example_stratsD} only has $0.89$.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_ind21.pdf}
    \caption{$\ns = 6$, $\toler = 0.2$, $\imprecision = 0$}
    \label{fig:example_stratsC}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_ind23.pdf}
    \caption{$\ns = 6$, $\toler = 0.2$, $\imprecision = 0$}
    \label{fig:example_stratsD}
  \end{subfigure}

  \caption{More example strategies under \rdd at stopping time of our simulations.}
  \label{fig:MoreExample_strats}
\end{figure}

Interestingly, this type of variability in evolutionary outcomes can
be weeded out by imprecision. To investigate this, we calculated the average
distance between evolved sender strategies within each group of trials
that had identical parameter values. We determined the distance
between strategies $\Sstrat$ and $\Sstrat'$ as the average Hellinger
distance between distributions $\Sstrat(\state)$ and
$\Sstrat'(\state)$ at each choice point $\state$:
\begin{align*}
  \text{HD}(\Sstrat,\Sstrat') & = \frac{1}{{\card{\States} \cdot
     \sqrt{2}}} \cdot  \sum_{\state \in \States} 
 \sqrt{\sum_{\messg \in  \Messgs}
         \left ( \sqrt{\Sstrat(\state,\messg)} -
         \sqrt{\Sstrat'(\state,\messg)} \right )^2 }\,.
\end{align*}
To compensate for the arbitrariness of message use, we set the distance between strategies
$\Sstrat$ and $\Sstrat'$ to be the maximum of $\text{HD}(\Sstrat,\Sstrat')$ and
$\text{HD}(\Sstrat^*,\Sstrat')$ where $\Sstrat^*$ is $\Sstrat$ with reversed message
indices. An example of the \emph{within group distance}, i.e., the average distances between
all sender strategies obtained for the same parameter values, is plotted in
Figure~\ref{fig:GroupMeasuresA} for $\beta = 0.1$ and $n_s = 10$. Despite some quantitative
differences, the general trend is the same for all other parameter settings that we tested:
with increasing imprecision, the resulting sender strategies were much more alike (modulo
swapping of messages). This means that perceptual imprecision can speed up and unify
evolutionary outcomes. It can amplify the emergence of sender strategies that are not only
convex, but also regular in that they induce a vague category split exactly in the middle of
the unit interval.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/WithinGroupDistanceConcise.pdf}
    \caption{Within group distance}
    \label{fig:GroupMeasuresA}
  \end{subfigure}
    \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/WithinGroupEUConcise.pdf}
    \caption{Within group expected utility}
    \label{fig:GroupMeasuresB}
  \end{subfigure}

    \caption{Within group measures for all runs with $\toler = 0.1$ and $n_s = 10$.}
  \label{fig:GroupMeasures}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}
\label{sec:discussion}

Our imprecise imitation dynamic leads to by-and-large successful signaling behavior, even in
the presence of noise, that shows the hallmarks of vagueness as desired. It also gives rise to
population-level behavior that looks as if agents are generalizing across similar
stimuli. Here, we would like to reflect briefly on some further conceptually relevant points
and compare our approach to related work.

\subsection{Levels of vagueness}
The imprecise imitation dynamic was introduced in Section~\ref{sec:repl-diff-dynam} as tracing
changes in the overall distribution of pure strategies in a population:
$\Sstrat(\messg \probbar \state)$ was said to represent the probability that a randomly sampled
sender would have a strategy that responds with $\messg$ to $\state$ (likewise for the
receiver). This is in line with the standard interpretation of the replicator equation, but we
should consider its philosophical implications. Based on this picture, vagueness in signal use
would seem to be characterized as a strictly population-level phenomenon since it arises in a
signaling system from the inability of individual agents to fully align their (non-vague)
strategies because of imprecision. This is, we believe, a plausible mechanism that can already
explain the existence of vagueness in a language even if we assume that each agent commands a
non-vague idiolect.

We would not, however, want to commit to the idea that vagueness does not exist at the level of
individual agents. True, our derivation of the imprecise imitation dynamic assumed that agents
carry and revise pure strategies. But that was an assumption of convenience, not of
conviction. Moreover, even if individual agents command a non-vague pure strategy, the
realization of that pure strategy, according to our model, is bound to be vague (e.g., the same
agent could signal differently in repeated exposure to the same state). We have used the term
``observation noise'' here, but it may well be that this could equally well be perceived as an
inseparable component of an agent's ``signaling faculty.'' In this sense, then, the model might
not be entirely incompatible with a picture of agents who have internalized a vague signaling
strategy. It would need to be seen, however, how revision of non-deterministic individual-level
behavior must be spelled out rigorously and whether the resulting population-level dynamic
would be equivalent to our present proposal in all relevant respects.

% Moreover, we believe that the model may be interpreted in yet alternative ways. The
% probability $\Sstrat(\messg \probbar \state)$ merely gives us the likelihood that $\messg$ will
% be used by a sender when observing state $\state$.  Whether the uncertainty represents
% variations in a population of agents playing pure strategies, a single sender using a
% non-deterministic strategy, or even a heterogeneous population of non-deterministic senders, is
% up for interpretation.  The first option supported what we believe to be a plausible motivation
% for the mechanism of imitation of success with imprecision, but other motivations could be
% imagined to better fit an interpretation that includes non-deterministic agents.  We thus
% believe that the model introduced in this paper does not exclude vagueness at the individual
% level, but to explore all possible interpretations would go beyond our current scope.

\subsection{Evolutionary benefits of imprecision}
The inclusion of confusability of similar states has noteworthy effects on the evolving meaning
of signals.  It transpired from our results that imprecision can have further accelerating and,
surprisingly, unifying effects on meaning evolution.  The unifying property of perceptual
imprecision could be considered an evolutionarily beneficial side-effect. A certain degree of
imprecision can lead to higher \emph{within group expected utility}, defined as the average
expected utility that each evolved language scored when playing against an arbitrary other
language obtained for the same parameter values. Figure~\ref{fig:GroupMeasuresB} gives a
representative example. The observation repeats for other parameter values: while imprecision
might decrease the communicative efficiency of individual languages, it increases the
conceptual coherence and communicative success between independently evolving strategies. It is
as if mere confusability of states imposes a regularity constraint on evolving categories.

The phenomenon could potentially not just be a side effect.  Given the benefit of a certain
amount of imprecision we observe when comparing within group expected utility, it would be
interesting to study whether, under certain conditions, this group level advantage could trump
the individual level disadvantage of a vague language, and thus actually select for a certain
amount of imprecision.  This could be achieved by letting the imprecision parameter be an
evolving part in the dynamics as well. The idea is for now mostly speculative, but we consider
it an interesting avenue for future research.  There is additional motivation to consider its
potential if we see it in the eyes of multilevel selection theory
\citep{Wilson1994,OGorman2008}.

\subsection{Related work}
\citet{OConnor2013:The-Evolution-o} makes a proposal related to ours based on a version of
reinforcement learning for sim-max games, in which successful play leads to reinforcement of
choice options also for states similar to the ones that actually occurred. This not only leads
to vague signaling of the appropriate kind, but also speeds up learning in such a way that,
especially for sim-max games with higher numbers of states, higher levels of communicative
success are reached in shorter learning periods. Our results complement and extend
\citeauthor{OConnor2013:The-Evolution-o}'s. The most important differences are that we obtain
similar regularizing effects also for cases with low numbers of states and do not assume that
agents have any kind of generalizing capacity in and of themselves, even if that is only
implicit, as in \citeauthor{OConnor2013:The-Evolution-o}'s generalized reinforcement learning.
State confusability has an effect on aggregate signaling behavior that can be described as
generalization without generalizers: the dynamics of imprecise imitation look as if
``conclusions'' about what works for one state are ``carried over'' to similar states. This,
however, is merely an epiphenomenon in the sense that no single agent genuinely generalizes
over stimuli or reasons about what a more systematic signaling strategy would be.

\citet{FrankeJager2010:Vagueness-Signa} suggested a number of ways in which
in\-for\-ma\-tion-processing limitations could lead to vague strategies. The model that is most
clearly related to the present approach uses the notion of a \emph{quantal response}, also
known as a \emph{soft-max function}
\citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,GoereeHolt2008:Quantal-Respons}. The
main difference between this and our present approach is in where stochastic noise is assumed
to reside. In case of a quantal response dynamic, it is the computation of expected utilities;
in case of imprecise imitation, it is perception and realization of similar states. There are
cases, then, where evolving signaling behavior, as predicted by these two approaches, is quite
different. Intuitively speaking, for a case with two messages, the further we venture away from
a prototypical interpretation of either message, the less discriminative a signaling strategy
would be when the source of ``trembles'' is the computation of expected utilities: to wit,
since both ``tall'' and ``short'' are almost equally bad descriptions for a giant, quantal
response dynamics predict that senders would be almost indifferent. Sender behavior that
evolves under confusion of states does not have this puzzling property, because a giant would
not likely be confused for a dwarf.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}
\label{sec:conclusion}

We set out to meet a technical challenge posed by Lipman's problem
\citep{Lipman2009:Why-is-Language}: is there a conceptually sound and mathematically coherent
formal model that shows how vague language can evolve under selective pressure for efficient
communication? On the assumption that agents are likely to mistake states for one another,
especially when they are perceptually similar, we derived a generalization of the replicator
dynamic from an agent-level process of imprecise imitation. The resulting population-level
dynamic produced signaling behavior that is at the same time regular and by-and-large
communicatively efficient, while also showing the crucial marks of vagueness. In a sense, the
model derives vagueness as a by-product of an arguably natural limitation on the discriminatory
power of signaling agents.  Although inability to sharply discriminate similar stimuli may lead
to vagueness and bring about a (slight) decrease in communicative efficiency, there may also be
an advantage, not of vagueness itself, but of its reason: systematicity in the confusability of
states (which may be a natural by-product of the perceptual system) supports
\emph{as-if}-generalization at the population-level without having to assume that agents
themselves have any generalization power.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Imprecise conditional imitation}
\label{sec:formal-stuff}

The goal of this section is to provide technical details for
Section~\ref{sec:repl-diff-dynam}. We first show, in Section~\ref{sec:cond-imit-succ}, how to
derive the standard replicator dynamic from noise-free imitation of success. Then, in
Section~\ref{sec:imit-succ-with}, we derive the imprecise imitation dynamic by a parallel
chain of arguments.

\subsection{Deriving the replicator dynamic from imitation of success}
\label{sec:cond-imit-succ}

The main idea behind \emph{imitation of success} is that agents imitate the behavior of other
agents at some choice point with a probability that is proportional to the expected utility of
the latter agents' choice. Since the sim-max games that we are looking at here have positive
utilities upper bound by 1, we can identify the switching probability with the expected
utility.

Let's consider sender strategies, as the receiver case is parallel. Call (misleadingly!) the
agent who gets a chance to change behavior ``learner'' and the possibly to-be-imitated agent
``teacher.'' A random learner is drawn from the population and given a chance to change
behavior at choice point $\state$. The probability that our learner plays $\messg$ is
$\Sstrat(\messg \probbar \state)$. The learner observes what a randomly sampled teacher does at
$\state$. That would be $\messg'$ with probability $\Sstrat(\messg' \probbar \state)$. The
learner then starts using $\messg'$ instead of $\messg$ with probability
$\EU(\messg', \state, \Rstrat)$. (Of course, $\messg'$ and $\messg$ could be the same; the
learner could even be the teacher as well, by random sampling.)

If agents get repeated update chances for their choice points, the \emph{expected change} of
frequency of $\messg$-choices at $\state$ becomes:
\begin{align}
  \label{eq:MeanChange}
  \dot{\Sstrat}(\messg \probbar \state) = P(\messg' \rightarrow m, \state) - P(\messg \rightarrow m', \state)\,,
\end{align}
where $P(\messg' \rightarrow m, \state)$ is the ``inflow'' probability that agents switch from
any $\messg'$ to $\messg$ and $P(\messg \rightarrow m', \state)$ is the ``outflow'' probability
that agents switch from $\messg$ to any $\messg'$. Since we are dealing with expectations in a
huge population, these can be spelled out as:
\begin{align*}
  P(\messg' \rightarrow m, \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
    \state)}_{\text{learner plays $\messg'$}} \cdot
  \underbrace{\Sstrat(\messg \probbar \state)}_{\text{teacher plays $\messg$}} \cdot
  \underbrace{\EU(\messg, \state, \Rstrat)}_{\text{EU teacher choice}} \\
  P(\messg \rightarrow m', \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg \probbar
    \state)}_{\text{learner plays $\messg$}} \cdot
  \underbrace{\Sstrat(\messg' \probbar \state)}_{\text{teacher plays $\messg'$}} \cdot
  \underbrace{\EU(\messg', \state, \Rstrat)}_{\text{EU teacher choice}}
\end{align*}
From this, we can simplify the expression of expected change in Equation~(\ref{eq:MeanChange})
to:
\begin{align*}
  \dot{\Sstrat}(\messg \probbar \state) & = \Sstrat(\messg \probbar \state) \cdot \EU(\messg,
  \state, \Rstrat) - \Sstrat(\messg \probbar \state) \cdot  \sum_{\messg'}
      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat) \\
& = \underbrace{\Sstrat(\messg \probbar
    \state)}_{\text{frequency of $\messg$ at $\state$}}  \Big ( \underbrace{\EU(\messg,
  \state, \Rstrat)}_{\text{EU of $\messg$ at $\state$}} -   \underbrace{\sum_{\messg'}
      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat)}_{\text{average
      EU at choice point $\state$}} \Big )\,.
\end{align*}
This latter formulation is the continuous-time version of the replicator dynamic. We obtain a
discrete-time formulation from it by assuming that discrete update steps are infinitesimally
small, so that:
 \begin{align*}
    \dot{\sigma}(\messg \probbar \state) & = \sigma'(\messg \probbar \state) - \sigma(\messg
    \probbar \state)\\ 
    & = \frac{\Sstrat(\messg \probbar
    \state)  \myts \EU(\messg,\state,\Rstrat)}{\sum_{\messg'}\Sstrat(\messg' \probbar
    \state)  \myts \EU(\messg,\state,\Rstrat)} - \sigma(\messg
      \probbar \state) \\
    & = \frac{\Sstrat(\messg \probbar
    \state)  \myts \EU(\messg,\state,\Rstrat) - \Sstrat(\messg \probbar \state) \sum_{\messg'}\Sstrat(\messg' \probbar
    \state)  \myts \EU(\messg,\state,\Rstrat) }{\sum_{\messg'}\Sstrat(\messg' \probbar
    \state)  \myts \EU(\messg,\state,\Rstrat)}
      \,.
  \end{align*}
By dropping the denominator, which is constant for all $\messg$ for fixed
$\state$, we obtain the above continuous-time formulation.

\subsection{Imitation of success with imprecision}
\label{sec:imit-succ-with}

The above derivation of the replicator dynamic assumes that agents can discriminate choices and
choice points perfectly. Let's dispense with that assumption. With an eye toward sim-max games,
we will assume that states, but not messages, may be confused for one another.\footnote{It is
  relatively straightforward to also incorporate confusability of messages, but this is
  irrelevant to our present purposes.} Confusability of states will affect how agents behave,
how they perceive the behavior of others, and the expected utilities of behavioral
dispositions.

To keep matters simple, let us assume that agents carry pure dispositions to act. Noise can
affect the realization of these strategies. As a sender, every agent maps states to messages:
these are subjectively perceived states, and no longer necessarily also the actually occurring
states. As a receiver, every agent maps messages to state interpretations: these are intended
interpretations that need not always be faithfully realized. This means that behavioral
strategies $\Sstrat$ and $\Rstrat$ represent the average proportions of \emph{actual}
behavioral dispositions in the population, the realization and observation of which can be
distorted by agents' confusion of similar states.

If $\mystate{a}$ is the actual state, let $P_o(\mystate{o} \probbar \mystate{a})$ be the
probability that a given agent observes state $\mystate{o}$. Similarly, if a given receiver
intends to select interpretation $\mystate{i}$, let $P_r(\mystate{r} \probbar \mystate{i})$ be
the probability with which state $\mystate{r}$ is realized. A single round of play of a sim-max
game is then governed by five pieces of stochastic information, where previously there were
only three (see Figures~\ref{fig:sim-max-illustration} and
\ref{fig:imprecise-sim-max-illustration}).

Expected utilities of choices at choice points should likewise take into account that actual states
need not be observed states, and intended interpretations need not be realized
interpretations. First, note that
\begin{align*}
  P_{\overline{o}}(\mystate{a} \probbar \mystate{o}) \propto P_a(\mystate{a}) P_o(\mystate{o}
    \probbar \mystate{a})
\end{align*}
is the probability that $\mystate{a}$ is actual if $\mystate{o}$ is observed by an agent. The
probability that a random sender produces $\messg$ when the actual state is $\mystate{a}$ is:
\begin{align*}
  P_{\Sstrat}(\messg \probbar \mystate{a}) = \sum_{\mystate{o}} P_o(\mystate{o} \probbar
  \mystate{a}) \Sstrat(\messg \probbar \mystate{o})\,. 
\end{align*}
The probability that the actual state is $\mystate{a}$ if a random sender produced $\messg$ is:
\begin{align*}
  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \propto P_a(\mystate{a})
  P_{\Sstrat}(\messg \probbar \mystate{a})\,. 
\end{align*}
The probability that $\mystate{r}$ is realized by a random receiver in response to
message $\messg$ is:
\begin{align*}
  P_{\Rstrat}(\mystate{r} \probbar \messg) = \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
  \mystate{i}) \Rstrat(\mystate{i} \probbar \messg)\,.
\end{align*}
This lets us capture the expected utilities for observed states (sender) and intended
interpretations (receiver) by taking into consideration what the likely actual states and
realized interpretations will be:
\begin{align*}
  \EU(\messg , \mystate{o}, \Rstrat) & = \sum_{\mystate{a}} P_{\overline{o}}(\mystate{a}
  \probbar \mystate{o}) \sum_{\mystate{r}} P_\Rstrat(\mystate{r} \probbar
  \messg) \utils(\mystate{a}, \mystate{r})\,, \\
  \EU(\mystate{i}, \messg, \Sstrat) & = \sum_{\mystate{a}} P_{\overline{\Sstrat}}(\mystate{a}
  \probbar \messg) \sum_{\mystate{r}} P_{r}(\mystate{r} \probbar \mystate{i})
  \utils(\mystate{a}, \mystate{r})\,.
\end{align*}
If the conditional probabilities $P_o$ and $P_r$ are trivial, i.e., assign probability 0 to the
confusability of non-identical states, above definitions reduce to the
previous definitions of expected utilities. This also legitimates the overload of notation.

Presence of potential imprecision in the form of non-trivial $P_o$ and $P_r$ will also affect
the dynamic that ensues from imitation of successes. Since imprecision works slightly
differently on senders and receivers (the former confuse choice points, the latter confuse
choices), we need to look separately at each case.

As before, suppose that senders receive a chance to change their behavior independently for a
given choice point. In the present case, this would be a chance to change how to respond to a
subjectively perceived state $\mystate{o}$, which need not be the actual one. We must then
consult the probability $P_o(\messg \probbar \mystate{o})$ that, given that the learner
observed $\mystate{o}$, he will simultaneously observe a randomly sampled teacher play
$\messg$. This is (with $P_{\overline{o}}$ and $P_{\Sstrat}$ as defined above):
\begin{align*}
  P_o(\messg \probbar \mystate{o}) = \sum_{\mystate{a}} P_{\overline{o}}(\mystate{a}
  \probbar \mystate{o}) \myts P_{\Sstrat}(\messg \probbar \mystate{a})\,.
\end{align*}
The ``inflow'' and ``outflow'' probabilities $P(\messg' \rightarrow \messg, \state)$ and
$P(\messg' \rightarrow \messg, \state)$ that a randomly sampled learner switches from any
$\messg'$ to $\messg$ or from $\messg$ to any $\messg'$ in subjectively perceived state
$\mystate{o}$ are therefore:
\begin{align*}
  P(\messg' \rightarrow \messg, \mystate{o}) & = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
    \mystate{o})}_{\text{learner plays $\messg'$ at $\mystate{o}$}} \cdot
  \underbrace{P_{o}(\messg \probbar \mystate{o})}_{\text{observe teacher play $\messg$}} \cdot
  \underbrace{\EU(\messg, \mystate{o}, \Rstrat)}_{\text{EU teacher choice in learner's view}}
  \\
  P(\messg \rightarrow \messg', \mystate{o}) & = \sum_{\messg'} \underbrace{\Sstrat(\messg \probbar
    \mystate{o})}_{\text{learner plays $\messg$ at $\mystate{o}$}} \cdot
  \underbrace{P_{o}(\messg' \probbar \mystate{o})}_{\text{observe teacher play $\messg'$}} \cdot
  \underbrace{\EU(\messg', \mystate{o}, \Rstrat)}_{\text{EU teacher choice in learner's view}}
\end{align*}
The mean change to the proportion of $\messg$ choices at state $\state$ are then:
\begin{align}
  \label{eq:nRDSender}
  \dot{\Sstrat}(\messg \probbar \mystate{o}) & = P(\messg' \rightarrow m, \mystate{o}) - P(\messg
  \rightarrow m', \mystate{o}) \nonumber \\
  & = \sum_{\messg'} \Sstrat(\messg' \probbar \mystate{o}) \myts P_o(\messg \probbar
  \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \sum_{\messg'} \Sstrat(\messg \probbar \mystate{o}) \myts
  P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat) \nonumber \\
  & = P_o(\messg \probbar \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \Sstrat(\messg \probbar \mystate{o}) \myts \sum_{\messg'} P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat)\,.
\end{align}

The case of the receiver is mostly analogous. Presented with an update opportunity for choice
point $\messg$, a learner will observe a random teacher choose interpretation $\mystate{o}$
with probability:\footnote{We assume that the noise $P_o(\mystate{o} \probbar \mystate{a})$
  that applies to the sender's observations also applies to the receiver's observations when
  trying to imitate other agents' strategies. We could easily make the model more complex and
  introduce another measure of noise for state confusability during receiver attempts to
  imitate. We refrain from it here, because we see no immediate theoretical gain.}
\begin{align*}
  P_o(\mystate{o} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
  \myts P_{\Rstrat}(\mystate{r} \probbar \messg) % = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
  % \myts \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
  % \mystate{i}) \myts \Rstrat(\mystate{i} \probbar \messg)
  \,.
\end{align*}
% The probability $P(\state' \rightarrow \state, \messg)$ that a randomly sampled learner switches
% from using $\state'$ to $\state$ in response to $\messg$ is therefore:
% \begin{align*}
%   P(\state' \rightarrow \state, \messg) = \sum_{\state'} \underbrace{\Rstrat(\state' \probbar
%     \messg)}_{\text{learner plays $\state'$}} \cdot
%   \underbrace{P_{o}(\state \probbar \messg)}_{\text{observe teacher play $\state$}} \cdot
%   \underbrace{\EU(\state, \messg, \Rstrat)}_{\text{EU teacher choice}}
% \end{align*}
Parallel to the sender case, this gives rise to:
\begin{align}
  \label{eq:nRDReceiver}
  \dot{\Rstrat}(\state \probbar \messg) & = \sum_{\state'} \Rstrat(\state' \probbar \messg)
  \myts P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \sum_{\state'}
  \Rstrat(\state \probbar \messg) \myts
  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) \nonumber \\
  & = P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \Rstrat(\state \probbar
  \messg) \myts \sum_{\state'}
  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) 
\end{align}

The continuous-time formulations in Equations~(\ref{eq:nRDSender}) and (\ref{eq:nRDReceiver})
have elegant and practical discrete-time solutions in:
\begin{align*}
  \Sstrat'(\messg \probbar \state) & \propto P_o(\messg \probbar \state) \myts \EU(\messg,
  \state, \Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto P_o(\state \probbar \messg)
  \myts \EU(\state, \messg, \Sstrat)\,,
\end{align*}
which is the discrete-time formulation of the imprecise imitation dynamic given in
Equation~(\ref{eq:nRD}). To see how the discrete-time formulation gives rise to the
continuous-time formulations above, let's assume that update steps are infinitesimally small,
so that, for the sender case:
\begin{align*}
  \dot{\sigma}(\messg \probbar \state) & = \sigma'(\messg \probbar \state) - \sigma(\messg
  \probbar \state)\\
  &= \frac{P_o(\messg \probbar \state) \myts \EU(\messg, \state, \Rstrat) - \Sstrat(\messg
    \probbar \state) \sum_{\messg'} P_o(\messg' \probbar \state) \myts \EU(\messg', \state,
    \Rstrat) }{ \sum_{\messg'} P_o(\messg' \probbar \state) \myts \EU(\messg', \state,
    \Rstrat)}\,.
\end{align*}
As before, we drop the denominator, which is constant for all $\messg$ for fixed $\state$, and
obtain the above continuous-time formulation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\printbibliography[heading=bibintoc]

\end{document}