### 2017-04-20 book_chapter/mf_notes.tex

\documentclass[fleqn,reqno,10pt]{article}

\usepackage{myarticlestyledefault}

\title{Notes on vagueness \& rationality}
\author{Michael Franke}
\date{April 19 2017}

\begin{document}
\maketitle

\section*{Where this is going}

\begin{itemize}
\item The main question to be addressed is:
  \begin{quote}
    How is vagueness in language compatible with rationality and/or evolutionary selection for
    efficient communication?
  \end{quote}
\item The answer I will try to give is something like:
  \begin{quote}
    If whatever we want to call ``language'' is not vague, it is likely quite a useless notion.
  \end{quote}
  With ``quite useless notion'' I do mean ``philosophically'' or ``theoretically useless'', much in
  the vein of Wittgenstein's later views on logical analysis of language: yes, we can do most
  of this, but it shows us little.
\item The way to give this answer is by a more careful analysis of rational or evolutionary
  efficient language use. We look at signaling games and models of rational agency /
  evolutionary selection. It is easy to identify multiple reasons why this idealized picture
  breaks by giving up a number of assumptions, none of which is necessarily that of idealized
  rationality or evolutionary optimization.
\item The main reason for vagueness in language argued for here is this:
  \begin{itemize}
  \item what counts as optimal crisp language use crucially depends on certain
    statistical properties of the world
    \begin{itemize}
    \item language must be optimal in different language games (see below)
      \begin{enumerate}
      \item sim-max games
      \item reference games without displacement
      \item reference games with displacement
      \end{enumerate}
    \item optimal behavior for all of these language games depends on specifics of the game
      situation, most importantly the utilities and the priors about what happens in the real
      world
    \end{itemize}
  \item neither rationality, nor evolutionary pressure can eradicate uncertainty about a
    complex and dynamically changing world
    \begin{itemize}
    \item the main implicit premisse here is that it is not irrational not to be omniscient;
      it's not irrational to have remaining uncertainty after any finite number of
      observations; finiteness of life is not to blame on irrationality either
    \end{itemize}
  \item neither rationality nor evolutionary selection could possibly guarantee an optimal
    crisp language
  \end{itemize}
\item Despite this uncertainty and the vagueness it gives rise to, there seems to be no need to
  postulate an idealized language as an explanatory ingredient of either agents beliefs (about
  language), nor about their actual linguistic practices.
\end{itemize}

\section*{Where do we find the meaning of a language?}

\begin{itemize}
\item There are different places where ``a language's meaning'' might reside:
  \begin{enumerate}
  \item at the \textbf{population-level}; or
  \item at the \textbf{agent-level}, where we might be interested in either:
  \begin{enumerate}
    \item an agent's behavior (language use \& interpretation across contexts); or
    \item her beliefs (about language use, or semantic meaning)
  \end{enumerate}
\end{enumerate}
\item it may seem natural to assume that population-level meaning supervenes on something at
  the agent-level (either behavior, beliefs or both) of individual agents;
  \begin{itemize}
  \item this is putting aside the option that meanings are innate concepts or any such
    esoteric moves
  \end{itemize}
\item We note that there is no generally accepted (formal) criterion for how exactly a
  language's meaning arises from anything more downstream (agent-level stuff, ways of
  interaction, etc.). We have to make do in this vacuum.
  \begin{itemize}
  \item Some attempts to fill the vacuum:
    \citet{Harms2004:Information-and,Harms2010:Determining-tru,Franke2013:An-adaptationis}
  \end{itemize}
\item Although speculative and provocative, we here suggest that everything of observational
  relevance (agents linguistic behavior, possibly their beliefs) can be explained well (i)
  without postulating an idealized non-vague language, and (ii) without abandoning the ideas of
  idealized individual-level rationality or idealized long-run evolutionary selection.
\end{itemize}

\section*{A closer look at an agent's beliefs and behavior}

\begin{itemize}
\item game theoretic models of signaling agents define agent behavior as possibly probabilistic
  functions:
  \begin{itemize}
  \item sender behavior $P_S(\messg \mid \state)$ assigns a probability to the choice of message
    $\messg$ to each state $\state$
  \item receiver behavior $P_R(\act \mid \messg)$ assigns a probability to the choice of
    reaction $\act$ after observing message $\messg$
  \end{itemize}
\item When we fix a context (think: sim-max signaling game), we can define which choices are
  rational. We see that, indeed, no vagueness shows.
  \begin{itemize}
  \item \textcolor{gray}{MF2JPC: I'll skip this here; you did most of this already and its
      clear to us anyway}
  \end{itemize}
\item This is all rather impoverished. You, I, she and he don't just play one game. Utilities
  differ in different situations. We also don't always play sim-max games, but may play a
  reference game. Speakers and listeners may have partial knowledge of the situation and of
  each other. They may also have beliefs about meaning, and their choices would depend on
  these. Recent probabilistic models of pragmatic use explicitly model agents' (beliefs) about
  each other's mental lexica
  \citep{BergenLevy2012:Thats-what-she-,BergenLevy2014:Pragmatic-Reaso,PottsLassiter2016:Embedded-implic}.
  \begin{itemize}
  \item More on dynamic adaptivity to speaker idiolect (Yildirim et al., \dots
    \citet{Davidson_1986:Derangement})
  \end{itemize}
\item The picture of language users which we should minimally adopt is this:
  \begin{itemize}
  \item speaker behavior is a probabilistic function $P_S(\messg \mid \state, B^S_c, B^S_l)$
    specifying a message choice probability for a meaning/state $\state$, a belief after a
    (possibly partial) observation of the context $B_c$ and a belief $B_l$ about the language's
    meaning
  \item listener interpretation behavior is similarly a probabilistic function
    $P_L(\state \mid \messg, B^L_c, B^L_l)$
    \begin{itemize}
    \item speaker and listener beliefs about context and meaning need not be the same; in fact,
      almost always they will be different
    \item the beliefs about contextual parameters $B_c$ which may influence choice behavior are
      bountiful:
      \begin{itemize}
      \item think: shape and parameters of the utility function in a sim-max game; most
        importantly: prior distribution of states (more on this below)
      \item imprecise perception of the referent set in a reference game
      \item \dots
      \end{itemize}
    \item the beliefs $B_l$ about the (lexical \& compositional) meanings of expressions could,
      for simplicity, be captured as a distribution over all possible Boolean lexica, where a
      Boolean lexicon $l$ assigns a subset set of states to each message: $\den{\messg}_l
      \subseteq \States$
      \begin{itemize}
      \item even if we do not want to have agents with ``fancy'' beliefs about meanings, a
        definition of rational choice would require a belief about opponent behavior; in a
        large population, this behavioral belief is good enough for our purposes: it will
        likewise be subject to heavy uncertainty, being obtained from finite (partial)
        observations about a dynamically changing world
      \end{itemize}

    \end{itemize}
  \end{itemize}
\end{itemize}

\section*{Different language games and natural uncertainty}

\begin{itemize}
\item Nothing has been said so far how beliefs are acquired, whether these beliefs are
  rational, how choice are made based on beliefs. There are many options. We cannot look at
  them all. We will browse a few and note how they all naturally give rise to something like
  ``vagueness'' without abandoning rationality.
\item We do this by noting that language games are plenty. We do not just ever engage in a
  single sim-max games (with everlastingly fixed priors, utilities and co-players). 
\item Let us distinguish at least three relevant kinds of games in which gradable properties
  could be used for immediate communication:
  \begin{enumerate}
  \item sim-max games
    \begin{itemize}
    \item clear how they look like
    \item a point which has so far not been stressed (at all? or enough?): the optimal strategy
      is a function of the prior; what counts as ``tall'' depends on (beliefs about) the
      distribution of states/degrees
    \item if priors track, say the actual frequency with which someone is, say, 180cm tall that
      real world frequency is impossible to know precisely
    \item as a result even perfectly rational agents would not be certain about the optimal
      strategy and will (given different life histories) have different strategies
    \item being rational, they would also know that they don't know for certain what the prior
      is and that there is no commonly known and temporally stable prior to begin with
    \end{itemize}
  \item reference games without displacement
    \begin{itemize}
    \item in each trial nature selects a variable number of referents (objects) as the
      \emph{context}; she also selects a designated object for the speaker to refer to; speaker
      and listener both observe the context, but only the speaker observes the designated
      object; speaker and listener obtain a positive payoff if they coordinate on the
      designated referent \citep{Franke2012:Scales-Salience,Franke2012:On-Scales-Salie}
    \item crucially, these could be considered as an infinite collection of games, all
      conditioned on the current context
    \item optimal behavior for these games could be defined as a function of the actual
      context; the problem for even a rational agent is to predict the co-player's behavior,
      for in almost any new trial there will not be a preceding trial in which the exact same
      context has been observed (adding perceptual noise and knowledge of perceptual
      limitations would make this even more pressing)
    \item as a result, even rational agents will be highly uncertain (and aware of their
      uncertainty) as to how language is used in these situations
    \item a rational response will take care of this uncertainty by preferring, if possible,
      expressions whose use conditions are clearer over those that are less clear: borderline
      case uses will be avoided (especially when the language is rich) and underdetermination
      by use creates further uncertainty creates further underdetermination etc.
    \end{itemize}
  \item reference games with displacement
    \begin{itemize}
    \item similar to the above; these games have been considered (informally) by
      \citet{Lipman2009:Why-is-Language} and \citet{Deemter2009:Utility-and-Lan}
    \item the sender observes a referent and describes it for the listener; but unlike before,
      the listener will have to pick out the designated referent in a future context (not known
      by the sender): e.g., picking up a person from the airport, identifying a robber or (from
      the main paper draft) picking up a book from the next room based on the color
    \item as the sender does not know the future context in which the designated referent will
      occur (the other objects co-present in the environment), the sender's optimal choice
      depends crucially on a conjecture about the likelihood of seeing particular contexts
    \item again the sender must needs know the statistical properties of a changing world for
      genuinely optimal crisp language use
    \end{itemize}
  \end{enumerate}
\end{itemize}

\printbibliography[heading=bibintoc]

\newpage

\section*{Snippets}

\begin{itemize}
\item - uncertainty about threshold: - learning from finite examples - rich language makes
  decisions about borderline cases obsolete; enhances problem of learning data scarcity -
  imperfect memory - uncertainty about QUD, utility, interlocutor's beliefs

\item   => why not just have a prior and integrate and have a fixed threshold

\item   1. Priors are non-introspective; sampling based assessment

\item   2. Choice is probabilistic -> utilities are noisy

\item   Why not adopt a fixed threshold? Why is an agent with uncertainty in the lexicon better off?
  => revision cost? surprisal cost?

\item   If a single agent's semantic representation is the outcome of rational learning and possibly
  an internal optimization towards efficient communication.

\item   Finiteness of life is not irrational.

\item   Convince by number of examples of natural sources for vagueness.

\item   Any precise threshold must be a function of priors about properties, QUDs, utilities, and
  interlocutor beliefs. But these are not stable over time. Vagueness might be optimal to
  compensate mistakes in a changing environment. Assumption here: revision cost high when too
  low prob on observable events. (van Deemter example: DECENT behavior)

\item   If speakers' uncertain beliefs come from natural probabilistic evidence, they will not be
  sharp step-functions. If we measure success of communication by KL-divergence or similar, the
  update of a literal listener with a vague threshold would give higher utility. But what about
  pragmatic listeners?

\item   Suppose agents play repeated reference games with different naturals kinds. Each kind has a
  mean and variance for each of several open and closed scale features. Learners entertain
  different hypotheses about lexical meaning (McNally?) They would like to assign a fixed
  degree to an expression because that is economical. They can do that for closed scale natural
  kinds unless their mean and variance is such that exemplars are frequently away from the end
  points. Think wine glasses! They cannot do this for open scales. They must consider a more
  difficult hypothesis, namely to use the prior after all.

\item   Good point by van Deemter: reliable metrics may be absent; the goal of communication may be
  expression of sentiment, and the real value may be assessed in a goal-oriented way (BIG when
  playing soccer not equal to big when painting).

\item   Languages supervene on agents' beliefs about lexica. Lexica are uncertain. They are still
  useful. Lexical uncertainty might even be rational in the light of belief revision in
  changing environments.

\item   Languages are optimized for communication and social interaction. Not necessarily for logical
  consistency.

\item   Plurality of language games forces uncertainty. Wittgenstein.

\item   So then logic breaks, so what? The possibility of visual illusions does not prove the visual
  system ill adapted or suboptimal.
\end{itemize}

\end{document}